---
title: "110 Season-adjusted annual data"
author: "DHJ"
date: "7 5 2020"
output: html_document
---



## 1. Libraries  
```{r}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(forcats)
library(mgcv)

```


## 2. Data  

### Read data
```{r}

dat <- readRDS("Data/100_Chemistry_wide.rds") %>%
  mutate(year = year(sample_date),
         month = month(sample_date),
         day_number = yday(sample_date))

df_deposition <- readRDS("Data/100_Deposition.rds")
df_stations <- readRDS("Data/100_Stations.rds")

station_selection <- readRDS("Data/100_Stations_selection.rds")

```

### Check which columns occur together with NO3   
NH4 and TOTN  
```{r}

df <- dat %>%
  filter(!is.na(NO3))

apply(!is.na(df), 2, sum) %>% sort() %>% rev()

```


## 3. Seasonal patterns  

### Seasonal distribution of data  
```{r}

# Seasonal distribution
gg <- dat %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  ggplot(aes(day_number)) +
  geom_histogram(binwidth = 10)

gg
gg + facet_wrap(vars(country), scales = "free_y")

dat %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  filter(country %in% c("Norway", "Poland", "Slovakia", "Sweden")) %>%
  ggplot(aes(day_number)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(vars(country), scales = "free_y")

# Norway has a peak at 290, Slovakia + Poland 260-270, Sweden 280-290

```
### Check seasonal patterns  
```{r}

df <- dat %>%
  filter(!is.na(NO3)) %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  group_by(country) %>%
  mutate(n = n()) %>%    # pull(n) %>% unique() %>% sort()    # for finding cut-off value = 1000
  filter(n > 1000)

gg <- ggplot(df, aes(x = factor(month), y = NO3)) +
  geom_boxplot() +
  facet_wrap(vars(country)) 

gg

# Zero values
mean(df$NO3 >= 0)  # 99.8%
table(df$NO3 > 0)  # 2634 not over zero
table(df$NO3 == 0) # 2473 equal zero
range(df$NO3)  # -5.6 11200.0

# For including zero values in box plot
tr <- scales::trans_new("", transform = function(x) log(x+0.1), inverse = function(x) exp(x)-0.1)

# Lot with log axis
gg + 
  scale_y_continuous(
    trans = tr, 
    breaks = c(0,1,10, 100, 1000),
    labels = function(lab) round(lab,0)
  )


```


### Test analysis of seasonal patterns   
Using log values  
```{r}
dat <- dat %>%
  mutate(station_id_f = factor(station_id),
         year_f = factor(year))

# Number of obs. per station
df <- dat %>%
  filter(!is.na(NO3) & NO3 > 0) %>%
  count(station_id)

ggplot(df, aes(n)) +
  geom_histogram(binwidth = 50)

# Sample 50 stations with n = 100-500
sample_stations <- df %>%
  filter(n > 100 & n < 500) %>%
  sample_n(50) %>%
  pull(station_id)

dat_sample <- dat %>%
  filter(!is.na(NO3) & station_id %in% sample_stations) %>%
  mutate(log_NO3 = log10(NO3 + 0.1))

m1 <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"), 
           data = dat_sample) 

# summary(m1$gam)
plot(m1)


```

### Seasonal pattern for all stations with at least 50 observations   
I.e., we assume same pattern (on log scale) for all stations  
A minute computing time or so  
```{r}

# Pick stations with n >= 50
pick_stations <- df %>%
  filter(n >= 50) %>%
  pull(station_id)

dat2 <- dat %>% 
  filter(!is.na(NO3) & station_id %in% pick_stations & year >= 1992) %>%
  mutate(log_NO3 = log10(NO3 + 0.1)) %>%
  filter(is.finite(log_NO3))
  
model_season <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"), 
           data = dat2) 

plot(model_season)

```

### Normalize all data to 15. October  
Day 288  
```{r}

yday(ymd("2019-10-15"))

# Add row number (for later left_join)
dat2 <- dat2 %>%
  mutate(row_no = 1:nrow(dat2))

# Predict NO2 given the actual data except that day_number = 288
df_for_predict <- dat2 %>%
  select(station_id_f, year_f, row_no) %>%
  mutate(day_number = 288)
pred <- predict.gam(model_season, newdata = df_for_predict)

# Add back the residuals
df_for_predict <- df_for_predict %>%
  mutate(log_NO3_adj = pred + residuals(model_season))

# Add log_NO3_adj column to our data set
dat2 <- dat2 %>%
  left_join(df_for_predict %>% select(row_no, log_NO3_adj), by = "row_no")
                       
```

### Reproduce plot above using season-adjusted data  
Some residual seasonal variation left - should be looked into?
```{r}

gg <- dat2 %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  ggplot(aes(x = factor(month), y = log_NO3_adj)) +
  geom_boxplot() +
  facet_wrap(vars(country)) 
gg

```

## 4. Data on annual level  
Simply using medians  
```{r}

dat_annual <- dat2 %>%
  group_by(station_id, year) %>%
  summarize(log_NO3_madj = median(log_NO3_adj)) %>%
  ungroup()

```



### Add deposition and metadata 
```{r}

dat_annual <- dat_annual %>%
  left_join(df_deposition, by = c("station_id", "year")) %>%
  left_join(df_stations %>% select(station_id:altitude, country, coniferous, deciduous, 
                                   heathlands, lake, mixed_forest, wetland)) %>%
  mutate(station_id = factor(station_id))


```

## 5. Save  
```{r}

saveRDS(model_season, "Data/110_model_season.rds")
saveRDS(dat_annual, "Data/110_dat_annual_season_adjusted.rds")


```


## 5. Analysis  
```{r}

# Random intercept
m1 <- lme(log_NO3_madj ~ TOTN_dep,
          random = ~1 | station_id, 
          data = dat_annual)
summary(m1)

# Random intercept and slope
m2 <- lme(log_NO3_madj ~ TOTN_dep,
          random = ~1 + TOTN_dep | station_id, 
          data = dat_annual)
summary(m2)

# Random effects
m3 <- lme(log_NO3_madj ~ 1,
          random = ~1 | station_id, 
          data = dat_annual)

summary(m3)


```

