---
title: "110 Season-adjusted annual data"
author: "DHJ"
date: "7 5 2020"
output: html_document
---



## 1. Libraries  
```{r}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(forcats)
library(mgcv)

```


## 2. Data  

### Read data
```{r}

dat <- readRDS("Data/100_Chemistry_wide.rds") %>%
  mutate(year = year(sample_date),
         month = month(sample_date),
         day_number = yday(sample_date))

df_deposition <- readRDS("Data/100_Deposition.rds")
df_stations <- readRDS("Data/100_Stations.rds")

station_selection <- readRDS("Data/100_Stations_selection.rds")

```

### Check which columns occur together with NO3   
NH4 and TOTN  
```{r}

df <- dat %>%
  filter(!is.na(NO3))

apply(!is.na(df), 2, sum) %>% sort() %>% rev()

```


## 3. One common seasonal pattern  

### Seasonal distribution of data  
```{r}

# Seasonal distribution
gg <- dat %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  ggplot(aes(day_number)) +
  geom_histogram(binwidth = 10)

gg
gg + facet_wrap(vars(country), scales = "free_y")

dat %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  filter(country %in% c("Norway", "Poland", "Slovakia", "Sweden")) %>%
  ggplot(aes(day_number)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(vars(country), scales = "free_y")

# Norway has a peak at 290, Slovakia + Poland 260-270, Sweden 280-290

```
### Check seasonal patterns  
```{r}

df <- dat %>%
  filter(!is.na(NO3)) %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  # Pick only countries with >1000 values
  group_by(country) %>%
  mutate(n = n()) %>%    # pull(n) %>% unique() %>% sort()    # for finding cut-off value = 1000
  filter(n > 1000)

gg <- ggplot(df, aes(x = factor(month), y = NO3)) +
  geom_boxplot() +
  facet_wrap(vars(country)) 

gg

# Zero values
mean(df$NO3 >= 0)  # 99.8%
table(df$NO3 > 0)  # 2634 not over zero
table(df$NO3 == 0) # 2473 equal zero
range(df$NO3)  # -5.6 11200.0

# For including zero values in box plot
tr <- scales::trans_new("", transform = function(x) log(x+0.1), inverse = function(x) exp(x)-0.1)

# Lot with log axis
gg + 
  scale_y_continuous(
    trans = tr, 
    breaks = c(0,1,10, 100, 1000),
    labels = function(lab) round(lab,0)
  )


```


### Test analysis of seasonal patterns   
Using log values  
```{r}
dat <- dat %>%
  mutate(station_id_f = factor(station_id),
         year_f = factor(year))

# Number of obs. per station
df <- dat %>%
  filter(!is.na(NO3) & NO3 > 0) %>%
  count(station_id)

ggplot(df, aes(n)) +
  geom_histogram(binwidth = 50)

# Sample 50 stations with n = 100-500
sample_stations <- df %>%
  filter(n > 100 & n < 500) %>%
  sample_n(50) %>%
  pull(station_id)

dat_sample <- dat %>%
  filter(!is.na(NO3) & station_id %in% sample_stations) %>%
  mutate(log_NO3 = log10(NO3 + 0.1))

m1 <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"), 
           data = dat_sample) 

# summary(m1$gam)
plot(m1)


```

### Seasonal pattern for all stations with at least 50 observations   
I.e., we assume same pattern (on log scale) for all stations  
A minute computing time or so  
```{r}

# Pick stations with n >= 50
pick_stations <- df %>%
  filter(n >= 50) %>%
  pull(station_id)

# dat2 = stations with >50 values and after 1992
dat_model <- dat %>% 
  filter(!is.na(NO3) & station_id %in% pick_stations & year >= 1992) %>%
  mutate(log_NO3 = log10(NO3 + 0.1)) %>%
  filter(is.finite(log_NO3))


model_season <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"), 
           data = dat_model) 

plot(model_season)

```

### Extract seasonal effect (day_effect)  
Set day 288 (15. October) as reference date  
```{r}

# summary(model_season)
yday(ymd("2019-10-15"))


# Pick a random station and year
# Which you choose doesn't matter (just try)
station_example <- dat2$station_id_f[10000]
year_example <- 2000

# Data useed for prediction
day_effect <- data.frame(
  station_id_f = station_example,
  year_f = year_example,
  day_number = 1:366)

# Predict
day_effect$log_NO3_pred <- predict.gam(model_season, newdata = day_effect)

# Deviance from day 288 (15. October)
sel <- day_effect$day_number == 288
day_effect$log_NO3_dev <- day_effect$log_NO3_pred - day_effect$log_NO3_pred[sel]

day_effect <- day_effect %>%
  select(day_number, log_NO3_dev)
  
# Plot day effect
ggplot(day_effect, aes(day_number, log_NO3_dev)) +
  geom_path()


```

### Normalize all data to 15. October  (dat2) 
Not only those in 'pick_stations'
```{r}

dat2 <- dat %>% 
  filter(!is.na(NO3) & year >= 1992) %>%
  left_join(day_effect, by = "day_number") %>%
  mutate(log_NO3 = log10(NO3 + 0.1),
         log_NO3_adj = log_NO3 - log_NO3_dev) %>%
  filter(is.finite(log_NO3))

```

### Reproduce plot above using season-adjusted data  
Some residual seasonal variation left - should be looked into?
```{r}

gg <- dat2 %>%
  left_join(df_stations %>% select(station_id, country), by = "station_id") %>%
  # Pick only countries with >850 values
  group_by(country) %>%
  mutate(n = n()) %>%  #  pull(n) %>% unique() %>% sort()    # for finding cut-off value = 850
  filter(n > 850) %>%
  ggplot(aes(x = factor(month), y = log_NO3_adj)) +
  geom_boxplot() +
  facet_wrap(vars(country)) 
gg

```

## 4. Separate pattern by country and altitude  

### Data
```{r}

dat3 <- dat %>%
  filter(!is.na(NO3) & year >= 1992) %>%
  mutate(log_NO3 = log10(NO3 + 0.1)) %>%
  left_join(df_stations %>% select(station_id, country, altitude), by = "station_id")
  
```

### Check patterns  
```{r, fig.height=4, fig.width = 9}

quantile(df_stations$altitude, na.rm = TRUE)

altitude_limits <- c(0, 200, 500, 2400)

df <- dat3 %>%
  mutate(altitude_f = cut(altitude, breaks = altitude_limits)) %>%
  filter(country %in% c("Norway", "Sweden", "Finland", "Germany", "United Kingdom", "Canada", "United States"))

# For including zero values in box plot
tr <- scales::trans_new("", transform = function(x) log(x+0.1), inverse = function(x) exp(x)-0.1)

gg <- ggplot(df, aes(x = factor(month), y = NO3)) +
  geom_boxplot() +
  scale_y_continuous(
    trans = tr, 
    breaks = c(0,1,10, 100, 1000),
    labels = function(lab) round(lab,0)
  ) +
  facet_grid(rows = vars(altitude_f), cols = vars(country))

gg

```

### Function 'seasonal_adjustment'  
Adds columns 'log_NO3_dev' and 'log_NO3_adj' to the data  
```{r}

seasonal_adjustment <- function(data, plot = FALSE){
  data_for_model <- get_data_for_model(data)
  model <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"),
               data = data_for_model) 
  day_effect <- extract_day_effect(model, data_for_model, plot = plot)
  result <- data %>% 
    left_join(day_effect$data, by = "day_number") %>%
    mutate(log_NO3_adj = log_NO3 - log_NO3_dev)
  invisible(
    list(data = result,
         day_effect = day_effect$data,
         ggplot = day_effect$ggplot)
  )
}

# Test
# debugonce(seasonal_adjustment)
# df <- seasonal_adjustment(dat3[1:20000,], plot = TRUE)$data


get_data_for_model <- function(data){
  data %>%
    group_by(station_id) %>%
    mutate(n = n()) %>%
    ungroup() %>%
    filter(n > 50) %>%
    mutate(station_id_f = factor(station_id),
           year_f = factor(year)) 
}

# Test
# dat3 %>% nrow()
# get_data_for_model(dat3) %>% nrow()

extract_day_effect <- function(model, data, plot = FALSE){
  
  station_example <- data$station_id_f[1]
  year_example <- 1992
  
  # Data useed for prediction
  day_effect <- data.frame(
    station_id_f = station_example,
    year_f = year_example,
    day_number = 1:366)
  
  # Predict
  day_effect$log_NO3_pred <- predict.gam(model, newdata = day_effect)
  
  # Deviance from day 288 (15. October)
  sel <- day_effect$day_number == 288
  day_effect$log_NO3_dev <- day_effect$log_NO3_pred - day_effect$log_NO3_pred[sel]
  
  day_effect <- day_effect %>%
    select(day_number, log_NO3_dev)
  
  # Plot day effect
  gg <- ggplot(day_effect, aes(day_number, log_NO3_dev)) +
    geom_path()

  if (plot){
    print(gg)
  }
    
  invisible(list(data = day_effect,
                 ggplot = gg)
  )
  
}
# Test
# df <- get_data_for_model(dat3[1:20000,])
# model <- gam(log_NO3 ~ station_id_f + year_f + s(day_number, bs = "cc"),
#               data = df) 
# extract_day_effect(model, df, plot = TRUE)


```

### Select data for modelling  
```{r, fig.width = 9, fig.height = 7}

altitude_limits <- c(0, 200, 500, 2400)

dat3 <- dat3 %>%
  mutate(altitude_f = cut(altitude, breaks = altitude_limits)) 

# Set missing leve,s to "(missing)"
dat3$altitude_f <- fct_explicit_na(dat3$altitude_f)

# Check missing altitude by coutry
xtabs(~is.na(altitude) + country, dat3)
# altitude = NA for Estonia and 2/3 of Ireland's lakes
# Estonia's highest point is 318 masl
# Ireland's highest point is 1038 masl

# Set all Ireland and Estonia lakes to (0,200]
sel <- dat3$country == "Ireland"
table(dat3$altitude_f[sel])
dat3$altitude_f[sel] <- "(0,200]"

sel <- dat3$country == "Estonia"
table(dat3$altitude_f[sel])
dat3$altitude_f[sel] <- "(0,200]"

# Filter so we keep only countries/altitude with a range of >= 9 months
#  with n >= 20 observations per month
# These data will be adjusted for season
dat_for_models1 <- dat3 %>%
  group_by(country, altitude_f, month) %>%
  mutate(n_per_month = n()) %>%
  filter(n_per_month >= 20) %>%
  ungroup() %>%
  group_by(country, altitude_f) %>%
  mutate(month_range = diff(range(month))) %>%
  filter(month_range >= 9) %>%
  ungroup()

# Filter so we keep only stations with n >= 50 
# These will be used for making the seasonal model for adjustment
dat_for_models2 <- dat_for_models1 %>%
  group_by(country, altitude_f) %>%
  mutate(n = n()) %>%
  filter(n >= 50) 

dat_for_models1 %>%
  ggplot(aes(month)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(vars(country, altitude_f), scales = "free_y")

nrow(dat3)
nrow(dat_for_models1)
nrow(dat_for_models2)  # which happens to be all of them

```

### Seasonal adjustment for each country/altitude - test 
```{r}

dat_for_models2_test <- dat_for_models2 %>% filter(country == "Norway")


# Split data seet intto a list
dat_for_models2_split <- dat_for_models2_test %>%
  group_by(country, altitude_f) %>%
  mutate(label = paste(country, altitude_f, sep = "_")) %>%
  group_split()

# Set names for the list
names(dat_for_models2_split) <- dat_for_models2_split %>% map_chr(~.$label[1])

# Make "safe" version of seasonal_adjustment
# => Continues looping even if seasonal_adjustment() fails
seasonal_adjustment_safe <- safely(seasonal_adjustment)

# Run adjustments
dat3_adjustment_list <- dat_for_models2_split %>%
  map(~seasonal_adjustment_safe(.)) %>%
  transpose()

# Structure: a list of 2 objects: 
# 1) list of results, 2) list of errors (NULL if analysis went ok)
str(dat3_adjustment_list, 1)

# Which adjustments did not retun an error?
adjustment_ok <- dat3_adjustment_list$error %>% map_lgl(is.null)

# Set names for result
names(dat3_adjustment_list$result) <- names(dat_for_models2_split)

# Check contents of $result
# str(dat3_adjustment_list$result, 1)
# str(dat3_adjustment_list$result[[1]], 1)

# Extract the "data" object and put together to a data frame
dat3_adjusted <- dat3_adjustment_list$result[adjustment_ok] %>% 
  map_df(~.$data)
# str(dat3_adjusteddata_list, 1)

# Plot resulting seasonal curves (version 1)
dat3_adjustment_list$result %>% 
  walk(~print(.$ggplot + labs(title = .$data$label[1])))

# Plot resulting seasonal curves (version 2)
dat3_day_effects <- dat3_adjustment_list$result %>% 
  map_df(function(X){
    df <- X$day_effect
    df %>% mutate(group = X$data$label[1])
    })

ggplot(dat3_day_effects, aes(day_number, log_NO3_dev)) +
  geom_path() +
  facet_wrap(vars(group))

```


### Seasonal adjustment for each country/altitude   
  
Calculate adjustments  
```{r}



# Split data seet intto a list
dat_for_models2_split <- dat_for_models2 %>%
  group_by(country, altitude_f) %>%
  mutate(label = paste(country, altitude_f, sep = "_")) %>%
  group_split()

# Set names for the list
names(dat_for_models2_split) <- dat_for_models2_split %>% map_chr(~.$label[1])

# Make "safe" version of seasonal_adjustment
# => Continues looping even if seasonal_adjustment() fails
seasonal_adjustment_safe <- safely(seasonal_adjustment)

# Run adjustments
t0 <- Sys.time()
dat3_adjustment_list <- dat_for_models2_split %>%
  map(~seasonal_adjustment_safe(.)) %>%
  transpose()

Sys.time() - t0

# Structure: a list of 2 objects: 
# 1) list of results, 2) list of errors (NULL if analysis went ok)
str(dat3_adjustment_list, 1)
```


Put results together 
```{r}

# Which adjustments did not retun an error?
adjustment_ok <- dat3_adjustment_list$error %>% map_lgl(is.null)

# Set names for result
names(dat3_adjustment_list$result) <- names(dat_for_models2_split)

# Check contents of $result
# str(dat3_adjustment_list$result, 1)
# str(dat3_adjustment_list$result[[1]], 1)

# Extract the "data" object and put together to a data frame
dat3_adjusted <- dat3_adjustment_list$result[adjustment_ok] %>% 
  map_df(~.$data)
# str(dat3_adjusteddata_list, 1)

# Make data set for seasonal curves used
dat3_day_effects <- dat3_adjustment_list$result %>% 
  map_df(function(X){
    df <- X$day_effect
    df %>% mutate(group = X$data$label[1])
    })

```

### Plot seasonal curves  
```{r}

ggplot(dat3_day_effects, aes(day_number, log_NO3_dev)) +
  geom_path() +
  facet_wrap(vars(group))

```

## 5. Data on annual level  
Simply using medians  
```{r}

dat_annual <- dat3_adjusted %>%
  group_by(station_id, year) %>%
  summarize(log_NO3_madj = median(log_NO3_adj)) %>%
  ungroup()

```



### Add deposition and metadata 
```{r}

dat_annual <- dat_annual %>%
  left_join(df_deposition, by = c("station_id", "year")) %>%
  left_join(df_stations %>% select(station_id:altitude, country, coniferous, deciduous, 
                                   heathlands, lake, mixed_forest, wetland)) %>%
  mutate(station_id = factor(station_id))


```

## 5. Save  
```{r}

saveRDS(model_season, "Data/110_model_season.rds")
saveRDS(dat_annual, "Data/110_dat_annual_season_adjusted.rds")


```


## 5. Analysis  
```{r}

# Random intercept
m1 <- lme(log_NO3_madj ~ TOTN_dep,
          random = ~1 | station_id, 
          data = dat_annual)
summary(m1)

# Random intercept and slope
m2 <- lme(log_NO3_madj ~ TOTN_dep,
          random = ~1 + TOTN_dep | station_id, 
          data = dat_annual)
summary(m2)

# Random effects
m3 <- lme(log_NO3_madj ~ 1,
          random = ~1 | station_id, 
          data = dat_annual)

summary(m3)


```

