---
title: "120_Time_series_modelling"
author: "DHJ"
date: "8 5 2020"
output: html_document
---


## 1. Libraries  
```{r}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(forcats)
library(mgcv)
library(nlme)

```


## 2. Data  

### Read data
```{r}

dat_annual <- readRDS("Data/110_dat_annual_season_adjusted.rds") %>%
  mutate(station_id = levels(station_id)[as.numeric(station_id)])  # remove factor 

str(dat_annual)

dat_climate <- readRDS("Data/100_Climate.rds")

station_selection <- readRDS("Data/100_Stations_selection.rds")
station_selection_no3 <- station_selection$trends_1992_2016_no3

```

### Add climate data  
```{r}

dat_climate_wide <- dat_climate %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(station_id = as.character(station_id))

nrow(dat_annual)

dat_annual <- dat_annual %>%
  left_join(dat_climate_wide, by = c("station_id", "year"))

nrow(dat_annual)  # check that row number hasn't changed

```

### Pick data  
```{r}

dat_annual_sel <- dat_annual %>%
  filter(station_id %in% station_selection_no3)

nrow(dat_annual_sel)

```


## 3. Time series analyses with gamm    
For each station separately

### Test with one time series
```{r}

dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[1])

mod <- list(
  gamm(log_NO3_madj ~ s(year, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(TOTN_dep, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(pre, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(tmp, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(pre, k = 3) + s(tmp, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(pre, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(tmp, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(pre, k = 3) + s(tmp, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(TOTN_dep ~ s(year, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(pre ~ s(year, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test),
  gamm(tmp ~ s(year, k = 3), 
            correlation = corAR1(form =∼ year),
            data = dat_annual_test)
  )

anova <- anova(mod[[1]]$lme, 
               mod[[2]]$lme,
               mod[[3]]$lme,
               mod[[4]]$lme,
               mod[[5]]$lme,
               mod[[6]]$lme,
               mod[[7]]$lme,
               mod[[8]]$lme)
anova

model_selected <- which.min(anova$AIC)
summary(mod[[model_selected]]$gam)

aic <- data.frame(Model = 1:8,
                  AIC = anova$AIC,
                  dAIC = anova$AIC - min(anova$AIC))

aic

par(mfrow = c(2,2), mar = c(4,5,2,1))
plot(mod[[1]]$gam)
plot(mod[[2]]$gam)
plot(mod[[3]]$gam)

par(mfrow = c(2,3), mar = c(4,5,2,1))
plot(mod[[5]]$gam)
plot(mod[[6]]$gam)
plot(mod[[7]]$gam)

# X <- plot(mod[[3]]$gam)
# str(X)

```

### All time series with TOC     
Ca 4 minutes  
```{r, cache = TRUE}

run_gamm_analyses <- function(data){
  list(
    gamm(log_NO3_madj ~ s(year, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(TOTN_dep, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(pre, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(tmp, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(pre, k = 3) + s(tmp, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(pre, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(tmp, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(log_NO3_madj ~ s(TOTN_dep, k = 3) + s(pre, k = 3) + s(tmp, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(TOTN_dep ~ s(year, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(pre ~ s(year, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data),
    gamm(tmp ~ s(year, k = 3), 
         correlation = corAR1(form =∼ year),
         data = data)
  )
}
# Test
# mod <- run_gamm_analyses(dat_annual_test)

# Safe version
run_gamm_analyses_safe <- safely(run_gamm_analyses)


# Split data
dat_annual_split <- dat_annual_sel %>%
  group_by(station_id) %>%
  group_split()

# Name the split data
names(dat_annual_split) <- dat_annual_split %>% map_chr(~.$station_id[1])

# Run all models  
t0 <- Sys.time()
gamm_model_list <- dat_annual_split %>%
  map(run_gamm_analyses_safe) %>%
  transpose()
Sys.time() - t0  

# Set names for result
names(gamm_model_list$error) <- names(dat_annual_split)
names(gamm_model_list$result) <- names(dat_annual_split)


```
## 4. Visual check of 'log_NO3_madj'    

### Check some random stations   
Some stations have extremely low 2006 values  
```{r}

# length(unique(dat_annual_sel$station_id))  # 398 stations  
i <- 1 
dat_annual_test$country[i]

dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[i])

plot(dat_annual_test[c("log_NO3_madj", "TOTN_dep", "pre", "tmp", "year")])

dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[i]) %>%
  ggplot(aes(year, log_NO3_madj)) +
  geom_line() +
  geom_point()

```


### Check time seres for each country 
```{r}

countries <- unique(dat_annual_sel$country)
# length(countries)  # 13 countries

for (cntr in countries){
  
  gg <- dat_annual_sel %>%
    filter(country %in% cntr) %>%
    ggplot(aes(year, log_NO3_madj, group = station_id)) +
    geom_line() +
    geom_point() +
    coord_cartesian(ylim = range(dat_annual_sel$log_NO3_madj, na.rm = TRUE)) +  # same y limit for all plots
    labs(title = cntr)
  
  print(gg)
}

```
### Clean some of the data  
```{r}

cat("Number of rows:", nrow(dat_annual_sel), "\n")

# CH
sel <- with(dat_annual_sel, country == "Switzerland" & year == 2006); sum(sel)
dat_annual_sel <- dat_annual_sel[!sel,]
# nrow(dat_annual_sel)

# D
sel <- with(dat_annual_sel, country == "Germany" & log_NO3_madj < 1); sum(sel)
dat_annual_sel <- dat_annual_sel[!sel,]

cat("Number of rows:", nrow(dat_annual_sel), "\n")

```

## 5. Time series analyses with gls (linear effects)        
For each station separately

```{r}

# length(unique(dat_annual_sel$station_id))  # 398 stations  
i <- 1 
dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[i])

plot(dat_annual_test[c("log_NO3_madj", "TOTN_dep", "pre", "tmp", "year")])

# https://stats.stackexchange.com/a/367973/13380
# https://m-clark.github.io/mixed-models-with-R/extensions.html#residual-structure
library(nlme)

mod1a = lm(
  log_NO3_madj ~ year,
  data = dat_annual_test
  )
mod1b = lme(
  log_NO3_madj ~ year,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)

summary(mod1a)
summary(mod1b)

# install.packages("effects")
# install.packages("estimability")
library(effects)
eff1a <- predictorEffects(mod1a)
eff1b <- predictorEffects(mod1b)
plot(eff1a)
plot(eff1b)

```


### Test with one time series
```{r}

dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[1])

mod <- list(
  lme(log_NO3_madj ~ year, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ TOTN_dep, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ pre, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ tmp, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ pre + tmp, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ TOTN_dep + pre, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ TOTN_dep + tmp, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(log_NO3_madj ~ TOTN_dep + pre + tmp, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(TOTN_dep ~ year, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(pre ~ year, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test),
  lme(tmp ~ year, 
      random = ~ 1 | station_id,   
      correlation = corAR1(form = ~ year | station_id),
      data = dat_annual_test)
)

anova <- anova(mod[[1]], 
               mod[[2]],
               mod[[3]],
               mod[[4]],
               mod[[5]],
               mod[[6]],
               mod[[7]],
               mod[[8]])
anova

model_selected <- which.min(anova$AIC)
summary(mod[[model_selected]]$gam)

aic <- data.frame(Model = 1:8,
                  AIC = anova$AIC,
                  dAIC = anova$AIC - min(anova$AIC))

aic


```


### All time series with TOC     
Ca 1.5 minute    
```{r, cache = TRUE}

# Changed from gls to lme because gls is supported by 'effects' package  
run_lme_analyses <- function(data){
  list(
      lme(log_NO3_madj ~ year, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ TOTN_dep, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ pre, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ tmp, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ pre + tmp, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ TOTN_dep + pre, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ TOTN_dep + tmp, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(log_NO3_madj ~ TOTN_dep + pre + tmp, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(TOTN_dep ~ year, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(pre ~ year, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data),
      lme(tmp ~ year, 
          random = ~ 1 | station_id,   
          correlation = corAR1(form = ~ year | station_id),
          data = data)
    )
}

# Test
# mod <- run_lme_analyses(dat_annual_test)

# Safe version
run_lme_analyses_safe <- safely(run_lme_analyses)


# Split data
dat_annual_split <- dat_annual_sel %>%
  group_by(station_id) %>%
  group_split()

# Name the split data
names(dat_annual_split) <- dat_annual_split %>% map_chr(~.$station_id[1])

# Run all models  
t0 <- Sys.time()
lme_model_list <- dat_annual_split %>%
  purrr::map(run_lme_analyses_safe) %>%
  transpose()
Sys.time() - t0  

# Set names for result
names(lme_model_list$error) <- names(dat_annual_split)
names(lme_model_list$result) <- names(dat_annual_split)

```


## 6. Save
```{r}

saveRDS(dat_annual_sel, "Data/120_dat_annual_sel.rds")
saveRDS(gamm_model_list, "Data/120_ts_model_list.rds")
saveRDS(lme_model_list, "Data/120_ts_model_list_linear.rds")

```


