---
title: "130_Time_series_results"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true

---



## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(forcats)
library(mgcv)
library(nlme)

library(MuMIn)
# install.packages("randomForest")
library(randomForest)

library(maps)
my_map <- map_data("world")


```


## 2. Data
```{r}

dat_annual_sel <- readRDS("Data/120_dat_annual_sel.rds")
ts_model_list <- readRDS("Data/120_ts_model_list.rds")
# ts_model_list_wout_TOC <- readRDS("Data/120_ts_model_list_wout_TOC.rds")

df_stations <- readRDS("Data/100_Stations.rds")
df_deposition <- readRDS("Data/100_Deposition.rds")
df_climate <- readRDS("Data/100_Climate.rds")

```



## 3. Make 'result_list'     

adjustment_ok
```{r}


# Which time series did not retun an error?
model_ok <- ts_model_list$error %>% map_lgl(is.null)
table(model_ok)

# Length of result: 4 if TOC is not included, 8 if TOC is included
result_length <- ts_model_list$result %>% map_int(length)

# Check contents of $result
# $result[[1]][1:3] %>% str(2)

result_list <- ts_model_list$result[model_ok]


```

### Check one station  
```{r}

st <- 1

# Station ID ( = station_ID in df_stations, df_deposition, df_climate...) 
names(result_list)[st]

# All models
str(result_list[[st]], 1)

# Model 1 - log_NO3_madj ~ year 
summary(result_list[[st]][[1]]$gam)

# Model 6 - log_NO3_madj ~ Deposition + precipitation  
summary(result_list[[st]][[6]]$gam)

# Model 9 - Deposition ~ year      
summary(result_list[[1]][[9]]$gam)

```


## 4. Explote AIC values 
All models:  
  c("Time", "Dep", "P", "T", "P+T", "Dep+P", "Dep+T", "Dep+P+T", "Dep_vs_time", "P_vs_time", "T_vs_time")

### Extract dAIC for NO3 models   
Not including the last three  
```{r}

model_labels <- 
  c("Time", "Dep", "P", "T", "P_T", "Dep_P", "Dep_T", "Dep_P_T")


get_anova <- function(model_list){
  anova(model_list[[1]]$lme, 
        model_list[[2]]$lme, 
        model_list[[3]]$lme,
        model_list[[4]]$lme,
        model_list[[5]]$lme, 
        model_list[[6]]$lme, 
        model_list[[7]]$lme,
        model_list[[8]]$lme
        )
}


# get_anova(result_list[[1]])

get_aic <- function(model_list){
  aic <- get_anova(model_list)$AIC
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

get_aic_without_time <- function(model_list){
  aic <- get_anova(model_list)$AIC[-1]
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels[-1]
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

# get_aic(result_list[[1]])

daic <- result_list %>% map_df(get_aic, .id = "station_id")
daic_without_time <- result_list %>% map_df(get_aic, .id = "station_id")



```

### Show all AICs  
```{r}

df <- daic %>%
  select(station_id, Time:Dep_P_T) %>%
  tidyr::pivot_longer(Time:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df_without_time <- daic_without_time %>%
  select(station_id, Dep:Dep_P_T) %>%
  tidyr::pivot_longer(Dep:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(tilte = "Best model")

df_without_time %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(tilte = "Best model, not including the 'time' model")

ggplot(df, aes(Model, dAIC)) +
  geom_boxplot()

ggplot(df, aes(Model, station_id, fill = dAIC)) +
  geom_raster()

```

### Check how often adding climate to deposition improves the model  
Improves model if the point is under the line  
- Conclusion: not so often  
```{r}

ggplot(daic, aes(Dep, Dep_P)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_P_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)


```


## 5. Rsq of the regression nitrate as function of deposition (model 2)


### Extract R-sq's   
Models  
- model 2: effect of deposition on NO3
- model 9: time trend in deposition   
```{r}

# summary(result_list[[1]][[2]]$gam)$r.sq

rsq_no3_vs_dep <- result_list %>% 
  map_dbl(~summary(.[[2]]$gam)$r.sq)
rsq_dep_vs_time <- result_list %>% 
  map_dbl(~summary(.[[9]]$gam)$r.sq)

rsq <- tibble(
  station_id = names(rsq_no3_vs_dep),
  rsq_no3_vs_dep,
  rsq_dep_vs_time
)

# rsq


```

### Plot R-sqs   
Interpretaton:  
- Upper right = models with strong time trend in deposition and strong effect of deposition on NO3   
- Lower right = models with strong time trend in deposition but nevertheless small effect of deposition on NO3  
- Lower left = models with weak time trend in deposition and (for that reason) small effect of deposition on NO3  
  
Dotted lines are used in part 6:  
- time series to the left of vertical dotted line are not used (low variation in deposition)  
- time series to the right of vertical dotted line: 
    * R-sq of deposition on NO3 (vertical scale) used in GAM and random forest regression
    * Classification by horizontal lines used in random forest classification

```{r}

rsq %>%
  mutate(
    rsq_no3_vs_dep = case_when(
      rsq_no3_vs_dep < 0 ~ 0,
      rsq_no3_vs_dep >= 0 ~ rsq_no3_vs_dep)
  ) %>%
  ggplot(aes(rsq_dep_vs_time, rsq_no3_vs_dep)) + 
  geom_point() +
  geom_hline(yintercept = c(0.2, 0.4), linetype = "dashed") +
  geom_vline(xintercept = 0.7, linetype = "dashed")
  

```

### Examples

#### Strong time trend in deposition (Rsq 0.91), strong effect of deposition on NO3 (Rsq 0.64)
```{r, fig.height=4, fig.width=8}

id <- "38299"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```


#### Strong time trend in deposition (Rsq 0.8), clear but weaker effect of deposition on NO3 (Rsq 0.4) 
```{r, fig.height=4, fig.width=8}

id <- "38090"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```

#### Strong time trend in deposition (0.9), little effect of deposition on NO3 (Rsq 0.17)
```{r, fig.height=4, fig.width=8}

id <- "23478"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```

#### Weak time trend in deposition (Rsq 0.35), weak effect of deposition on NO3 
```{r, fig.height=4, fig.width=8}

id <- "38478"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```



## 6. What separates stations with little/strong effect of deposition on NO3
Analysis of time series with relatively strong time trend in deposition   

### Select cases  
```{r}

rsq <- rsq %>%
  mutate(Dep_effect = 
           case_when(
             rsq_dep_vs_time < 0.7 ~ "Expected_low",
             rsq_no3_vs_dep < 0.20 ~ "Low_effect",
             rsq_no3_vs_dep < 0.40 ~ "Medium_effect",
             rsq_no3_vs_dep >= 0.40 ~ "High_effect")
  ) %>%
  mutate(station_id = as.numeric(station_id))

table(rsq$Dep_effect)

```

### Data   

#### Check land cover NAs
```{r}

df_lacking_landcover <- df_stations %>% select(urban:other)
df_lacking_landcover <- is.na(df_lacking_landcover)

sel_lacking_all <- apply(df_lacking_landcover, 1, mean) == 1
# df_stations[sel_lacking_all,] %>% View()

apply(df_lacking_landcover[!sel_lacking_all,], 2, sum)

# df_stations %>% filter(is.na(grasslands)) %>% View()

```


#### Make data set 'data_select'
Adding mean deposition and mean precipitation + temeprature   
- Exclude time series with "Expected_low" - i.e., weak time series trends in deposition  
```{r}

df_deposition_mean <-
  df_deposition %>% 
  group_by(station_id) %>% 
  summarise(Mean_dep = mean(TOTN_dep))

df_climate_mean <- df_climate %>%
  group_by(station_id, variable) %>% 
  summarise(mean = mean(value)) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mean", names_prefix = "Mean_")


# df_stations %>% filter(is.na(coniferous))
# apply(is.na(df_stations), 2, mean)

df_stations2 <- df_stations %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier) %>%
  select(station_id, 
         coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse) %>%
  filter(!is.na(coniferous))


data_select <- df_stations2 %>% 
  left_join(df_deposition_mean, by  = "station_id") %>%
  left_join(df_climate_mean, by  = "station_id") %>%
  left_join(rsq, by  = "station_id") %>%
  # Exclude time series with "Expected_low" - i.e., weak time series trends in deposition
  filter(Dep_effect %in% c("Low_effect", "Medium_effect", "High_effect") ) %>%
  select(station_id, Dep_effect, rsq_no3_vs_dep, coniferous:Mean_tmp) %>%
  mutate(
    rsq_no3_vs_dep = case_when(
      rsq_no3_vs_dep < 0 ~ 0,
      rsq_no3_vs_dep >= 0 ~ rsq_no3_vs_dep),
    Dep_effect = factor(Dep_effect, 
                             levels = c("Low_effect", "Medium_effect", "High_effect")
                             ),
    coniferous = log(coniferous + 1), 
    deciduous = log(deciduous + 1), 
    lake = log(lake + 1), 
    mixed_forest = log(mixed_forest + 1), 
    wetland = log(wetland + 1),
    bare_sparse = log(bare_sparse + 1)
    ) %>%
  as.data.frame() 
# %>%
#   mutate(Dep_effect2 = case_when(
#     Dep_effect == "Low_effect" ~ 1,
#     Dep_effect == "Medium_effect" ~ 2,
#     Dep_effect == "High_effect" ~ 3)
#   ) %>% 
#   select(-Dep_effect)

rownames(data_select) <- data_select$station_id
data_select$station_id <- NULL

# Check missing values 
apply(is.na(data_select), 2, sum)

# 
table(addNA(data_select$Dep_effect))

# For making mutate code
# x1 <- data_select %>% names()
# x2 <- paste0(
#   paste0(x, " = "), paste0("log(", x, " + 1)")
# )
# x2 %>% paste(collapse = ", ")


```

### Multiple GAM  


#### Plot 1
```{r}
ggplot(data_select, aes(Dep_effect, rsq_no3_vs_dep)) +
  geom_boxplot() +
  geom_jitter()

```

#### Plot 2
```{r, fig.width=9, fig.height=9}

plot(data_select %>% select(-Dep_effect))

```

#### Data
```{r}

data_mumin <- data_select %>% 
  select(-Dep_effect)

# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# saveRDS(data_mumin, "Data/130_data_mumin.rds")

```


#### Find best models
```{r}

# For making formula
# x <- data_select %>% names()
# paste0("s(", x, ", k = 3)") %>% paste(collapse = " + ")

# data_mumin <- readRDS("Data/130_data_mumin.rds")

# Check for NAs
# apply(is.na(data_mumin), 2, sum)

# Check mean values
# apply(data_mumin, 2, mean)

mod_complete <- gam(rsq_no3_vs_dep ~ 
             s(coniferous, k = 3) + s(deciduous, k = 3) + s(lake, k = 3) + 
             s(mixed_forest, k = 3) + s(wetland, k = 3) +  + s(bare_sparse, k = 3) +
             s(Mean_dep, k = 3) + s(Mean_pre, k = 3) + s(Mean_tmp, k = 3), 
           data = data_mumin,
           na.action = "na.fail")
# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# summary(mod_complete)
# dd <- dredge(mod_complete)
# saveRDS(dd, "Data/130_mumin_object.rds")

dd <- readRDS("Data/130_mumin_object.rds")

```

#### Best models
```{r}

subset(dd, delta < 2)

```

#### Plot 3 best models  
PLus ANOVA table for best model;  
```{r, fig.width=9, fig.height = 7}
#'Best' models

# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# best_models <- list(
#   get.models(dd, 1)[[1]],
#   get.models(dd, 2)[[1]],
#   get.models(dd, 3)[[1]],
#   get.models(dd, 4)[[1]]
# )
# saveRDS(best_models, "Data/130_best_models.rds")

best_models <- readRDS("Data/130_best_models.rds")

summary(best_models[[1]])

par(mfrow = c(4,4), mar = c(4,5,2,1))
plot(best_models[[1]], res = TRUE, pch = 1)
par(mfg = c(2,1))
plot(best_models[[2]], res = TRUE, pch = 1)
par(mfg = c(3,1))
plot(best_models[[3]], res = TRUE, pch = 1)
par(mfg = c(4,1))
plot(best_models[[4]], res = TRUE, pch = 1)

```



### Random forest, classification

#### Split into training and validation data
```{r}

# data_select$Dep_effect2 <- as.factor(data_select$Dep_effect2)

set.seed(123)

x <- runif(nrow(data_select))
train <- ifelse(x < 0.8, TRUE, FALSE)

train_set <- data_select[train,] %>% select(-rsq_no3_vs_dep)
valid_set <- data_select[!train,] %>% select(-rsq_no3_vs_dep) 

```

#### Analysis
```{r}

model1 <- randomForest(Dep_effect ~ ., 
                       data = train_set, 
                       mtry = 4,
                       importance = TRUE)

model1

```

#### Predict on training data
```{r}

# Predicting on train set
pred_valid <- predict(model1, valid_set, type = "class")
# Checking classification accuracy
table(pred_valid, valid_set$Dep_effect)  

```

#### Importance  
High MeanDecreaseGini = high importance in model  
```{r}

randomForest::importance(model1)
varImpPlot(model1)

```

#### Plot partial effects (standard)   
```{r, fig.height=8, fig.width=8}

imp <- randomForest::importance(model1)
impvar1 <- rownames(imp)[order(imp[, 2], decreasing=TRUE)]
saveRDS(impvar1, "Data/130_impvar1.rmd")
# impvar1 <- readRDS("Data/130_impvar1.rmd")

par(mfrow=c(3,3))

for (i in seq_along(impvar1)) {
  partialPlot(model1, train_set, impvar1[i], xlab=impvar1[i],
              main=paste("Effect of", impvar1[i]),
              ylim=c(0, 1))
}


if (FALSE){
  
  
  # dir.create("Figures")

  png("Figures/Random_forest_class_partial.png", height = 17, width = 17, units = 'cm', 
     res = 300, type="cairo", antialias="default")

  par(mfrow=c(3,3), mar = c(5,4,2,1))
  
  for (i in seq_along(impvar1)) {
    partialPlot(model1, train_set, impvar1[i], xlab=impvar1[i],
              main=paste("Effect of", impvar1[i]),
                ylim=c(0.18, 0.36))
    
  }
  
  dev.off()
  
    
}

```

#### Partial plots - functions   
```{r}

# Random sample of variable number 'var_no'
random_sample <- function(var_no, 
                          data = train_set, 
                          n = 20) {
  base::sample(data[[var_no]], n, replace = TRUE)
}
# random_sample(2)

# Make variable going from min to max
make_focus_variable <- function(pred_data,
                                var, 
                                actual_data = train_set){
  n <- nrow(pred_data)
  pred_data[[var]] <- seq(min(actual_data[[var]]), max(actual_data[[var]]), length = n)
  pred_data
}

# Returns data for plot
# - where the 'focus variable' var is set to name 'x_var' and varies from min to max
# - the otehr variables 
plot_effect_get_data_class <- function(var, 
                                       data = train_set,
                                       model = model1,
                                       var_nos = 2:ncol(data), 
                                       n_categories = 12, n = 1000){
  
  sample_for_predict <- var_nos %>% 
    purrr::map_dfc(random_sample, n = n) %>%
    set_names(names(train_set)[var_nos]) %>%
    make_focus_variable(var = var)
  
  # Make predicted data
  sample_for_predict$Pred_class <- predict(model, sample_for_predict, type = "class")
  sample_for_predict$i <- 1:nrow(sample_for_predict)
  
  sample_for_predict <- sample_for_predict %>%
    mutate(Group = cut(i, breaks = n_categories))
  
  # Changes name of 'focus variable' to name 'x_var'
  sel <- names(sample_for_predict) == var
  names(sample_for_predict)[sel] <- "x_var"
  
  sample_for_predict_grouped <- sample_for_predict %>%
    group_by(Group) %>%
    summarize(x_var = mean(x_var),
              Low_effect = mean(Pred_class == "Low_effect"),
              Medium_effect = mean(Pred_class == "Medium_effect"),
              High_effect = mean(Pred_class == "High_effect")
    ) %>%
    pivot_longer(Low_effect:High_effect, names_to = "Class", values_to = "Fraction") %>%
    mutate(Class = forcats::fct_inorder(Class))
  
  sample_for_predict_grouped
}
# Test
# plot_effect_get_data("coniferous")

plot_effect_class <- function(var, 
                              data = train_set,
                              model = model1,
                              var_nos = 2:ncol(data), 
                              n_categories = 12, n = 1000,
                              backtransform = FALSE){
  
  sample_for_predict_grouped <- plot_effect_get_data_class(
    var = var,
    data = data,
    model = model,
    var_nos = var_nos, 
    n_categories = n_categories, 
    n = n)
  
  colors <- RColorBrewer::brewer.pal(n = 5, name = "YlOrRd")
  
  df <- sample_for_predict_grouped %>%
    mutate(Class = forcats::fct_rev(Class),)
  
  if (backtransform){
    df <- df %>%
      mutate(x_var = factor(round(exp(x_var) - 1,1)))
      xlabel <- paste0(var, ", percent")
  } else {
    df <- df %>%
      mutate(x_var = factor(round(x_var, 1)))
      xlabel <- paste0(var, ", log(percent + 1)")
  }
  ggplot(df, 
         aes(x = x_var, y = Fraction, fill = Class)) +
    geom_col() +
    scale_fill_manual(values = rev(colors)) +
    labs(x = xlabel)
  
} 

```



#### Partial plots
```{r}

# dir.create("Figures/Random_forest_class")
# dir.create("Figures/Random_forest_regr")

for (i in 1:length(impvar1)){
  var <- impvar1[i]
  # show log-transformed numbers (a)
  fn <- sprintf("Figures/Random_forest_class/Random_forest_class_%02ia_%s.png", i, var)
  gg <- plot_effect_class(var, data = train_set, model = model1)
  ggsave(fn, gg, height = 4, width = 7)
  # Original numbers (b)
  if (!var %in% c("Mean_tmp", "Mean_pre", "Mean_dep")){
    
    fn <- sprintf("Figures/Random_forest_class/Random_forest_class_%02ib_%s.png", i, var)
    gg <- plot_effect_class(var, data = train_set, model = model1, 
                           backtransform = TRUE)
    ggsave(fn, gg, height = 4, width = 7)
  }
} 

```

#### Partial plot for bare_sparse  
```{r}
plot_effect_class("bare_sparse", backtransform = TRUE)
```




### Random forest, regression  

#### Split into training and validation data
```{r}

# data_select$Dep_effect2 <- as.factor(data_select$Dep_effect2)

set.seed(123)

x <- runif(nrow(data_select))
train <- ifelse(x < 0.8, TRUE, FALSE)

train_set <- data_select[train,] %>% select(-Dep_effect)
valid_set <- data_select[!train,] %>% select(-Dep_effect) 

```

#### Analysis
```{r}

model2 <- randomForest(rsq_no3_vs_dep ~ ., 
                       data = train_set, 
                       mtry = 4,
                       importance = TRUE)

model2


```

#### Actual vs Predicted, training set
```{r}
plot(model2$pred, train_set$rsq_no3_vs_dep)

# R-sq
summary(lm(train_set$rsq_no3_vs_dep ~ model2$pred))$r.sq

```

#### Predict on training data
```{r}

# ?predict.randomForest

# Predicting on train set
pred_valid <- predict(model2, valid_set, type = "response")
# Checking classification accuracy
plot(pred_valid, valid_set$rsq_no3_vs_dep)  
abline(0, 1, lty = 2)

```


#### Importance
High MeanDecreaseGini = high importance in model  
```{r}

randomForest::importance(model2)
varImpPlot(model2)

png("Figures/Random_forest_regr_importance.png", height = 14, width = 25, units = 'cm', 
    res = 300, type="cairo", antialias="default")

varImpPlot(model2)

dev.off()


```



#### Plot partial effects (standard plot)
```{r, fig.height=8, fig.width=8}

imp <- randomForest::importance(model2)
impvar2 <- rownames(imp)[order(imp[, 1], decreasing=TRUE)]
saveRDS(impvar2, "Data/130_impvar2.rmd")
# impvar2 <- readRDS("Data/130_impvar2.rmd")

op = par(mfrow=c(3,3))

for (i in seq_along(impvar2)) {
  partialPlot(model2, train_set, impvar2[i], xlab=impvar2[i],
              main=paste("Effect of", impvar2[i]),
              ylim=c(0.18, 0.36))
}

if (FALSE){
  
  
  # dir.create("Figures")

  png("Figures/Random_forest_regr_partial.png", height = 17, width = 17, units = 'cm', 
     res = 300, type="cairo", antialias="default")

  par(mfrow=c(3,3), mar = c(5,4,2,1))
  
  for (i in seq_along(impvar2)) {
    partialPlot(model2, train_set, impvar2[i], xlab=impvar2[i],
              main=paste("Effect of", impvar2[i]),
                ylim=c(0.18, 0.36))
    
  }
  
  dev.off()
  
    
}
```


#### Partial plots - functions   
```{r}

# random_sample() - AS ABOVE
# make_focus_variable() - AS ABOVE <- function(pred_data,

# Returns data for plot
# - where the 'focus variable' var is set to name 'x_var' and varies from min to max
# - the otehr variables 

# type  = "class" for classification, "response" for regression
plot_effect_get_data_regr <- function(var, 
                                       data = train_set,
                                       model = model2,
                                       var_nos = 2:ncol(data), 
                                       n_categories = 50, n = 2000){
  
  sample_for_predict <- var_nos %>% 
    purrr::map_dfc(random_sample, n = n) %>%
    set_names(names(data)[var_nos]) %>%
    make_focus_variable(var = var)
  
  # Make predicted data
  sample_for_predict$Pred_response <- predict(model, sample_for_predict, type = "response")
  sample_for_predict$i <- 1:nrow(sample_for_predict)
  
  sample_for_predict <- sample_for_predict %>%
    mutate(Group = cut(i, breaks = n_categories))
  
  # Changes name of 'focus variable' to name 'x_var'
  sel <- names(sample_for_predict) == var
  names(sample_for_predict)[sel] <- "x_var"

  sample_for_predict_grouped <- sample_for_predict %>%
    group_by(Group) %>%
    summarize(x_var = mean(x_var),
              Pred_response = mean(Pred_response)
    )
  sample_for_predict_grouped
}


# plot_effect_get_data_regr("bare_sparse", data = train_set, model2)
# 
# plot_effect_get_data_regr("bare_sparse", data = train_set, model2,
#                           n_categories = 50, n = 2000)

plot_effect_regr <- function(var, 
                             data = train_set,
                             model = model2,
                             backtransform = FALSE,
                             var_nos = 2:ncol(data), 
                             n_categories = 12, n = 1000){
  
  sample_for_predict_regr <- plot_effect_get_data_regr(
    var = var, 
    data = data,
    model = model,
    var_nos = var_nos, 
    n_categories = n_categories, 
    n = n)
  
  colors <- RColorBrewer::brewer.pal(n = 5, name = "YlOrRd")
  
  df <- sample_for_predict_regr
  
  if (backtransform){
    df <- df %>%
      mutate(x_var = round(exp(x_var) - 1,1))
      xlabel <- paste0(var, ", percent")
  } else {
      xlabel <- paste0(var, ", log(percent + 1)")
  }
  ggplot(df, 
         aes(x = x_var, y = Pred_response)) +
    geom_smooth(span = 0.3) +
    geom_point() +
    labs(x = xlabel)
  
} 

```

#### Partial plots
```{r}

for (i in 1:length(impvar2)){
  var <- impvar2[i]
  # show log-transformed numbers (a)
  fn <- sprintf("Figures/Random_forest_regr/Random_forest_regr_%02ia_%s.png", i, var)
  gg <- plot_effect_regr(var, data = train_set, model = model2)
  ggsave(fn, gg, height = 4, width = 7)
  # Original numbers (b)
  if (!var %in% c("Mean_tmp", "Mean_pre", "Mean_dep")){
    fn <- sprintf("Figures/Random_forest_regr/Random_forest_regr_%02ib_%s.png", i, var)
    gg <- plot_effect_regr(var, data = train_set, model = model2, 
                           backtransform = TRUE)
    ggsave(fn, gg, height = 4, width = 7)
  }
} 


```

#### Partial plot for bare_sparse
```{r}

plot_effect_regr("bare_sparse", data = train_set, model = model2,
                          n_categories = 50, n = 2000)
plot_effect_regr("bare_sparse", data = train_set, model = model2,
                          n_categories = 50, n = 2000,
                 backtransform = TRUE)

```






