---
title: "130_Time_series_results"
author: "DHJ"
date: "8 5 2020"
output: 
  github_document:
    toc: true
  html_document:
    toc: true    
    toc_float: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })    
  
---



## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(forcats)
library(mgcv)
library(nlme)

library(MuMIn)
# install.packages("randomForest")
library(randomForest)

library(maps)
my_map <- map_data("world")


```


## 2. Data
```{r}

dat_annual_sel <- readRDS("Data/120_dat_annual_sel.rds")
ts_model_list <- readRDS("Data/120_ts_model_list.rds")
# ts_model_list_wout_TOC <- readRDS("Data/120_ts_model_list_wout_TOC.rds")

df_stations <- readRDS("Data/100_Stations.rds")
df_deposition <- readRDS("Data/100_Deposition.rds")
df_climate <- readRDS("Data/100_Climate.rds")

```



## 3. Make 'result_list'     

adjustment_ok
```{r}


# Which time series did not retun an error?
model_ok <- ts_model_list$error %>% map_lgl(is.null)
table(model_ok)

# Length of result: 4 if TOC is not included, 8 if TOC is included
result_length <- ts_model_list$result %>% map_int(length)

# Check contents of $result
# $result[[1]][1:3] %>% str(2)

result_list <- ts_model_list$result[model_ok]


```


## 4. Explote AIC values 
All models:  
  c("Time", "Dep", "P", "T", "P+T", "Dep+P", "Dep+T", "Dep+P+T", "Dep_vs_time", "P_vs_time", "T_vs_time")


### Extract dAIC for NO3 models   
Not including the last three  
```{r}

model_labels <- 
  c("Time", "Dep", "P", "T", "P_T", "Dep_P", "Dep_T", "Dep_P_T")


get_anova <- function(model_list){
  anova(model_list[[1]]$lme, 
        model_list[[2]]$lme, 
        model_list[[3]]$lme,
        model_list[[4]]$lme,
        model_list[[5]]$lme, 
        model_list[[6]]$lme, 
        model_list[[7]]$lme,
        model_list[[8]]$lme
        )
}


# get_anova(result_list[[1]])

get_aic <- function(model_list){
  aic <- get_anova(model_list)$AIC
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

get_aic_without_time <- function(model_list){
  aic <- get_anova(model_list)$AIC[-1]
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels[-1]
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

# get_aic(result_list[[1]])

daic <- result_list %>% map_df(get_aic, .id = "station_id")
daic_without_time <- result_list %>% map_df(get_aic, .id = "station_id")



```

### Show all AICs  
```{r}

df <- daic %>%
  select(station_id, Time:Dep_P_T) %>%
  tidyr::pivot_longer(Time:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df_without_time <- daic_without_time %>%
  select(station_id, Dep:Dep_P_T) %>%
  tidyr::pivot_longer(Dep:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(tilte = "Best model")

df_without_time %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(tilte = "Best model, not including the 'time' model")

ggplot(df, aes(Model, dAIC)) +
  geom_boxplot()

ggplot(df, aes(Model, station_id, fill = dAIC)) +
  geom_raster()

```

### Check how often adding climate to deposition improves the model  
Improves model if the point is under the line  
- Conclusion: not so often  
```{r}

ggplot(daic, aes(Dep, Dep_P)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_P_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)


```


## 5. Rsq of the regression nitrate as function of deposition (model 2)


### Extract R-sq's   
Models  
- model 2: effect of deposition on NO3
- model 9: time trend in deposition   
```{r}

# summary(result_list[[1]][[2]]$gam)$r.sq

rsq_no3_vs_dep <- result_list %>% 
  map_dbl(~summary(.[[2]]$gam)$r.sq)
rsq_dep_vs_time <- result_list %>% 
  map_dbl(~summary(.[[9]]$gam)$r.sq)

rsq <- tibble(
  station_id = names(rsq_no3_vs_dep),
  rsq_no3_vs_dep,
  rsq_dep_vs_time
)

# rsq


```

### Plot R-sqs   
Interpretaton:  
- Upper right = models with strong time trend in deposition and strong effect of deposition on NO3   
- Lower right = models with strong time trend in deposition but nevertheless small effect of deposition on NO3  
- Lower left = models with weak time trend in deposition and (for that reason) small effect of deposition on NO3  
  
Dotted lines are used in part 6:  
- time series to the left of vertical dotted line are not used (low variation in deposition)  
- time series to the right of vertical dotted line: 
    * R-sq of deposition on NO3 (vertical scale) used in GAM and random forest regression
    * Classification by horizontal lines used in random forest classification

```{r}

rsq %>%
  mutate(
    rsq_no3_vs_dep = case_when(
      rsq_no3_vs_dep < 0 ~ 0,
      rsq_no3_vs_dep >= 0 ~ rsq_no3_vs_dep)
  ) %>%
  ggplot(aes(rsq_dep_vs_time, rsq_no3_vs_dep)) + 
  geom_point() +
  geom_hline(yintercept = c(0.2, 0.4), linetype = "dashed") +
  geom_vline(xintercept = 0.7, linetype = "dashed")
  

```

### Examples

#### Strong time trend in deposition (Rsq 0.91), strong effect of deposition on NO3 (Rsq 0.64)
```{r, fig.height=4, fig.width=8}

id <- "38299"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```


#### Strong time trend in deposition (Rsq 0.8), clear but weaker effect of deposition on NO3 (Rsq 0.4) 
```{r, fig.height=4, fig.width=8}

id <- "38090"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```

#### Strong time trend in deposition (0.9), little effect of deposition on NO3 (Rsq 0.17)
```{r, fig.height=4, fig.width=8}

id <- "23478"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```

#### Weak time trend in deposition (Rsq 0.35), weak effect of deposition on NO3 
```{r, fig.height=4, fig.width=8}

id <- "38478"

filter(rsq, station_id == id)

par(mfrow = c(1,2), mar = c(4,5,2,1))
plot(result_list[[id]][[2]]$gam, res = TRUE, pch = 1)
plot(result_list[[id]][[9]]$gam, res = TRUE, pch = 1)

```



## 6. What separates stations with little/strong effect of deposition on NO3
Analysis of time series with relatively strong time trend in deposition   

### Select cases  
```{r}

rsq <- rsq %>%
  mutate(Dep_effect = 
           case_when(
             rsq_dep_vs_time < 0.7 ~ "Expected_low",
             rsq_no3_vs_dep < 0.20 ~ "Low_effect",
             rsq_no3_vs_dep < 0.40 ~ "Medium_effect",
             rsq_no3_vs_dep > 0.40 ~ "High_effect")
  ) %>%
  mutate(station_id = as.numeric(station_id))

table(rsq$Dep_effect)

```

### Data   
  
Adding mean deposition and mean precipitation + temeprature   
- Exclude time series with "Expected_low" - i.e., weak time series trends in deposition  

```{r}

df_deposition_mean <-
  df_deposition %>% 
  group_by(station_id) %>% 
  summarise(Mean_dep = mean(TOTN_dep))

df_climate_mean <- df_climate %>%
  group_by(station_id, variable) %>% 
  summarise(mean = mean(value)) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mean", names_prefix = "Mean_")


# df_stations %>% filter(is.na(coniferous))

df_stations2 <- df_stations %>%
  select(station_id, 
         coniferous, deciduous, lake, mixed_forest, wetland) %>%
  filter(!is.na(coniferous))
apply(is.na(df_stations2), 2, sum)


data_select <- df_stations2 %>% 
  left_join(df_deposition_mean, by  = "station_id") %>%
  left_join(df_climate_mean, by  = "station_id") %>%
  left_join(rsq, by  = "station_id") %>%
  # Exclude time series with "Expected_low" - i.e., weak time series trends in deposition
  filter(Dep_effect %in% c("Low_effect", "Medium_effect", "High_effect") ) %>%
  select(station_id, Dep_effect, rsq_no3_vs_dep, coniferous:Mean_tmp) %>%
  mutate(
    rsq_no3_vs_dep = case_when(
      rsq_no3_vs_dep < 0 ~ 0,
      rsq_no3_vs_dep >= 0 ~ rsq_no3_vs_dep),
    Dep_effect = factor(Dep_effect, 
                             levels = c("Low_effect", "Medium_effect", "High_effect")
                             ),
    coniferous = log(coniferous + 1), deciduous = log(deciduous + 1), lake = log(lake + 1), 
    mixed_forest = log(mixed_forest + 1), wetland = log(wetland + 1)
    ) %>%
  as.data.frame() 
# %>%
#   mutate(Dep_effect2 = case_when(
#     Dep_effect == "Low_effect" ~ 1,
#     Dep_effect == "Medium_effect" ~ 2,
#     Dep_effect == "High_effect" ~ 3)
#   ) %>% 
#   select(-Dep_effect)

rownames(data_select) <- data_select$station_id
data_select$station_id <- NULL

# Check missing values 
apply(is.na(data_select), 2, sum)

# 
table(addNA(data_select$Dep_effect))

# For making mutate code
# x1 <- data_select %>% names()
# x2 <- paste0(
#   paste0(x, " = "), paste0("log(", x, " + 1)")
# )
# x2 %>% paste(collapse = ", ")


```

### Multiple GAM  


#### Plot 
```{r, fig.height=7}

ggplot(data_select, aes(Dep_effect, rsq_no3_vs_dep)) +
  geom_boxplot() +
  geom_jitter()

plot(data_select %>% select(-Dep_effect))

```

#### Data
```{r}

data_mumin <- data_select %>% 
  select(-Dep_effect)

# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# saveRDS(data_mumin, "Data/130_data_mumin.rds")

```


#### Find best models
```{r}

# For making formula
# x <- data_select %>% names()
# paste0("s(", x, ", k = 3)") %>% paste(collapse = " + ")

data_mumin <- readRDS("Data/130_data_mumin.rds")

# Check for NAs
# apply(is.na(data_mumin), 2, sum)

# Check mean values
# apply(data_mumin, 2, mean)

mod_complete <- gam(rsq_no3_vs_dep ~ 
             s(coniferous, k = 3) + s(deciduous, k = 3) + s(lake, k = 3) + 
             s(mixed_forest, k = 3) + s(wetland, k = 3) + 
             s(Mean_dep, k = 3) + s(Mean_pre, k = 3) + s(Mean_tmp, k = 3), 
           data = data_mumin,
           na.action = "na.fail")
# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# summary(mod_complete)
# dd <- dredge(mod_complete)
# saveRDS(dd, "Data/130_mumin_object.rds")
dd <- readRDS("Data/130_mumin_object.rds")

```

#### Best models
```{r}

subset(dd, delta < 2)

```

#### Plot 3 best models  
PLus ANOVA table for best model;  
```{r, fig.width=9, fig.height = 7}
#'Best' models

# RUN AND SAVED ONLY INTERACTIVELY - because this fails when knitting Rmd documnet
# best_models <- list(
#   get.models(dd, 1)[[1]],
#   get.models(dd, 2)[[1]],
#   get.models(dd, 3)[[1]]
# )
# saveRDS(best_models, "Data/130_best_models.rds")

best_models <- readRDS("Data/130_best_models.rds")

summary(best_models[[1]])$s.table

par(mfrow = c(3,4), mar = c(4,5,2,1))
plot(best_models[[1]], res = TRUE, pch = 1)
par(mfg = c(2,1))
plot(best_models[[2]], res = TRUE, pch = 1)
par(mfg = c(3,1))
plot(best_models[[3]], res = TRUE, pch = 1)

```



### Random forest, classification

#### Split into training and validation data
```{r}

# data_select$Dep_effect2 <- as.factor(data_select$Dep_effect2)

set.seed(123)

x <- runif(nrow(data_select))
train <- ifelse(x < 0.8, TRUE, FALSE)

train_set <- data_select[train,] %>% select(-rsq_no3_vs_dep)
valid_set <- data_select[!train,] %>% select(-rsq_no3_vs_dep) 

```

#### Analysis
```{r}

model1 <- randomForest(Dep_effect ~ ., 
                       data = train_set, 
                       mtry = 4,
                       importance = TRUE)

model1

```

#### Predict on training data
```{r}

# Predicting on train set
pred_valid <- predict(model1, valid_set, type = "class")
# Checking classification accuracy
table(pred_valid, valid_set$Dep_effect)  

```

#### Importance  
High MeanDecreaseGini = high importance in model  
```{r}

randomForest::importance(model1)
varImpPlot(model1)

```

### Random forest, regression  

#### Split into training and validation data
```{r}

# data_select$Dep_effect2 <- as.factor(data_select$Dep_effect2)

set.seed(123)

x <- runif(nrow(data_select))
train <- ifelse(x < 0.8, TRUE, FALSE)

train_set <- data_select[train,] %>% select(-Dep_effect)
valid_set <- data_select[!train,] %>% select(-Dep_effect) 

```

#### Analysis
```{r}

model2 <- randomForest(rsq_no3_vs_dep ~ ., 
                       data = train_set, 
                       mtry = 4,
                       importance = TRUE)

model2

```

#### Importance
High MeanDecreaseGini = high importance in model  
```{r}

randomForest::importance(model2)
varImpPlot(model2)

```

