---
title: "130_Time_series_results"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true

---



## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of these packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(forcats)
library(mgcv)
library(nlme)
library(mapview)
library()

library(MuMIn)
# install.packages("randomForest")
library(randomForest)

library(maps)
my_map <- map_data("world")

library(effects)    # handles lme models  
library(readxl)

knitr::opts_chunk$set(results = 'hold')
knitr::opts_chunk$set(warning = FALSE)

```


## 2. Data
```{r}

dat_annual_sel <- readRDS("Data/120_dat_annual_sel.rds")
ts_model_list <- readRDS("Data/120_ts_model_list_linear.rds")
# ts_model_list_wout_TOC <- readRDS("Data/120_ts_model_list_wout_TOC.rds")

df_stations <- readRDS("Data/100_Stations.rds")
df_deposition <- readRDS("Data/100_Deposition.rds")
df_climate <- readRDS("Data/100_Climate.rds")

```

### Station metadata
```{r}

df_station <- read_excel(
  "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_2020_05_04.xlsx") %>%
  mutate(station_id = as.character(station_id))

```

## . 
## HAR KOMMI HIT
## .

## 3. Make 'result_list'     

adjustment_ok
```{r}


# Which time series did not retun an error?
model_ok <- ts_model_list$error %>% map_lgl(is.null)
table(model_ok)

# Length of result: 4 if TOC is not included, 8 if TOC is included
result_length <- ts_model_list$result %>% map_int(length)

# Check contents of $result
# $result[[1]][1:3] %>% str(2)

result_list <- ts_model_list$result[model_ok]


```

### Check one station  
```{r}

st <- 1

# Station ID ( = station_ID in df_stations, df_deposition, df_climate...) 
names(result_list)[st]

# All models
# str(result_list[[st]], 1)

# Model 1 - log_NO3_madj ~ year 
# summary(result_list[[st]][[1]]$gam)

# Model 6 - log_NO3_madj ~ Deposition + precipitation  
# summary(result_list[[st]][[6]]$gam)

# Model 9 - Deposition ~ year      
# summary(result_list[[1]][[9]]$gam)

```


## 4. Explote AIC values 
All models:  
  c("Time", "Dep", "P", "T", "P+T", "Dep+P", "Dep+T", "Dep+P+T", "Dep_vs_time", "P_vs_time", "T_vs_time")

### Extract dAIC for NO3 models   
Not including the last three  
```{r, warning=FALSE}

model_labels <- 
  c("Time", "Dep", "P", "T", "P_T", "Dep_P", "Dep_T", "Dep_P_T")


get_anova <- function(model_list){
  anova(model_list[[1]], 
        model_list[[2]], 
        model_list[[3]],
        model_list[[4]],
        model_list[[5]], 
        model_list[[6]], 
        model_list[[7]],
        model_list[[8]]
        )
}
# Test:
# get_anova(result_list[[1]])

# get_anova(result_list[[1]])

get_aic <- function(model_list){
  aic <- get_anova(model_list)$AIC
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

get_aic_without_time <- function(model_list){
  aic <- get_anova(model_list)$AIC[-1]
  daic <- aic - min(aic)
  df <- daic %>% matrix(nrow = 1) %>% data.frame() 
  names(df) <- model_labels[-1]
  data.frame(df, aic_min = min(aic), aic_best = which.min(aic))
}

# get_aic(result_list[[1]])

daic <- result_list %>% map_df(get_aic, .id = "station_id")
daic_without_time <- result_list %>% map_df(get_aic, .id = "station_id")



```

### Show all AICs  
```{r, fig.width=5, fig.height=3}

df <- daic %>%
  select(station_id, Time:Dep_P_T) %>%
  tidyr::pivot_longer(Time:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df_without_time <- daic_without_time %>%
  select(station_id, Dep:Dep_P_T) %>%
  tidyr::pivot_longer(Dep:Dep_P_T, names_to = "Model", values_to = "dAIC") %>%
  mutate(Model = forcats::fct_inorder(Model))

df %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(title = "Best model")

df_without_time %>%
  filter(dAIC == 0) %>%
  ggplot(aes(Model)) +
  geom_histogram(stat = "count") +
  labs(tilte = "Best model, not including the 'time' model")

ggplot(df, aes(Model, dAIC)) +
  geom_boxplot()

ggplot(df, aes(Model, station_id, fill = dAIC)) +
  geom_raster()

```

### Check how often adding climate to deposition improves the model  
Improves model if the point is under the line  
- Conclusion: in this case (linear model): almost never
```{r}

ggplot(daic, aes(Dep, Dep_P)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)
ggplot(daic, aes(Dep, Dep_P_T)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)


```

### Examples  
```{r}


i <- 1 


dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[i])

mod_no3_year = lme(
  log_NO3_madj ~ year,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)
mod_no3_tmp = lme(
  log_NO3_madj ~ tmp,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)

mod_tmp_year = lme(
  tmp ~ year,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)

summary(mod_no3_year)
# summary(result_list[[i]][[1]])
summary(mod_no3_tmp)

# Get slope (estimate and SE) 
# coef(summary(result_list[[i]][[1]]))[2, 1:2]

eff1 <- predictorEffects(mod_no3_year)
eff2 <- predictorEffects(mod_no3_tmp)
eff3 <- predictorEffects(mod_tmp_year)
plot(eff1)
plot(eff2)
plot(eff3)

```

## 5. Slope estimates   

### Extract slope estimates   
Models  
- model 2: effect of deposition on NO3
- model 9: time trend in deposition   
```{r}

# summary(result_list[[1]][[2]]$gam)$r.sq

slope_no3_vs_dep <- result_list %>% 
  map_dbl(~coef(summary(.[[1]]))[2, 1])

slope_dep_vs_time <- result_list %>% 
  map_dbl(~coef(summary(.[[9]]))[2, 1])

slope_1 <- tibble(
  station_id = names(slope_no3_vs_dep),
  slope_no3_vs_dep,
  slope_dep_vs_time
)


names(dat_annual_sel) %>% paste(collapse = ", ")

df_meta <- dat_annual_sel %>%
  as_tibble() %>%
  group_by(station_id, country) %>%
  summarise(across(c(TOTN_dep, latitude, longitude, catchment_area, altitude, 
                     coniferous, deciduous, heathlands, lake,
                     mixed_forest, wetland, pre, tmp), mean)) %>%
  left_join(df_station[c("station_id", "region")], by = "station_id")


slope <- slope_1 %>%
  left_join(df_meta)
# slope


```
## 6. Plot estimates  

```{r, fig.width=8, fig.height=8}

ggplot(slope, aes(slope_dep_vs_time, slope_no3_vs_dep, color = country)) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) 
  
ggplot(slope, aes(slope_dep_vs_time, slope_no3_vs_dep)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")

slope %>%
  filter(!country %in% c("Latvia","Ireland","Italy","Netherlands")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_dep)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")


```



## 7. Plot all on map    
```{r, fig.width = 9, fig.height=9}

# slope <- dplyr::sample_n(slope, 50)
  
slope_sf <- sf::st_as_sf(
  slope,
  coords = c("longitude", "latitude"), 
  crs = "+proj=longlat +ellps=WGS84", 
  agr = "constant") 

# Classify 'slope_no3_vs_dep' -> 'slope_no3_vs_dep_class'
slope_sf <- slope_sf %>%
  mutate(slope_no3_vs_dep_class = 
           cut(slope_no3_vs_dep, breaks = c(seq(-0.05, 0.02, 0.01), 0.11)))

mapview(slope_sf, zcol = "slope_no3_vs_dep_class")

```



## 8. Analysis by region    

### Plot by region, N America    
```{r, fig.width=8, fig.height=3.5}

slope %>%
  filter(country %in% c("Canada", "United States")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_dep, color = region)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2)

```

### USA, without land cover   
```{r, results = 'hold'}

slope %>%
  filter(country %in% "United States") %>%
  xtabs(~is.na(coniferous), .)

df <- slope %>%
  filter(country %in% "United States") %>%
  select(station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, pre, tmp)
df <- df[complete.cases(df),]
# colnames(slope)

cat("\nNumber of rows:", nrow(df), "\n")

options(na.action = "na.fail")
fm1 <- lm(slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + pre + tmp,
          data = df)
dd <- dredge(fm1)

# coniferous, deciduous, heathlands, lake, mixed_forest, wetland, heathlands + lake + mixed_forest #+ wetland + pre + tmp
                                                 
```
#### Showing the best models (dAIC < 2)  
```{r}
subset(dd, delta < 2)

```

```{r, fig.width=8, fig.height=3}
par(mfrow = c(1,3), mar = c(4,5,2,1))
visreg::visreg(get.models(dd, 1)[[1]])

```


```{r}

df_sel <- slope %>%
  filter(country %in% "United States")
df_sel$station_id[1]

dat_annual_test <- dat_annual_sel %>%
  filter(station_id %in% unique(dat_annual_sel$station_id)[i])

mod_no3_year = lme(
  log_NO3_madj ~ year,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)
mod_no3_tmp = lme(
  log_NO3_madj ~ tmp,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)

mod_tmp_year = lme(
  tmp ~ year,
  data = dat_annual_test,
  random = ~ 1 | station_id,
  correlation = corAR1(form = ~ year | station_id)
)



```




### USA + Canada  , incl land cover   
Excluding 'catchment_area' as it lacks from 1/3 of the data    
```{r}

slope %>%
  filter(country %in% c("United States", "Canada")) %>%
  xtabs(~is.na(coniferous), .)

slope %>%
  filter(country %in% c("United States", "Canada")) %>%
  xtabs(~is.na(catchment_area), .)   # lacking for 37 stations (ca 1/3)

df <- slope %>%
  filter(country %in% "United States") %>%
  select(station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, 
         pre, tmp,
         coniferous, deciduous, heathlands, lake, mixed_forest, wetland)
df <- df[complete.cases(df),]


cat("\nNumber of rows:", nrow(df), "\n")

# colnames(slope)
# plot(select(df, -station_id))
  
               
```

#### Including latitude, longitude   
```{r, results='hold'}
options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + 
                     pre + tmp +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
dd1 <- dredge(fm)

subset(dd1, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd1, 1)[[1]])$r.sq

```

#### Excluding latitude, longitude  
```{r, results  = "hold"}
options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time +  
                     pre + tmp +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
dd2 <- dredge(fm)

subset(dd2, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd2, 1)[[1]])$r.sq

```


#### Best models incl/excl latitude and longitude  
```{r, fig.width=8, fig.height=2.5}

par(mfrow = c(1,4), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(1,4), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 1)[[1]])
mtext("Best model excl latitude and longitude", outer = TRUE)

```



### Nordics , incl land cover   

#### Plot by region        
```{r, fig.width=6, fig.height=5}

slope %>%
  filter(country %in% c("Norway", "Sweden", "Finland", "Latvia")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_dep, color = region)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2)

```


#### Prepare data  
```{r}

slope %>%
  filter(country %in% c("Norway", "Sweden", "Finland", "Latvia")) %>%
  xtabs(~is.na(coniferous), .)

slope %>%
  filter(country %in% c("Norway", "Sweden", "Finland", "Latvia")) %>%
  xtabs(~is.na(catchment_area), .)   # lacking for none

df <- slope %>%
  filter(country %in% c("Norway", "Sweden", "Finland", "Latvia")) %>%
  filter(slope_no3_vs_dep <= 0.04)  %>%   # exclude one Lativa station with very high value
  select(station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, 
         pre, tmp, catchment_area,
         coniferous, deciduous, heathlands, lake, mixed_forest, wetland)
df <- df[complete.cases(df),]

cat("\n\nNumber of rows:", nrow(df), "\n")
# colnames(slope)

# plot(select(df, -station_id))
  
               
```

#### Including latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + 
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
# dd1 <- dredge(fm)                           # only once
# saveRDS(dd1, "Data/140_Nordics_dd1.rds")    # save it as it takes a couple of minutes
dd1 <- readRDS("Data/140_Nordics_dd1.rds")

subset(dd1, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd1, 1)[[1]])$r.sq

```

#### Excluding latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + 
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
dd2 <- dredge(fm)

subset(dd2, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd2, 1)[[1]])$r.sq

```


#### Best models incl/excl latitude and longitude  
```{r, fig.width=9, fig.height=2.2}

par(mfrow = c(1,5), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(1,5), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 1)[[1]])
mtext("Best model excl latitude and longitude", outer = TRUE)

```


#### Best models incl/excl latitude and longitude  
```{r, fig.width=8, fig.height=2.5}

par(mfrow = c(1,4), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(1,4), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 1)[[1]])
mtext("Best model excl latitude and longitude", outer = TRUE)

```


### Europe except Nordics , excl land cover   

#### Plot by region        
```{r, fig.width=6, fig.height=5}

slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_dep, color = region)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2)

```


#### Prepare data  
```{r}

slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  xtabs(~is.na(coniferous), .)    # lacking for 26 rows = almost haf of the data  

slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  xtabs(~is.na(catchment_area), .)   # lacking for 37 stations (ca 1/3)

df <- slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  filter(slope_no3_vs_dep <= 0.04)  %>%   # exclude one Lativa station with very high value
  select(station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, 
         pre, tmp)
df <- df[complete.cases(df),]

cat("\n\nNumber of rows:", nrow(df), "\n")
# colnames(slope)

# plot(select(df, -station_id))
  
               
```

#### Including latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + 
                     pre + tmp,
          data = df)
dd1 <- dredge(fm)  
# dd1 <- readRDS("Data/140_Nordics_dd1.rds")

subset(dd1, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd1, 1)[[1]])$r.sq

```

#### Excluding latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + 
                     pre + tmp,
          data = df)
dd2 <- dredge(fm)

subset(dd2, delta < 2)

cat("\n\nR2 model 1 (intercept only....): \n")
summary(get.models(dd2, 1)[[1]])$r.sq

cat("\n\nR2 model 2: \n")
summary(get.models(dd2, 2)[[1]])$r.sq

```


#### Best models incl/excl latitude and longitude  
```{r, fig.width=8, fig.height=2.5}

par(mfrow = c(1,4), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(1,4), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 2)[[1]])
mtext("SECOND best model excl latitude and longitude", outer = TRUE)

```




### Europe except Nordics , including land cover   

#### Plot by region        
```{r, fig.width=6, fig.height=3.5}

slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  filter(!is.na(coniferous)) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_dep, color = region)) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2)

```


#### Prepare data  
```{r}

slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  filter(!is.na(coniferous)) %>%
  xtabs(~is.na(catchment_area), .)   # lacking for 37 stations (ca 1/3)

df <- slope %>%
  filter(!country %in% c("Norway", "Sweden", "Finland", "Latvia", "Canada", "United States")) %>%
  filter(slope_no3_vs_dep <= 0.04)  %>%   # exclude one Lativa station with very high value
  select(station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, 
         pre, tmp, catchment_area,
         coniferous, deciduous, heathlands, lake, mixed_forest, wetland)
df <- df[complete.cases(df),]

cat("\n\nNumber of rows:", nrow(df), "\n")
# colnames(slope)

# plot(select(df, -station_id))
  
               
```




#### Including latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + 
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
# dd1 <- dredge(fm)  
# saveRDS(dd1, "Data/140_Nordics_dd1.rds")    # save it as it takes a couple of minutes
dd1 <- readRDS("Data/140_Nordics_dd1.rds")

subset(dd1, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd1, 1)[[1]])$r.sq

```

#### Excluding latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + 
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
dd2 <- dredge(fm)

subset(dd2, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd2, 1)[[1]])$r.sq

```

#### Best models incl/excl latitude and longitude  
```{r, fig.width=8, fig.height=2.5}

par(mfrow = c(1,4), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(1,4), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 1)[[1]])
mtext("Best model excl latitude and longitude", outer = TRUE)

```



## 9. Analysis for all regions  

#### Prepare data  
```{r}

df <- slope %>%
  mutate(main_region = case_when(
    country %in% c("Canada", "United States") ~ "North America",
    country %in% c("Norway", "Sweden", "Finland", "Latvia") ~ "Nordic/Baltic",
    TRUE ~ "Continent / British Isles")
  ) %>%
  filter(slope_no3_vs_dep <= 0.04)  %>%   # exclude two stations with very high slope value
  select(main_region, station_id, slope_no3_vs_dep, slope_dep_vs_time, latitude, longitude, altitude, 
         pre, tmp, catchment_area,
         coniferous, deciduous, heathlands, lake, mixed_forest, wetland)
df <- df[complete.cases(df),]

cat("\n\nNumber of rows:", nrow(df), "\n")
# colnames(slope)

# plot(select(df, -station_id))
  
               
```




#### Including latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + latitude + longitude + altitude + main_region +
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
# dd1 <- dredge(fm)  
# saveRDS(dd1, "Data/140_All_dd1.rds")    # save it as it takes a couple of minutes
dd1 <- readRDS("Data/140_All_dd1.rds")

subset(dd1, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd1, 1)[[1]])$r.sq

```

#### Excluding latitude, longitude  
```{r, results  = "hold"}

options(na.action = "na.fail")
fm <- lm(
  slope_no3_vs_dep ~ slope_dep_vs_time + 
                     pre + tmp + catchment_area +
                     coniferous + deciduous + heathlands + lake + mixed_forest + 
                     wetland,
          data = df)
dd2 <- dredge(fm)

subset(dd2, delta < 2)

cat("\n\nR2: \n")
summary(get.models(dd2, 1)[[1]])$r.sq

```

#### Best models incl/excl latitude and longitude  
```{r, fig.width=8, fig.height=4}

par(mfrow = c(2,4), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg::visreg(get.models(dd1, 1)[[1]])
mtext("Best model incl latitude and longitude", outer = TRUE)

par(mfrow = c(2,4), mar = c(4,5,2,1))
visreg::visreg(get.models(dd2, 1)[[1]])
mtext("Best model excl latitude and longitude", outer = TRUE)

```





```{r}


```






