---
title: "159parm_get_landcover"
author: "DHJ"
date: "26 4 2021"
output: 
  html_document:
    keep_md: true
    toc: true  
    toc_float: true
---


## 1. Packages  
```{r, results='hide', warning=FALSE, message=FALSE}

library(readxl)
library(dplyr)

knitr::opts_chunk$set(results = 'hold') # collect the results from a chunk  
knitr::opts_chunk$set(out.width = "120")  


```


### Data source  
```{r}

localfiles <- TRUE  # if FALSE: reading from K  

```


## 2. Metadata - full station list  
```{r}

# Put in newer metadata (variable Continent has values, includes variable Group)
if (localfiles){
  fn <- "Data_landcover/all_icpw_sites_may_2020.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/all_icpw_sites_may_2020.xlsx"
}

df_meta_without_landcover <- read_excel(fn)

```


## 3. Land cover  

### Get land cover and catchment_area    
* Main land cover file has lines for UK data, but they are mostly empty  
* Read UK data separately and add to data  
```{r}

#
# Define names of land cover variables
# 
vars_landcover <- c("catchment_area",
                    "urban", "cultivated", "total_forest", "deciduous", "coniferous", 
                    "mixed_forest", "total_shrub_herbaceous", "grasslands", "heathlands", 
                    "transitional_woodland_shrub", "bare_rock", "sparsely_vegetated", 
                    "glacier", "wetland", "lake", "water_ex_lake", "other")

#
# MAIN LAND COVER FILE  
#

if (localfiles){
  fn <- "Data_landcover/ICPW_All_Stations_2020_land cover_2020_05_04.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_land cover_05_04.xlsx"
}

df_landcover1 <- read_excel(fn)
cat(nrow(df_landcover1), "lines of data read into df_landcover1 \n")
# names(df_landcover1); cat("\n")

```


### Extra UK data

#### Read file  
```{r}
#
# EXTRA UK DATA  
#
# UK data 'df_landcover1_new'

if (localfiles){
  fn <- "Data_landcover/icpw_land_cover_United Kingdom_29-01-2021.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Excel documents - received/icpw_land_cover_United Kingdom_29-01-2021.xlsx"
}

df_landcover1_new1 <- read_excel(fn, 
                               sheet = "uk", 
                               range = "A1:U23",    # NOTE: specific range! 
                               col_types = c("numeric", rep("text", 2), rep("numeric", 18))) %>%
  as.data.frame()
cat(nrow(df_landcover1_new1), "lines of UK data read into df_landcover1_new1 \n")

```


#### Remove lines from main data   
```{r}
#
# REMOVE UK LINES FROM MAIN DATA
#

# Existing data in 'df_landcover1'  
stations <- df_landcover1_new1$station_id
sel_to_remove <- df_landcover1$station_id %in% stations

# Check existing data - just have altitide, a few (erronous) catchment area, and no land cover: 
# unique(df_landcover1$station_id[sel_to_remove])
# df_landcover1[sel_to_remove,] %>% View()
# df_landcover1[sel_to_remove,] %>% select(-(station_code:longitude), -(altitude:region)) %>% View()

# Remove existing UK data:
df_landcover1b <- df_landcover1[!sel_to_remove,]
cat(sum(sel_to_remove), "lines of data removed from df_landcover1 \n")
cat(nrow(df_landcover1b), "lines of data in df_landcover1b \n")

```


#### Make data ready to be added     
```{r}
#
# MAKE UK DATA READY FOR BEING ADDED  
#

# All empty cells are replaced with zeros
df_landcover1_new1[is.na(df_landcover1_new1)] <- 0

# Get new column names for UK data 
# str(df_landcover1_new1)
new_names <- names(df_landcover1_new1) %>% 
  tolower() %>% 
  gsub(" (%)", "", ., fixed = TRUE) %>%
  gsub(" ", "_", ., fixed = TRUE)
# Check (compare):
# names(df_landcover)
# new_names

# Change all names (most are correct)
df_landcover1_new2 <- df_landcover1_new1
names(df_landcover1_new2) <- new_names

# Change the rest of the names
df_landcover1_new2 <- df_landcover1_new2 %>%
  select(-station_code, -station_name) %>%
  rename(
    catchment_area = `catchment_area_(km2)`,
    total_shrub_herbaceous = `total_shrub_and/or_herbaceous_vegetation`,
    transitional_woodland_shrub = `transitional_woodland/scrub`,
    water_ex_lake = `water_(excl._lake)`
  )

# Check column names
check <- !(names(df_landcover1_new2) %in% c("station_id", vars_landcover))
if (sum(check) > 0){
  names(df_landcover1_new2)[check] %>% print()
  stop("Some names don't fit")
} 


```


#### Add data      
```{r}

#
# ADD UK DATA TO MAIN DATA (without metadata)
#

# Combine 
df_landcover2 <- bind_rows(
  df_landcover1b[c("station_id", vars_landcover)],
  df_landcover1_new2[c("station_id", vars_landcover)],
)
cat(nrow(df_landcover2), "lines of data in df_landcover2 after adding UK data \n")

```

### Extra Canada data

#### Read file  
```{r}

if (localfiles){
  fn <- "Data_landcover/20211026 Canada combined.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Excel documents - received/Canada/20211026 Canada combined.xlsx"
}

df_landcover2_new1 <- read_excel(fn, 
                               sheet = "canada")

# Set this to numeric:
df_landcover2_new1$`Total shrub and/or herbaceous vegetation (%)` <- df_landcover2_new1$`Total shrub and/or herbaceous vegetation (%)` %>% as.numeric()

# Number of rows (should be the same)
cat(nrow(df_landcover2_new1), " lines of data read into df_landcover2_new1 \n")

cat("Number of rows in df_landcover2: ", 
        df_landcover2 %>% 
               filter(station_id %in% df_landcover2_new1$station_id) %>%
          nrow(),
    "\n")

```


#### Remove lines from main data   
```{r}

# Existing data in 'df_landcover2'  
stations <- df_landcover2_new1$station_id
sel_to_remove <- df_landcover2$station_id %in% stations

# Check existing data - just have altitide, a few (erronous) catchment area, and no land cover: 
# unique(df_landcover2$station_id[sel_to_remove])
# df_landcover2[sel_to_remove,] %>% View()
# df_landcover2[sel_to_remove,] %>% select(-(station_code:longitude), -(altitude:region)) %>% View()

# Remove existing data:
df_landcover2b <- df_landcover2[!sel_to_remove,]
cat(sum(sel_to_remove), "lines of data removed from df_landcover2 \n")
cat(nrow(df_landcover2b), "lines of data in df_landcover2b \n")

```


#### Make data ready to be added     
```{r}

# All empty cells are replaced with zeros
df_landcover2_new1[is.na(df_landcover2_new1)] <- 0

# Get new column names for the new data 
# str(df_landcover2_new1)
new_names <- names(df_landcover2_new1) %>% 
  tolower() %>% 
  gsub(" (%)", "", ., fixed = TRUE) %>%
  gsub(" ", "_", ., fixed = TRUE)
# Check (compare):
# names(df_landcover)
# new_names

# Change all names (most are correct)
df_landcover2_new2 <- df_landcover2_new1
names(df_landcover2_new2) <- new_names

# Change the rest of the names
df_landcover2_new2 <- df_landcover2_new2 %>%
  select(-station_code, -station_name) %>%
  rename(
    catchment_area = `catchment_area_(km2)`,
    total_shrub_herbaceous = `total_shrub_and/or_herbaceous_vegetation`,
    transitional_woodland_shrub = `transitional_woodland/scrub`,
    water_ex_lake = `water_(excl._lake)`
  )

# Special for Canada data:
df_landcover2_new2 <- df_landcover2_new2 %>%
  select(-total)

# Check column names
check <- !(names(df_landcover2_new2) %in% c("station_id", vars_landcover))
if (sum(check) > 0){
  names(df_landcover2_new2)[check] %>% print()
  stop("Some names don't fit")
} 


```


#### Add data      
```{r}

#
# ADD DATA TO MAIN DATA (without metadata)
#

# Combine 
df_landcover2 <- bind_rows(
  df_landcover2b[c("station_id", vars_landcover)],
  df_landcover2_new2[c("station_id", vars_landcover)],
)
cat(nrow(df_landcover2), "lines of data in df_landcover2 after adding Canada data \n")

```



## 4a. Add land cover to metadata  
```{r}

n1 <- nrow(df_meta_without_landcover)

df_meta1 <- df_meta_without_landcover %>% 
  left_join(df_landcover2, by = "station_id")

n2 <- nrow(df_meta1)

# Check that join is ok
if (n2 != n1)
  stop("Number of rows changed; check whether there are only unique values of 'station_id'")

# Order columns
df_meta1 <- df_meta1[c(names(df_meta_without_landcover), vars_landcover)]

# names(df_landcover1); cat("\n")

cat(nrow(df_meta1), "lines of data in df_meta1 after adding new metadata \n")

```
### Add altitude for Ireland  

#### Read file  
```{r}


if (localfiles){
  fn <- "Data_landcover/Ireland_icpsites_may2020.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Ireland_icpsites_may2020.xlsx"
}

df_landcover_ie <- read_excel(fn)

cat("Number of rows: ", 
        nrow(df_landcover_ie),
    "\n\n")

cat("Number of rows in df_meta1: ", 
        df_meta1 %>% 
               filter(station_id %in% df_landcover_ie$station_id) %>%
          nrow(),
    "\n\n")

```

#### Add new altitude
```{r}

df_meta2 <- df_meta1 %>%
  left_join(
    df_landcover_ie %>% 
      select(station_id, altitude) %>%
      rename(altitude_ie = altitude)
  ) %>%
  mutate(
    altitude = case_when(
      is.na(altitude_ie) ~ altitude,
      !is.na(altitude_ie) ~ altitude_ie)
    )

xtabs(~is.na(altitude), df_meta1)
xtabs(~is.na(altitude), df_meta2)

```


### Save     
```{r}

saveRDS(df_meta2, "Data/159_df_meta1.rds")
writexl::write_xlsx(df_meta2, "Data/159_df_meta1.xlsx")

# df_landcover2 <- readRDS("Data/159_df_landcover2.rds")

cat("\n------------------------------------------\n")
cat("Coverage per country\n")
xtabs(~country + is.na(total_forest), df_meta1)

```

## 4b. Check/fix land cover sums   

- The first part (forest) goes from df_meta2 -> df_meta2b -> df_meta2c

- The second part (shrub / grass) goes from df_meta2c -> df_meta2d -> df_meta2e


### Total forest vs. coniferous + deciduous + mixed forest 

#### Tables     
- 29 records have neither   
- 9 records have total forest, but not each type   
```{r}

xtabs(~is.na(total_forest) + is.na(coniferous), df_meta2)  
xtabs(~is.na(total_forest) + is.na(deciduous), df_meta2)  
xtabs(~is.na(total_forest) + is.na(mixed_forest), df_meta2)  

```
#### But many have total_forest = 0
```{r, results='hold'}

cat("For records with total_forest not NA: \n")
xtabs(~ (total_forest == 0) + is.na(coniferous), df_meta2 %>% filter(!is.na(total_forest)))  

cat("\n\nFor records with coniferous not NA: \n")
xtabs(~ (coniferous == 0) + (total_forest == 0) + is.na(total_forest), df_meta2 %>% filter(!is.na(coniferous)))  

```
#### Table for checking  
```{r}

df_check1 <- df_meta2 %>%
  filter(!is.na(total_forest) | !is.na(coniferous)) %>%
  select(total_forest, coniferous, deciduous, mixed_forest) %>%
  mutate(across(.fns = round, digits = 2)) %>%
  mutate(
    check_sum = coniferous + deciduous + mixed_forest,
    check_diff = round(total_forest - check_sum, 1)) %>%
  arrange(check_diff)


# View(df_check1)

cat("Records with total_forest OR coniferous: \n")
nrow(df_check1)


```

#### Check records with total_forest OR coniferous   
- 328 are OK (both total_forest and coniferous, and sum is OK)  
- 65 more are also OK (zero for both total_forest and coniferous)  
- 4 have check_sum = zero but total > 0 (not OK, sub-types be set to NA)     
- 4 have check_sum = NA but total > 0 (this is OK)         
- 121 have data for check_sum but total = 0  
- 4 have check_sum = NA and total = 0 (can we trust that this is actually no forest?)    
```{r}


df_check2 <- df_check1 %>%
  mutate(
    total_forest_na = is.na(total_forest),
    total_forest_zero = total_forest == 0,
    check_sum_na = is.na(check_sum),
    check_sum_zero = check_sum == 0,
    sum_OK = abs(check_diff) < 0.01) 

View(df_check2)

df_check2 %>%
  count(total_forest_na, total_forest_zero, check_sum_na, check_sum_zero, sum_OK)

```

#### Fix total forest    
- New result  
    - 449 are still OK (total and check_sum are both > 0, total = check_sum)    
    - 65 are hopefully OK (total and check_sum are both = 0)     
    - 8 have total > 0, coniferous + deciduous + mixed_forest are set to NA  
    - 5 have total = 0, coniferous + deciduous + mixed_forest are set to NA  
```{r}

df_meta2b <- df_meta2 %>%
  mutate(check_sum = coniferous + deciduous + mixed_forest) %>%
  mutate(
    total_forest_new = case_when(
      !is.na(check_sum) & check_sum > 0 ~ coniferous + deciduous + mixed_forest,
      TRUE ~ total_forest)
    )

# Those with total > 0 but check_sum = 0, we set coniferous, deciduous. mixed to NA
sel <- !is.na(df_meta2b$check_sum) & df_meta2b$check_sum == 0 & df_meta2b$total_forest_new > 0  
sum(sel)
# View(df_meta2b[sel,])
df_meta2b$coniferous[sel] <- NA
df_meta2b$deciduous[sel] <- NA
df_meta2b$mixed_forest[sel] <- NA

df_check3 <- df_meta2b %>%
  filter(!is.na(total_forest_new) | !is.na(coniferous)) %>%
  select(total_forest, total_forest_new, coniferous, deciduous, mixed_forest) %>%
  mutate(across(.fns = round, digits = 2)) %>%
  mutate(
    check_sum = coniferous + deciduous + mixed_forest,
    check_diff = round(total_forest_new - check_sum, 1)) %>%
  arrange(check_diff)

df_check4 <- df_check3 %>%
  mutate(
    total_forest_na = is.na(total_forest_new),
    total_forest_zero = total_forest_new == 0,
    check_sum_na = is.na(check_sum),
    check_sum_zero = check_sum == 0,
    sum_OK = abs(check_diff) < 0.01) 

# View(df_check4)

df_check4 %>%
  count(total_forest_na, total_forest_zero, check_sum_na, check_sum_zero, sum_OK)

```

#### Corrected data set  
```{r}

df_meta2c <- df_meta2b %>%
  select(-total_forest) %>%
  rename(total_forest = total_forest_new)


```




### Total shrub/herbacous vs. grasslands + heathlands +  transitional_woodland_shrub

#### Tables     
- 29 records have neither   
- 28 records have total_shrub_herbaceous, but not each type   
```{r}

xtabs(~is.na(total_shrub_herbaceous) + is.na(grasslands), df_meta2c)  
xtabs(~is.na(total_shrub_herbaceous) + is.na(heathlands), df_meta2c)  
xtabs(~is.na(total_shrub_herbaceous) + is.na(transitional_woodland_shrub), df_meta2c)  

```


#### But many have total_shrub_herbaceous = 0

```{r, results='hold'}

cat("For records with total_shrub_herbaceous not NA: \n")
xtabs(~ (total_shrub_herbaceous == 0) + is.na(grasslands), df_meta2c %>% filter(!is.na(total_shrub_herbaceous)))  

cat("\n\nFor records with grasslands not NA: \n")
xtabs(~ (grasslands == 0) + (total_shrub_herbaceous == 0) + is.na(total_shrub_herbaceous), df_meta2c %>% filter(!is.na(grasslands)))  

```


#### Table for checking  

```{r}

df_check1 <- df_meta2c %>%
  filter(!is.na(total_shrub_herbaceous) | !is.na(grasslands)) %>%
  select(total_shrub_herbaceous, grasslands, heathlands, transitional_woodland_shrub) %>%
  mutate(across(.fns = round, digits = 2)) %>%
  mutate(
    check_sum = grasslands + heathlands + transitional_woodland_shrub,
    check_diff = round(total_shrub_herbaceous - check_sum, 1)) %>%
  arrange(check_diff)


# View(df_check1)

cat("Records with total_shrub_herbaceous OR grasslands: \n")
nrow(df_check1)


```

#### Check records with total_shrub_herbaceous OR grasslands   
- 207 are OK (both total_shrub_herbaceous > 0 and check_sum > 0, and sum is OK)  
- 144 more are also OK (both total_shrub_herbaceous = 0 and check_sum = 0  
- 56 have check_sum = zero but total > 0  (not OK, sub-types be set to NA)     
- 26 have check_sum = NA but total > 0 (this is OK)         
- 92 have check_sum > 0 but total = 0  
- 2 have check_sum = NA and total = 0 (can we trust that this is actually no total_shrub_herbaceous?)    
```{r}


df_check2 <- df_check1 %>%
  mutate(
    total_shrub_herbaceous_na = is.na(total_shrub_herbaceous),
    total_shrub_herbaceous_zero = total_shrub_herbaceous == 0,
    check_sum_na = is.na(check_sum),
    check_sum_zero = check_sum == 0,
    sum_OK = abs(check_diff) < 0.01) 

View(df_check2)

df_check2 %>%
  count(total_shrub_herbaceous_na, total_shrub_herbaceous_zero, check_sum_na, check_sum_zero, sum_OK)

```

#### Fix total_shrub_herbaceous    
- New result  
    - 299 are still OK (total and check_sum are both > 0, total = check_sum)    
    - 144 are hopefully OK (total and check_sum are both = 0)     
    - 82 have total > 0, grasslands + heathlands + transitional_woodland_shrub are set to NA  
    - 2 have total = 0, grasslands + heathlands + transitional_woodland_shrub are set to NA  
```{r}

df_meta2d <- df_meta2c %>%
  mutate(check_sum = grasslands + heathlands + transitional_woodland_shrub) %>%
  mutate(
    total_shrub_herbaceous_new = case_when(
      !is.na(check_sum) & check_sum > 0 ~ grasslands + heathlands + transitional_woodland_shrub,
      TRUE ~ total_shrub_herbaceous)
    )

# Those with total > 0 but check_sum = 0, we set grasslands, heathlands. mixed to NA
sel <- !is.na(df_meta2d$check_sum) & df_meta2d$check_sum == 0 & df_meta2d$total_shrub_herbaceous_new > 0  
sum(sel)
# View(df_meta2d[sel,])
df_meta2d$grasslands[sel] <- NA
df_meta2d$heathlands[sel] <- NA
df_meta2d$transitional_woodland_shrub[sel] <- NA

df_check3 <- df_meta2d %>%
  filter(!is.na(total_shrub_herbaceous_new) | !is.na(grasslands)) %>%
  select(total_shrub_herbaceous, total_shrub_herbaceous_new, grasslands, heathlands, transitional_woodland_shrub) %>%
  mutate(across(.fns = round, digits = 2)) %>%
  mutate(
    check_sum = grasslands + heathlands + transitional_woodland_shrub,
    check_diff = round(total_shrub_herbaceous_new - check_sum, 1)) %>%
  arrange(check_diff)

df_check4 <- df_check3 %>%
  mutate(
    total_shrub_herbaceous_na = is.na(total_shrub_herbaceous_new),
    total_shrub_herbaceous_zero = total_shrub_herbaceous_new == 0,
    check_sum_na = is.na(check_sum),
    check_sum_zero = check_sum == 0,
    sum_OK = abs(check_diff) < 0.01) 

# View(df_check4)

df_check4 %>%
  count(total_shrub_herbaceous_na, total_shrub_herbaceous_zero, check_sum_na, check_sum_zero, sum_OK)

```

#### Corrected data set  
```{r}

df_meta2e <- df_meta2d %>%
  select(-total_shrub_herbaceous) %>%
  rename(total_shrub_herbaceous = total_shrub_herbaceous_new)


```




## 5. Combine land cover types 'df_meta2'      
* bare_sparse = bare_rock + sparsely_vegetated + glacier   
* Select: coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse   
* Also: removes all metadata (keeps only station_id, catchment_area, and land cover columns)  
```{r}

# df_landcover2 %>% select(-(station_code:longitude), -(altitude:region)) %>% names() %>% dput()

# Select columns
sel_columns <- c("station_id",
                 "station_code", "station_name", "latitude", "longitude", 
                 "altitude", "continent", "country", "region", "group",
                 "catchment_area", 
                 "urban", "cultivated", "total_forest", 
                 "deciduous", "coniferous", "mixed_forest", "total_shrub_herbaceous", 
                 "grasslands", "heathlands", "transitional_woodland_shrub", "bare_rock", 
                 "sparsely_vegetated", "glacier", "wetland", "lake", "water_ex_lake", "other")

# Add new combined data
df_meta3 <- df_meta2e[sel_columns] %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier,
         decid_mixed = deciduous + mixed_forest,
         lake_water = lake + water_ex_lake)

# Remove 'old' columns
df_meta4 <- df_meta3 %>%
  select(-c(bare_rock, sparsely_vegetated, glacier, deciduous, mixed_forest, lake, water_ex_lake))

cat("df_meta3, n =", nrow(df_meta2), "\n")


```



## 6. Save  
```{r}

# Metadata including all original columns
saveRDS(df_meta3, "Data/159_df_meta3.rds")
writexl::write_xlsx(df_meta3, "Data/159_df_metadata_merged_landcover_allcols.xlsx")

# Metadata excluding columns that have been combined 
saveRDS(df_meta4, "Data/159_df_meta4.rds")
writexl::write_xlsx(df_meta4, "Data/159_df_metadata_merged_landcover.xlsx")

```


## 7. Statistics   

### Lacking variables      
* 76 stations lack only catchment_area   
* 11 stations lack catchment_area + land cover (14 variables in total)   
* 17 stations have catchment_area, lack land cover   
* 1 station lack catchment_area + land cover + altitude (lacking 15 variables in total)   
* 1 station lack catchment_area + altitude (lacking 2 variables in total)    
* 28 stations have land cover but lack grasslands, heathlands and transitional_woodland_shrub  
* 9 stations have land cover but lack only coniferous + decid_mixed   

```{r}

cat("Number of stations which lack x variables")
tab1 <- apply(is.na(df_meta4), 1, sum)
tab2 <- table(tab1)
tab2

cat("\n\n")

for (i in as.numeric(names(tab2))){
  cat("Variables lacking for stations lacking", i, "variables: \n")
  tab_i <- apply(is.na(df_meta4[tab1 == i,]), 2, sum)
  print(
    tab_i[tab_i > 0])
  cat("\n")
}

```


### Variables lacking, by country    
- USA lacks catchment_area (79 stations) and partially land cover (16 stations)   
- Sweden partially (28 stations) lack 'grasslands' (i.e. land use types are lumped)    
- Switzerland (9 stations) lack 'coniferous' (i.e. land use types are lumped)    
```{r}

cat("Availablility of altitude: \n")
df_meta4 %>%
  xtabs(~is.na(altitude) + country, .) %>% print()

cat("-----------------------------------------------------------------------------------------\n")
cat("Availablility of catchment_area: \n")
df_meta4 %>%
  xtabs(~is.na(catchment_area) + country, .) %>% print()

cat("-----------------------------------------------------------------------------------------\n")
cat("Availablility of catchment_area, for stations having land cover: \n")
df_meta4 %>%
  filter(!is.na(total_forest)) %>%
  xtabs(~is.na(catchment_area) + country, .)

cat("-----------------------------------------------------------------------------------------\n")
cat("Availablility of land cover, for stations having catchment_area: \n")
df_meta4 %>%
  filter(!is.na(catchment_area)) %>%
  xtabs(~is.na(total_forest) + country, .)

cat("-----------------------------------------------------------------------------------------\n")
cat("Availablility of 'grasslands', for stations having land cover: \n")
df_meta4 %>%
  filter(!is.na(total_forest)) %>%
  xtabs(~is.na(grasslands) + country, .)

cat("-----------------------------------------------------------------------------------------\n")
cat("Availablility of 'coniferous', for stations having land cover: \n")
df_meta4 %>%
  filter(!is.na(total_forest)) %>%
  xtabs(~is.na(coniferous) + country, .)


```

## Appendix: compare with other metadata    
```{r}

run_this <- FALSE
# run_this <- TRUE

### Check other land cover files
if (run_this){
  
  # Station metadata
  # THESE WILL BE ADDED TOGETHER WITH LAND COVER  
  
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
  file <- "trends_1992-2016_no3_stations.csv"
  fn <- paste0(folder, "/", file)
  df1 <- read.csv(fn, encoding = "UTF-8") %>% 
    arrange(station_id)
  nrow(df1)
  
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/all_icpw_sites_may_2020.xlsx"
  df2 <- read_excel(fn) %>% 
    arrange(station_id)
  nrow(df2)
  
  # Column names:
  n1 <- names(df1)
  n2 <- names(df2)
  cat("Missing in df1: \n")
  n2[!n2 %in% n1]
  cat("Missing in df2: \n")
  n1[!n1 %in% n2]
  
  cat("\n")
  cat("---------------------------------------------------------------\n")
  cat("Compare common variables: \n")
  cat("---------------------------------------------------------------\n")
  n_common <- n1[n1 %in% n2]
  for (col in n_common){
    # col <- "station_id"
    x1 <- sum(!df1[[col]] == df2[[col]], na.rm = TRUE)
    x2a <- sum(is.na(df1[[col]]))
    x2b <- sum(is.na(df2[[col]]))
    x3 <- sum(is.na(df1[[col]]) != is.na(df2[[col]]))
    cat(col, ": \n  n differences (rows with values):", x1, "\n")
    cat("  Rows lacking values:", x2a, "/", x2b, "\n")
    cat("  Rows lacking values different places:", x3, "\n")
  }


# View side by side
df1_forjoin <- df1[n_common]
names(df1_forjoin) <- paste0(n_common, "_1")  
df2_forjoin <- df2[n_common]
names(df2_forjoin) <- paste0(n_common, "_2")
# bind_cols(df1_forjoin, df2_forjoin) %>% View()
bind_cols(df1_forjoin$region_1, df2_forjoin$region_2) %>% View()
# fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_land cover_2020_05_04.xlsx"
  # df3 <- read_excel(fn)

}


```

