---
title: "159parm_get_landcover"
author: "DHJ"
date: "26 4 2021"
output: html_document
---


## 1. Packages  
```{r}

library(readxl)
library(dplyr)

```


### Data source  
```{r}

localfiles <- TRUE  # if FALSE: reading from K  

```


## 2. Metadata - full station list  
```{r}

# Put in newer metadata (variable Continent has values, includes variable Group)
if (localfiles){
  fn <- "Data_landcover/all_icpw_sites_may_2020.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/all_icpw_sites_may_2020.xlsx"
}

df_meta_without_landcover <- read_excel(fn)

```


## 3. Land cover  

### Get land cover and catchment_area    
* Main land cover file has lines for UK data, but they are mostly empty  
* Read UK data separately and add to data  
```{r}

#
# Define names of land cover variables
# 
vars_landcover <- c("catchment_area",
                    "urban", "cultivated", "total_forest", "deciduous", "coniferous", 
                    "mixed_forest", "total_shrub_herbaceous", "grasslands", "heathlands", 
                    "transitional_woodland_shrub", "bare_rock", "sparsely_vegetated", 
                    "glacier", "wetland", "lake", "water_ex_lake", "other")

#
# MAIN LAND COVER FILE  
#

if (localfiles){
  fn <- "Data_landcover/ICPW_All_Stations_2020_land cover_2020_05_04.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_land cover_05_04.xlsx"
}

df_landcover1 <- read_excel(fn)
cat(nrow(df_landcover1), "lines of data read into df_landcover1 \n")
# names(df_landcover1); cat("\n")

```


### Extra UK data

#### Read file  
```{r}
#
# EXTRA UK DATA  
#
# UK data 'df_landcover1_new'

if (localfiles){
  fn <- "Data_landcover/icpw_land_cover_United Kingdom_29-01-2021.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Excel documents - received/icpw_land_cover_United Kingdom_29-01-2021.xlsx"
}

df_landcover1_new1 <- read_excel(fn, 
                               sheet = "uk", 
                               range = "A1:U23",    # NOTE: specific range! 
                               col_types = c("numeric", rep("text", 2), rep("numeric", 18))) %>%
  as.data.frame()
cat(nrow(df_landcover1_new1), "lines of UK data read into df_landcover1_new1 \n")

```


#### Remove lines from main data   
```{r}
#
# REMOVE UK LINES FROM MAIN DATA
#

# Existing data in 'df_landcover1'  
stations <- df_landcover1_new1$station_id
sel_to_remove <- df_landcover1$station_id %in% stations

# Check existing data - just have altitide, a few (erronous) catchment area, and no land cover: 
# unique(df_landcover1$station_id[sel_to_remove])
# df_landcover1[sel_to_remove,] %>% View()
# df_landcover1[sel_to_remove,] %>% select(-(station_code:longitude), -(altitude:region)) %>% View()

# Remove existing UK data:
df_landcover1b <- df_landcover1[!sel_to_remove,]
cat(sum(sel_to_remove), "lines of data removed from df_landcover1 \n")
cat(nrow(df_landcover1b), "lines of data in df_landcover1b \n")

```


#### Make data ready to be added     
```{r}
#
# MAKE UK DATA READY FOR BEING ADDED  
#

# All empty cells are replaced with zeros
df_landcover1_new1[is.na(df_landcover1_new1)] <- 0

# Get new column names for UK data 
# str(df_landcover1_new1)
new_names <- names(df_landcover1_new1) %>% 
  tolower() %>% 
  gsub(" (%)", "", ., fixed = TRUE) %>%
  gsub(" ", "_", ., fixed = TRUE)
# Check (compare):
# names(df_landcover)
# new_names

# Change all names (most are correct)
df_landcover1_new2 <- df_landcover1_new1
names(df_landcover1_new2) <- new_names

# Change the rest of the names
df_landcover1_new2 <- df_landcover1_new2 %>%
  select(-station_code, -station_name) %>%
  rename(
    catchment_area = `catchment_area_(km2)`,
    total_shrub_herbaceous = `total_shrub_and/or_herbaceous_vegetation`,
    transitional_woodland_shrub = `transitional_woodland/scrub`,
    water_ex_lake = `water_(excl._lake)`
  )

# Check column names
check <- !(names(df_landcover1_new2) %in% c("station_id", vars_landcover))
if (sum(check) > 0){
  names(df_landcover1_new2)[check] %>% print()
  stop("Some names don't fit")
} 


```


#### Add data      
```{r}

#
# ADD UK DATA TO MAIN DATA (without metadata)
#

# Combine 
df_landcover2 <- bind_rows(
  df_landcover1b[c("station_id", vars_landcover)],
  df_landcover1_new2[c("station_id", vars_landcover)],
)
cat(nrow(df_landcover2), "lines of data in df_landcover2 after adding UK data \n")

```

### Extra Canada data

#### Read file  
```{r}

if (localfiles){
  fn <- "Data_landcover/20211026 Canada combined.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Excel documents - received/Canada/20211026 Canada combined.xlsx"
}

df_landcover2_new1 <- read_excel(fn, 
                               sheet = "canada")

# Set this to numeric:
df_landcover2_new1$`Total shrub and/or herbaceous vegetation (%)` <- df_landcover2_new1$`Total shrub and/or herbaceous vegetation (%)` %>% as.numeric()

# Number of rows (should be the same)
message(nrow(df_landcover2_new1), " lines of data read into df_landcover2_new1")

message("Number of rows in df_landcover2: ", 
        df_landcover2 %>% 
               filter(station_id %in% df_landcover2_new1$station_id) %>%
          nrow()
        )

```


#### Remove lines from main data   
```{r}

# Existing data in 'df_landcover2'  
stations <- df_landcover2_new1$station_id
sel_to_remove <- df_landcover2$station_id %in% stations

# Check existing data - just have altitide, a few (erronous) catchment area, and no land cover: 
# unique(df_landcover2$station_id[sel_to_remove])
# df_landcover2[sel_to_remove,] %>% View()
# df_landcover2[sel_to_remove,] %>% select(-(station_code:longitude), -(altitude:region)) %>% View()

# Remove existing data:
df_landcover2b <- df_landcover2[!sel_to_remove,]
cat(sum(sel_to_remove), "lines of data removed from df_landcover2 \n")
cat(nrow(df_landcover2b), "lines of data in df_landcover2b \n")

```


#### Make data ready to be added     
```{r}

# All empty cells are replaced with zeros
df_landcover2_new1[is.na(df_landcover2_new1)] <- 0

# Get new column names for the new data 
# str(df_landcover2_new1)
new_names <- names(df_landcover2_new1) %>% 
  tolower() %>% 
  gsub(" (%)", "", ., fixed = TRUE) %>%
  gsub(" ", "_", ., fixed = TRUE)
# Check (compare):
# names(df_landcover)
# new_names

# Change all names (most are correct)
df_landcover2_new2 <- df_landcover2_new1
names(df_landcover2_new2) <- new_names

# Change the rest of the names
df_landcover2_new2 <- df_landcover2_new2 %>%
  select(-station_code, -station_name) %>%
  rename(
    catchment_area = `catchment_area_(km2)`,
    total_shrub_herbaceous = `total_shrub_and/or_herbaceous_vegetation`,
    transitional_woodland_shrub = `transitional_woodland/scrub`,
    water_ex_lake = `water_(excl._lake)`
  )

# Special for Canada data:
df_landcover2_new2 <- df_landcover2_new2 %>%
  select(-total)

# Check column names
check <- !(names(df_landcover2_new2) %in% c("station_id", vars_landcover))
if (sum(check) > 0){
  names(df_landcover2_new2)[check] %>% print()
  stop("Some names don't fit")
} 


```


#### Add data      
```{r}

#
# ADD DATA TO MAIN DATA (without metadata)
#

# Combine 
df_landcover2 <- bind_rows(
  df_landcover2b[c("station_id", vars_landcover)],
  df_landcover2_new2[c("station_id", vars_landcover)],
)
cat(nrow(df_landcover2), "lines of data in df_landcover2 after adding Canada data \n")

```



## 4. Add land cover to metadata  
```{r}

n1 <- nrow(df_meta_without_landcover)

df_meta1 <- df_meta_without_landcover %>% 
  left_join(df_landcover2, by = "station_id")

n2 <- nrow(df_meta1)

# Check that join is ok
if (n2 != n1)
  stop("Number of rows changed; check whether there are only unique values of 'station_id'")

# Order columns
df_meta1 <- df_meta1[c(names(df_meta_without_landcover), vars_landcover)]

# names(df_landcover1); cat("\n")

cat(nrow(df_meta1), "lines of data in df_meta1 after adding new metadata \n")

```
### Add altitude for Ireland  

#### Read file  
```{r}


if (localfiles){
  fn <- "Data_landcover/Ireland_icpsites_may2020.xlsx"
} else {
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/Ireland_icpsites_may2020.xlsx"
}

df_landcover_ie <- read_excel(fn)

message("Number of rows: ", 
        nrow(df_landcover_ie))

message("Number of rows in df_meta1: ", 
        df_meta1 %>% 
               filter(station_id %in% df_landcover_ie$station_id) %>%
          nrow()
        )
```

#### Add new altitude
```{r}

df_meta2 <- df_meta1 %>%
  left_join(
    df_landcover_ie %>% 
      select(station_id, altitude) %>%
      rename(altitude_ie = altitude)
  ) %>%
  mutate(
    altitude = case_when(
      is.na(altitude_ie) ~ altitude,
      !is.na(altitude_ie) ~ altitude_ie)
    )

xtabs(~is.na(altitude), df_meta1)
xtabs(~is.na(altitude), df_meta2)

```


### Save     
```{r}

#
# SAVE
#

saveRDS(df_meta1, "Data/159_df_meta1.rds")
writexl::write_xlsx(df_meta1, "Data/159_df_meta1.xlsx")

# df_landcover2 <- readRDS("Data/159_df_landcover2.rds")

cat("\n------------------------------------------\n")
cat("Coverage per country\n")
xtabs(~country + is.na(total_forest), df_meta1)

```


## 5. Combine land cover types 'df_meta2'      
* bare_sparse = bare_rock + sparsely_vegetated + glacier   
* Select: coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse   
* Also: removes all metadata (keeps only station_id, catchment_area, and land cover columns)  
```{r}

# df_landcover2 %>% select(-(station_code:longitude), -(altitude:region)) %>% names() %>% dput()

# Select columns
sel_columns <- c("station_id", "catchment_area", "urban", "cultivated", "total_forest", 
                   "deciduous", "coniferous", "mixed_forest", "total_shrub_herbaceous", 
                   "grasslands", "heathlands", "transitional_woodland_shrub", "bare_rock", 
                   "sparsely_vegetated", "glacier", "wetland", "lake", "water_ex_lake", "other")
  
df_meta2 <- df_meta1[sel_columns] %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier,
         decid_mixed = deciduous + mixed_forest,
         lake_water = lake + water_ex_lake) %>%
  select(-c(bare_rock, sparsely_vegetated, glacier, deciduous, mixed_forest, lake, water_ex_lake))

cat("df_meta2, n =", nrow(df_meta2), "\n")


```



## 6. Save  
```{r}

saveRDS(df_meta2, "Data/159_df_meta2.rds")
writexl::write_xlsx(df_meta2, "Data/159_df_metadata_merged_landcover.xlsx")

```


## Appendix: compare with other metadata    
```{r}

run_this <- FALSE
# run_this <- TRUE

### Check other land cover files
if (run_this){
  
  # Station metadata
  # THESE WILL BE ADDED TOGETHER WITH LAND COVER  
  
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
  file <- "trends_1992-2016_no3_stations.csv"
  fn <- paste0(folder, "/", file)
  df1 <- read.csv(fn, encoding = "UTF-8") %>% 
    arrange(station_id)
  nrow(df1)
  
  fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/all_icpw_sites_may_2020.xlsx"
  df2 <- read_excel(fn) %>% 
    arrange(station_id)
  nrow(df2)
  
  # Column names:
  n1 <- names(df1)
  n2 <- names(df2)
  cat("Missing in df1: \n")
  n2[!n2 %in% n1]
  cat("Missing in df2: \n")
  n1[!n1 %in% n2]
  
  cat("\n")
  cat("---------------------------------------------------------------\n")
  cat("Compare common variables: \n")
  cat("---------------------------------------------------------------\n")
  n_common <- n1[n1 %in% n2]
  for (col in n_common){
    # col <- "station_id"
    x1 <- sum(!df1[[col]] == df2[[col]], na.rm = TRUE)
    x2a <- sum(is.na(df1[[col]]))
    x2b <- sum(is.na(df2[[col]]))
    x3 <- sum(is.na(df1[[col]]) != is.na(df2[[col]]))
    cat(col, ": \n  n differences (rows with values):", x1, "\n")
    cat("  Rows lacking values:", x2a, "/", x2b, "\n")
    cat("  Rows lacking values different places:", x3, "\n")
  }


# View side by side
df1_forjoin <- df1[n_common]
names(df1_forjoin) <- paste0(n_common, "_1")  
df2_forjoin <- df2[n_common]
names(df2_forjoin) <- paste0(n_common, "_2")
# bind_cols(df1_forjoin, df2_forjoin) %>% View()
bind_cols(df1_forjoin$region_1, df2_forjoin$region_2) %>% View()
# fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_land cover_2020_05_04.xlsx"
  # df3 <- read_excel(fn)

}


```

