---
title: "160_Time_series_results_James"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true

---

Analysis of James' trend results    
* Taken from `https://github.com/JamesSample/icpw2/tree/master/thematic_report_2020/results`   
* Sen slope of NO3, TOTN, TOC/TON etc.  
* Note that these contains no data fra United States!  


## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of tehse packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)

# Too many packages, not all are used
# library(forcats)
# library(mgcv)
# library(nlme)
library(mapview)
library(visreg)     # visreg
library(rkt)        # Theil -Sen Regression

library(MuMIn)      

# Trees and forests
# install.packages("party")
# install.packages("modeltools")
# install.packages("coin")
# install.packages("multcomp")
# install.packages("TH.data")
library(party)
# install.packages("evtree")
library(evtree)
# install.packages("randomForest")
library(randomForest)

library(maps)
my_map <- map_data("world")

library(effects)    # handles lme models  
library(readxl)

knitr::opts_chunk$set(results = 'hold')
knitr::opts_chunk$set(warning = FALSE)

```


## 2. Data
```{r}

dat_annual_sel <- readRDS("Data/120_dat_annual_sel.rds")
# ts_model_list <- readRDS("Data/120_ts_model_list_linear.rds")
# ts_model_list_wout_TOC <- readRDS("Data/120_ts_model_list_wout_TOC.rds")

df_stations <- readRDS("Data/100_Stations.rds")
df_deposition <- readRDS("Data/100_Deposition.rds")
df_climate <- readRDS("Data/100_Climate.rds")

```

### Station metadata
```{r}

df_station <- read_excel(
  "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_2020_05_04.xlsx") %>%
  mutate(station_id = as.character(station_id))

```

### James results  
```{r}

#
# Regression results
#
fn <- "https://raw.githubusercontent.com/JamesSample/icpw2/master/thematic_report_2020/results/trends_1992-2016_toc_totn_no3_relax_italy/trends_1992-2016_toc_totn_no3_relax_italy_results.csv"
reg_no3 <- read.csv(fn, encoding = "UTF-8")
nrow(reg_no3)

# Station metadata
fn <- "https://raw.githubusercontent.com/JamesSample/icpw2/master/thematic_report_2020/results/trends_1992-2016_toc_totn_no3_relax_italy/trends_1992-2016_toc_totn_no3_relax_italy_stations.csv"
reg_no3_st <- read.csv(fn, encoding = "UTF-8")

# xtabs(~variable, reg_no3)  

#
# Medians
#
fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016/medians_2012-2016_no3.csv"
medians_no3 <- read.csv(fn, encoding = "UTF-8")
nrow(medians_no3)

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016/medians_2012-2016_toc_totn_no3_nh4.csv"
medians_toc <- read.csv(fn, encoding = "UTF-8")
nrow(medians_toc)


```
### Start 'dat'  
With slope regression data  
* Make one line per station  
```{r}

df1 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, sen_slp) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "sen_slp")
df2 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, mk_p_val) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mk_p_val")

names(df1)[2:3] <- c("slope_no3_vs_time", "slope_tocton_vs_time")
names(df2)[2:3] <- c("p_no3_vs_time", "p_tocton_vs_time")

dat <- full_join(df1, df2, by = "station_id")

```

### Deposition Theil-Senslope   
```{r}

### Get deposition  
df_deposition <- readRDS("Data/100_Deposition.rds") %>%
  as.data.frame()

# TEST Senslope
# df_test <- df_deposition %>% filter(station_id == 38115)
# result <- rkt(df_test$year, df_test$TOTN_dep)
# data.frame(station_id = 38115, TOTN_dep_P = result$sl, TOTN_dep_slope = result$B)

# Function for Sen slope
get_senslope <- function(data){
  result <- rkt(data$year, data$TOTN_dep)
  data.frame(station_id = data$station_id[1], slope_dep_vs_time = result$B, p_dep_vs_time = result$sl)
}

df_deposition_slope <- df_deposition %>%
  split(.$station_id) %>% # str()
  purrr::map_dfr(get_senslope)  

```


### Add deposition slope to data  
```{r}

nrow(dat)
dat <- left_join(dat, 
                 df_deposition_slope,
                 by = "station_id")

nrow(dat)


```
### Add medians and station metadata   
```{r}

nrow(dat)
dat <- dat %>%
  left_join(medians_no3, by = "station_id") %>%
  # left_join(medians_toc, by = "station_id") %>%   # not needed, overlap in variables  
  left_join(reg_no3_st, by = "station_id")
nrow(dat)

# Simplify names by removing units
# names(dat)
names(dat) <- sub(".N_µg.l.N", "", names(dat))
names(dat) <- sub("_mg.C.l", "", names(dat))
names(dat) <- sub("_µg.l.P", "", names(dat))
names(dat)


```

### Add climate and deposition means 
```{r}

df_climate <- readRDS("Data/100_Climate.rds")

df_climate_mean <- df_climate %>%
  group_by(station_id, variable) %>% 
  summarise(mean = mean(value)) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mean")

df_deposition_mean <-
  df_deposition %>% 
  group_by(station_id) %>% 
  summarise(Mean_dep = mean(TOTN_dep))

# Add
dat <- dat %>%
  left_join(df_climate_mean, by = "station_id") %>%
  left_join(df_deposition_mean, by = "station_id")
nrow(dat)

```

### Add land cover    
```{r}

fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Land cover/ICPW_All_Stations_2020_2020_05_04.xlsx"

df_station <- read_excel(fn)

nrow(dat)
dat <- left_join(dat, 
                 df_station %>% select(-(station_code:longitude), -(altitude:region)), 
                 by = "station_id"
)

nrow(dat)

```



## 3. Plot estimates  

Note: **No USA**

```{r, fig.width=9, fig.height=7}

ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time)) + 
  geom_point(data = dat %>% filter(p_no3_vs_time < 0.05), size = rel(2)) +
  geom_point(aes(color = country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) 
  
ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time,
                color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")

dat %>%
  filter(!country %in% c("Latvia","Ireland","Italy","Netherlands")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_time,
             color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")


```


## 4. Significant / non-significant NO3 decline  

### Data for analysis  
```{r}

# Data for analyses
df_analysis <- dat %>%
  mutate(
    no3_decline = case_when(
      slope_no3_vs_time < 0 & p_no3_vs_time <= 0.05 ~ 1,
      TRUE ~ 0)
  ) %>%
  select(no3_decline, slope_dep_vs_time,
         NO3, TOC,
    latitude, longitude, altitude, 
    pre, tmp, catchment_area,
    coniferous, deciduous, lake, mixed_forest, wetland)

# Complete cases
apply(is.na(df_analysis), 2, sum)
table(complete.cases(df_analysis))

# Keep only complete cases
df_analysis <- df_analysis[complete.cases(df_analysis),]

```


## 5. Logistic regression, all variables  
```{r, results  = "hold"}


# Full model  
fm <- glm(
  no3_decline ~ .,
  data = df_analysis, 
  family = "binomial",
  na.action = "na.fail")

# Dredging for best model
# dd1 <- dredge(fm)                           # only once
# saveRDS(dd1, "Data/160_all_dd1.rds")    # save it as it takes a couple of minutes

# Read dredge result
dd1 <- readRDS("Data/160_all_dd1.rds")

# Check best models  
subset(dd1, delta < 1)

cat("\n\nR2: \n")
dd1_mod1 <- get.models(dd1, 1)[[1]]  
# summary(dd1_mod1)  

par(mfrow = c(2,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(dd1_mod1, scale = "response")
mtext("All data, include coordinates", outer = TRUE)

```

## 6. Logistic regression, exclude coordinates    
```{r, results  = "hold"}

fm <- glm(
  no3_decline ~ slope_dep_vs_time + NO3 + TOC +
    altitude + 
    pre + tmp + catchment_area +
    coniferous + deciduous + lake + mixed_forest + 
    wetland,
  data = df_analysis, 
  family = "binomial",
  na.action = "na.fail")

# dd2 <- dredge(fm)                       # only once
# saveRDS(dd2, "Data/160_all_dd2.rds")    # save it as it takes a couple of minutes

dd2 <- readRDS("Data/160_all_dd2.rds")

subset(dd2, delta < 1)

cat("\n\nR2: \n")
dd2_mod1 <- get.models(dd2, 1)[[1]]  
# summary(dd2_mod1)  

par(mfrow = c(2,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(dd2_mod1, scale = "response")
mtext("All data, exclude coordinates", outer = TRUE)

```




## 7. Tree and forest classification



### Split into training and validation data
```{r}

set.seed(123)

x <- runif(nrow(df_analysis))
train <- ifelse(x < 0.9, TRUE, FALSE)

train_set <- df_analysis[train,]  %>% 
  mutate(no3_decline_f = factor(no3_decline)) %>% select(-no3_decline, -longitude, - latitude) %>%
  as.data.frame()
valid_set <- df_analysis[!train,] %>% 
  mutate(no3_decline_f = factor(no3_decline)) %>% select(-no3_decline, -longitude, - latitude) %>%
  as.data.frame()

```


### a. Tree classification using 'party'   
```{r, fig.width=8, fig.height=8}

(ct = ctree(no3_decline_f ~ ., data = train_set))

plot(ct, main="Conditional Inference Tree")

cat("\n\n")
cat("Table of prediction errors \n")
table(predict(ct), train_set$no3_decline_f)
cat("\n\n")

cat("Classification of training set \n")
tr.pred = predict(ct, newdata = valid_set, type="prob")
table(tr.pred[,"1"] > 0.5, valid_set$no3_decline_f)


```

### b. Evtree (Evolutionary Learning)   
```{r, fig.width=8, fig.height=8}

ev.raw = evtree(no3_decline_f ~ ., data = train_set)
plot(ev.raw)
table(predict(ev.raw), train_set$no3_decline_f)
1-mean(predict(ev.raw) == train_set$no3_decline_f)

```


### c. Random forest  
* *For results/interpretation, see separate document '160_randomforest_James_data.html'*  
```{r}

model1 <- randomForest(no3_decline_f ~ ., 
                       data = train_set, 
                       mtry = 5,
                       importance = TRUE)

model1

```


#### Random forest, predict on training data
```{r}

# Predicting on train set
pred_valid <- predict(model1, valid_set, type = "class")
# Checking classification accuracy
table(pred_valid, valid_set$no3_decline_f)  

```

#### Random forest, importance (but see separate result file)  
High MeanDecreaseGini = high importance in model  
```{r}

randomForest::importance(model1)
varImpPlot(model1)

```

#### Random forest, partial effects   
* NOt easy to interpret    
```{r, fig.height=7, fig.width = 10}

imp <- randomForest::importance(model1)
impvar1 <- rownames(imp)[order(imp[, 2], decreasing=TRUE)]

# impvar1 <- readRDS("Data/130_impvar1.rmd")

par(mfrow=c(3,4))

for (i in seq_along(impvar1)) {
  partialPlot(x = model1, 
              pred.data = train_set, 
              x.var = impvar1[i], 
              which.class = "1",
              main=paste("Effect of", impvar1[i]))
}



```
#### randomForestExplainer
```{r}

# install.packages("randomForestExplainer")
library(randomForestExplainer)

# Copy to console and run (this will create an Rmd file and render an HTML file):
# explain_forest(model1, interactions = TRUE, data = train_set, path = "160_randomforest_James_data.html")

```




