---
title: "`r params$document_title`"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true
params:
  document_title: 
    value: '160x Analyse NO3 decline - test run'
  text_dataset: 
    value: 'Data with slope_dep_vs_time, NO3, and TOTN_dep'
  extra_selection_variable: 'continent'
  extra_selection_values: 'Europe;North America'
  selected_vars: 
    value: 'no3_decline,catchment_area, TOC,slope_dep_vs_time, NO3, TOTN_dep, latitude, longitude, altitude,pre, tmp, Slope_pre, Slope_tmp, urban, cultivated, coniferous, decid_mixed, total_shrub_herbaceous,wetland, lake_water, bare_sparse'
  extra_pairwise_plots:
    value: 'TOC,NO3; slope_dep_vs_time,TOTN_dep; altitude,decid_mixed'
  pairwise_plots_same_scale:
    value: 'FALSE'
  logistic_formula: 
    value: 'no3_decline ~ TOC*altitude + TOTN_dep*slope_dep_vs_time + NO3 + decid_mixed + coniferous + tmp + lake_water + wetland'

---

```{r, echo = FALSE}  

#
# This file is designed to be run using '160parm_run_markdown.R'  
# The scripts in that file renders 3 result html file: 19\60a, b and c (different variable sets and data)  
#

```


**Analysis of NO3 decrease (categorical, "decrease or not"), based on James' trend results**  
  
**`r params$text_dataset`**   

* Response variable: 'Significant /NO3 decline' (locations with signif. *increase* are *not* excluded)  
* Data from https://github.com/JamesSample/icpw2/tree/master/thematic_report_2020/results      
* Sen slope of NO3, TOTN, TOC/TON etc. 1992-2016
* Response variable in all analyses are *whether NO3 decreases or not*     
* Predictors:
    - slope_dep_vs_time: Trend in Tot-N deposition 1992-2016    
    - NO3, TOTN_dep: Medians of NO3, TOTN_dep (Tot-N deposition) 1992-2016   
    - catchment_area (if included in data)      
    - TOC: Medians of TOC       
    - pre, tmp: mean precipitation + temp   
    - Land cover 
  
Technical details: This html file was rendered with `160parm_run_markdown.R` which runs the script `160parm_Time_series_results_James.Rmd` with different inputs, resulting in html files 160a, 160b and 160c.    

## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of these packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)      # pivot_wider
library(purrr)      # 'map' functions  
library(lubridate)  
library(ggplot2)

# Too many packages, not all are used
# library(mapview)
library(visreg)     # visreg
library(rkt)        # Theil -Sen Regression

library(MuMIn)      

# Trees and forests
library(party)                  # ctree
library(evtree)                 # evtree
library(randomForest)
library(randomForestExplainer)  # measure_importance, plot_multi_way_importance
library(pdp)                    # partial, autoplot

library(maps)
my_map <- map_data("world")

library(effects)    # handles lme models  
library(readxl)
library(readr)

source("002_Functions.R")
source("160parm_functions.R")

knitr::opts_chunk$set(results = 'hold') # collect the results from a chunk  
knitr::opts_chunk$set(warning = FALSE)  

options(width = 95)

```



## 2. Data

### James' trends and medians     
```{r}

#
# Regression results
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
file <- "trends_1992-2016_no3_results.csv"
fn <- paste0(folder, "/", file)

reg_no3 <- read.csv(fn, encoding = "UTF-8")
cat("Regression results:", sQuote(file), ",n =", nrow(reg_no3), "\n\n")

# Station metadata

# THESE WILL BE ADDED TOGETHER WITH LAND COVER  

# file <- "trends_1992-2016_no3_stations.csv"
# fn <- paste0(folder, "/", file)
# 
# df_metadata <- read.csv(fn, encoding = "UTF-8")
# cat("Regression result metadata:", sQuote(file), ",n =", nrow(df_metadata), "\n\n")
# 
# cat("Countries with trends: \n")
# xtabs(~country, df_metadata)  

#
# Medians NO3
#

if (FALSE){

  # OLD: 2012-2016 medians NO3  

  cat("--------------------------------------------------------------------------\n")
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
  file <- "medians_2012-2016_no3.csv"
  fn <- paste0(folder, "/", file)
  # data:
  medians_no3 <- read.csv(fn, encoding = "UTF-8")
  cat("Medians NO3:", sQuote(file), ",n =", nrow(medians_no3), "\n\n")
  file <- "medians_2012-2016_no3_stations.csv"
  fn <- paste0(folder, "/", file)
  # metadata:
  medians_no3_st <- read.csv(fn, encoding = "UTF-8")
  cat("Median metadata (not used):", sQuote(file), "n =", nrow(medians_no3_st), "\n\n")
  # stats:
  cat("Countries with medians: \n")
  xtabs(~country, medians_no3_st)  
  
  #
  #  OLD: 2012-2016 Medians TOC
  #
  cat("--------------------------------------------------------------------------\n")
  
  file <- "medians_2012-2016_toc_totn_no3_nh4.csv"
  fn <- paste0(folder, "/", file)
  
  medians_toc <- read.csv(fn, encoding = "UTF-8")
  cat("Medians TOC:", sQuote(file), ",n =", nrow(medians_toc), "\n\n")
  
  cat("Countries with medians: \n")
  xtabs(~country, medians_no3_st)  
  
}


```

### Start 'dat'  

With slope regression data  
* Make one line per station  
```{r}

# table(reg_no3$variable)

# Slope 
df1 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, sen_slp) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "sen_slp") %>%
  rename(slope_no3_vs_time = `NO3-N_µg/l N`, 
         slope_tocton_vs_time = `TOC/TON`)
  
# Slope p-value
df2 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, mk_p_val) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mk_p_val") %>%
  rename(p_no3_vs_time = `NO3-N_µg/l N`, 
         p_tocton_vs_time = `TOC/TON`)

# Medians
df3 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC_mg C/l")) %>%
  select(station_id, variable, median) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "median") %>%
  rename(NO3 = `NO3-N_µg/l N`, 
         TOC = `TOC_mg C/l`)

cat("\n")
cat("df1, n =", nrow(df1), "\n")
cat("df2, n =", nrow(df2), "\n")
cat("df3, n =", nrow(df3), "\n")

dat_1 <- df1 %>%
  full_join(df2, by = "station_id") %>%
  full_join(df3, by = "station_id")

cat("dat_1, n =", nrow(dat_1), "\n")

```

```{r}
# dat_1
# 
# str <- "slope_no3_vs_time ~ p_no3_vs_time"
# plot(as.formula(str), data = dat_1)
# lm(as.formula(str), data = dat_1)

```

### Deposition trends and median 1992-2006     
```{r}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/deposition/totn_dep_trends_icpw_stns.csv"  

df_deposition <- read.csv(fn) %>% 
  filter(variable == "totn_mgNpm2")  

cat("n =", nrow(df_deposition), "\n")

```


### Add deposition slope and medians to data  
```{r}

# debugonce(left_join2)
dat_2 <- dat_1 %>% 
  left_join2(df_deposition %>% 
              select(station_id, median, sen_slp, mk_p_val) %>%
              rename(TOTN_dep = median,
                     slope_dep_vs_time = sen_slp,
                     p_dep_vs_time = mk_p_val),
             by = "station_id",
             print_vars = TRUE)

```

### Add medians and station metadata   

```{r}

# dat_2 <- dat_2 %>%
#   left_join(df_metadata, by = "station_id")

# cat("dat_2, n =", nrow(dat_2), "\n")

# Simplify names by removing units
# names(dat_2)
# names(dat_2) <- sub(".N_µg.l.N", "", names(dat_2))
# names(dat_2) <- sub("_mg.C.l", "", names(dat_2))
# names(dat_2) <- sub("_µg.l.P", "", names(dat_2))

# cat("\nVariable names: \n")
# names(dat_2)


```

### Add climate and deposition medians and slopes  
```{r}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/climate/cru_climate_trends_icpw_stns.csv"

df_climate_mean <- read_csv(fn) %>% 
  select(station_id, variable, median) %>%
  pivot_wider(names_from = "variable", values_from = "median")
cat("\n")
# names(df_climate_mean)

df_climate_slope <- read_csv(fn) %>%
  select(station_id, variable, sen_slp) %>%
  pivot_wider(names_from = "variable", values_from = "sen_slp", names_prefix = "Slope_")

# Add
dat_3 <- dat_2 %>%
  left_join2(df_climate_mean, by = "station_id", print_vars = TRUE) %>%
  left_join2(df_climate_slope, by = "station_id", print_vars = TRUE)

```

### Combine land cover types   
* Data including UK read using script 159   
* Note: also includes metadata (country, etc.)
* bare_sparse = bare_rock + sparsely_vegetated + glacier   
* Select: coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse   

```{r}

# df_landcover3_OLD <- readRDS("Data/159_df_landcover3.rds")
df_landcover3 <- readRDS("Data/159_df_meta3.rds")

df_landcover3 <- df_landcover3 %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier,
         decid_mixed = deciduous + mixed_forest,
         lake_water = lake + water_ex_lake) %>%
  select(-c(bare_rock, sparsely_vegetated, glacier, deciduous, mixed_forest, lake, water_ex_lake))

```


### Add land cover columns to main data    
```{r}

dat_4 <- left_join2(dat_3, 
                   df_landcover3, 
                   by = "station_id", 
                   print_vars = TRUE
)


```

### Drop locations with >10% cultivated    
```{r}

dat_5 <- dat_4 %>%
  filter(cultivated <= 10)

cat(nrow(dat_4) - nrow(dat_5), "stations with > 10% cultivated deleted \n")

```
### Extra selection  
```{r}

selection_values <- params$extra_selection_values %>% strsplit(split = ";") %>% .[[1]]
sel <- dat_5[[params$extra_selection_variable]] %in% selection_values
sum(sel)
mean(sel)
dat_6 <- dat_5[sel,]

```


### Data set used  
```{r}

dat <- dat_6

```


## 3. Plot slopes    

```{r, fig.width=9, fig.height=7}

ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time)) + 
  geom_point(data = dat %>% filter(p_no3_vs_time < 0.05), size = rel(2)) +
  geom_point(aes(color = country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) 
  
ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time,
                color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")

dat %>%
  filter(!country %in% c("Latvia","Ireland","Italy","Netherlands")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_time,
             color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries") + 
  ylim(-50, 25)


```


## 4. Select data   

### a. Selection of variables  
* Select variables to use, and thereby also cases  
```{r}

get_df_no3_decline <- function(data, variable_string){
  variable_string <- gsub(" ", "", variable_string)
  variables <- strsplit(variable_string, split = ",")[[1]]
  # Data for analyses
  df <- data %>%
    mutate(
      no3_decline = case_when(
        slope_no3_vs_time < 0 & p_no3_vs_time <= 0.05 ~ 1,
        TRUE ~ 0)
    )
  df[variables]
}

cat("-------------------------------------------------------------\n")
cat("Variables: \n")
cat(params$selected_vars)
cat("\n-------------------------------------------------------------\n")

df_analysis <- get_df_no3_decline(dat, params$selected_vars)  

# names(dat) %>% paste(collapse = ", ")

cat("Number of missing values per variable: \n")
apply(is.na(df_analysis), 2, sum) 
cat("\n")

# What is missing? (long output)
if (FALSE){
dat %>% 
  split(.$country) %>%
  purrr::map(~apply(is.na(.), 2, mean))
}

cat("Number of complete observations: \n")
complete <- complete.cases(df_analysis)
table(complete)

cat("\n\n")
cat("Number of complete observations by country: \n")
table(dat$country, complete)

# Keep only complete cases
df_analysis <- df_analysis[complete.cases(df_analysis),]

cat("\n\n")
cat("Original data: n =", nrow(dat), "\n")
cat("Analysis: n =", nrow(df_analysis), "\n")


```


### b. Correlations   
```{r, fig.width=10, fig.height=10}

gg <- GGally::ggcorr(
  df_analysis, 
  method = c("complete.obs", "kendall"), 
  label = TRUE,
  hjust = 0.9, angle = -30) # +                    # slanted labels
gg + coord_cartesian(x = c(-2, 20), y = c(-2,22))  # fix margins


```  



## 5. Tree and forest classification


### Split into training and validation data
```{r}

set.seed(123)

x <- runif(nrow(df_analysis))
train <- ifelse(x < 0.9, TRUE, FALSE)

train_set <- df_analysis[train,]  %>% 
  mutate(no3_decline_f = factor(no3_decline)) %>% select(-no3_decline, -longitude, - latitude) %>%
  as.data.frame()
valid_set <- df_analysis[!train,] %>% 
  mutate(no3_decline_f = factor(no3_decline)) %>% select(-no3_decline, -longitude, - latitude) %>%
  as.data.frame()

```


### a. Tree classification using 'party'   
```{r, fig.width=8, fig.height=8}

(ct = ctree(no3_decline_f ~ ., data = train_set))

plot(ct, main="Conditional Inference Tree")

cat("\n\n")
cat("Table of prediction errors \n")
table(predict(ct), train_set$no3_decline_f)
cat("\n\n")

cat("Classification of training set \n")
tr.pred = predict(ct, newdata = valid_set, type="prob")
colnames(tr.pred) <- c("P0", "P1")
# tr.pred <- tr.pred %>% map_dfr(~data.frame(P0 = .[1], P1 = .[2]))
table(tr.pred[,"P1"] > 0.5, valid_set$no3_decline_f)


```

### b. Evtree (Evolutionary Learning)   
```{r, fig.width=8, fig.height=8}

ev.raw = evtree(no3_decline_f ~ ., data = train_set)

plot(ev.raw)
cat("Predicted in training data: \n")
table(predict(ev.raw), train_set$no3_decline_f)

cat("\n\nPrediction errors in training data: \n")
1-mean(predict(ev.raw) == train_set$no3_decline_f)
```


### c. Random forest  
* *For results/interpretation, see separate document '160_randomforest_James_data.html'*  


#### c1a. Predict on training data
```{r}  

model1 <- randomForest(no3_decline_f ~ ., 
                       data = train_set, 
                       mtry = 5,
                       importance = TRUE)

# Predicting on train set
pred_valid <- predict(model1, valid_set, type = "class")
# Checking classification accuracy
table(pred_valid, valid_set$no3_decline_f)  

error_fraction <- mean(
  (pred_valid == 0 & valid_set$no3_decline_f == 1) | 
    (pred_valid == 1 & valid_set$no3_decline_f == 0))
cat("Error rate for training data:", round(error_fraction*100, 1), "%\n")

```

#### c1b. Model for all data    
```{r}  

full_set <- df_analysis  %>% 
  mutate(no3_decline_f = factor(no3_decline)) %>% 
  select(-no3_decline, -longitude, - latitude) %>%
  as.data.frame()

model1 <- randomForest(no3_decline_f ~ ., 
                       data = full_set, 
                       mtry = 5,
                       importance = TRUE)

# Predicting on train set
pred_valid <- predict(model1, valid_set, type = "class")
# Checking classification accuracy
table(pred_valid, valid_set$no3_decline_f)  

error_fraction <- mean(
  (pred_valid == 0 & valid_set$no3_decline_f == 1) | 
    (pred_valid == 1 & valid_set$no3_decline_f == 0))
cat("Error rate for training data:", round(error_fraction*100, 1), "%\n")

```

#### c2. Importance of variables
```{r 5c2_calculate_importance}

# Calculation
importance <- measure_importance(model1)

```



```{r}

plot_multi_way_importance(importance, size_measure = "no_of_nodes", no_of_labels = 6)  
plot_multi_way_importance(importance, 
                          x_measure = "accuracy_decrease", 
                          y_measure = "gini_decrease", 
                          size_measure = "p_value", no_of_labels = 6)

```



#### c3. Random forest, show partial effects  

```{r 5c3_calculate_partial_effects1, warning=FALSE, message=FALSE, results='hide'}

# Which variables to include:
variables_for_plot <- importance %>%
  mutate(variable = levels(variable)[as.numeric(variable)]) %>%
  arrange(desc(gini_decrease)) %>%
  pull(variable) %>%
  head(12)   # pick the first 12 variables (or less)

# Calculation

plotdata <- NULL  # will be list for storing results

max_number_of_plots <- length(variables_for_plot)/2 %>% floor()

for (i in 1:max_number_of_plots){
  varno1 <- c(1,3,5,7,9,11)[i]
  varno2 <- varno1 + 1
  plotdata[[i]] <- model1 %>%
    partial(pred.var = variables_for_plot[c(varno1, varno2)], chull = TRUE, progress = "text",
            which.class = "1", prob = TRUE)
}

```


```{r 5c3_calculate_partial_effects2, warning=FALSE, message=FALSE, results='hide'}

### Extra plots

i <- max_number_of_plots

plotpairs <- params$extra_pairwise_plots %>%
  gsub(" ", "", ., fixed = TRUE) %>%
  strsplit(split = ";") %>%
  .[[1]] %>%
  purrr::map(~strsplit(., split = ",")[[1]])

for (plotvar in plotpairs){
  # print(plotvar)
  if (plotvar[1] %in% names(full_set) & plotvar[2] %in% names(full_set)){
    i <- i + 1
    plotdata[[i]] <- model1 %>%
      partial(pred.var = c(plotvar[1], plotvar[2]), chull = TRUE, progress = "text",
             which.class = "1", prob = TRUE)
  }
}


```

```{r 5c3_plot_partial_effects2}

# Find range of predicted values for each graph
ranges <- plotdata %>% purrr::map_dfc(~range(.$yhat))

# use range of all the ranges
for (i in 1:length(plotdata)){
  
  if (params$pairwise_plots_same_scale == "TRUE"){
  
    gg <- plot_pair_number(i, zrange = range(ranges))
    print(gg)
    
  } else {
    
    gg <- plot_pair_number(i)
    print(gg)

  }
  
}
  

```


```{r, echo = FALSE, results  = "hold"}

# #   no3_decline ~ as.formula(params$logistic_formula),
# 
# fm <- glm(
#   as.formula(params$logistic_formula),
#   data = df_analysis, 
#   family = "binomial",
#   na.action = "na.fail")
# 
# dd1b <- dredge(fm)                       # only once
# saveRDS(dd1b, "Data/160_all_dd1b.rds")    # save it as it takes a couple of minutes
# # dd1b <- readRDS("Data/160_all_dd1b.rds")
# 
# # subset(dd1b, delta < 1)
# subset(dd1b, delta < 2)
# 
# cat("\n\nR2: \n")
# dd1b_mod1 <- get.models(dd1b, 1)[[1]]  
# # summary(dd1b_mod1)  
# 
# par(mfrow = c(2,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
# visreg(dd1b_mod1, scale = "response")

```


## 6. Logistic regression       
Start model: **`r params$logistic_formula`**
```{r, results  = "hold"}

fm <- glm(
  as.formula(params$logistic_formula),
  data = df_analysis, 
  family = "binomial",  
  na.action = "na.fail")

dredged_models <- dredge(fm)                       # only once
# saveRDS(dredged_models, "Data/162_all_dredged_models.rds")    # save it as it takes a couple of minutes
# dredged_models <- readRDS("Data/162_all_dredged_models.rds")

# cat("\n\nR2: \n")
# mod1 <- get.models(dredged_models, 1)[[1]]  
# summary(mod1)  

```  

### Best models  
```{r}
# subset(dredged_models, delta < 1)

subset(dredged_models, delta < 2)

# Alternative way of showing result (didn't become any better)
# df <- subset(dredged_models, delta < 2)
# select(as.data.frame(df) %>% round(6), -`(Intercept)`, -logLik, -AICc)

```


### Plots  
```{r, message=FALSE, results  = "hold"}

# Pick model with lowest AICc
mod1 <- get.models(dredged_models, 1)[[1]]  

summ <- summary(mod1)
dev_expl <- summ$deviance/summ$null.deviance
cat("Percentage of deviance explained:", round(100*dev_expl, 1), "% \n")

modelvars <- get_model_variables(mod1)

# Interactions: 3D plot 
# visreg2d(mod1, xvar = vars[1], yvar = vars[2], 
#          type = 'conditional', scale = "response") 

# Interactions: 2D plot 
if (length(modelvars$interaction_list) > 0){
  modelvars$interaction_list %>% purrr::walk(
    ~visreg(mod1, .x[1], by = .x[2], scale = "response")
  )
}

# Additive effects: 1D plot
if (length(modelvars$additive_vars) > 0){
  par(mfrow = c(2,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
  for (var in modelvars$additive_vars)
    visreg(mod1, var, scale = "response")  
}

```



