---
title: "`r params$document_title`"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true
params:
  document_title: 
    value: '162x Analyse NO3 status - test run'
  text_line1: 
    value: 'Analysis of NO3 medians (2012-2016)'
  text_line2: 
    value: 'Data with NO3, TOTN_dep, slope_dep_vs_time, TOC, tmp (temperature), pre (precipitation)'
  medians_filename:
    value: 'medians_2012-2016_no3.csv'        
  selected_vars: 
    value: 'log_median_no3,catchment_area, log_median_toc,slope_dep_vs_time, TOTN_dep, latitude, longitude, altitude,pre, tmp, urban, cultivated, coniferous, decid_mixed, total_shrub_herbaceous,wetland, lake_water, bare_sparse'
  tree_formula:
    value: 'log_median_no3 ~ .'
  extra_pairwise_plots:
    value: 'TOC,NO3; slope_dep_vs_time,TOTN_dep; altitude,decid_mixed'
  logistic_formula: 
    value: 'log_median_no3 ~ TOTN_dep + slope_dep_vs_time + TOTN_dep:slope_dep_vs_time + log_median_toc + TOTN_dep:log_median_toc + tmp + pre + altitude + decid_mixed + bare_sparse + coniferous + catchment_area + lake_water + total_shrub_herbaceous'

---

```{r, echo = FALSE}  

#
# This file is designed to be run using '162parm_run_markdown.R'  
# The scripts in that file renders 3 result html file: 19\60a, b and c (different variable sets and data)  
#

```


**`r params$text_line1`**   

**`r params$text_line2`**   
**Name of dataset: `r params$medians_filename`**   


* Response variable: 'Current NO3 level' (locations with signif. *increase* are *not* excluded)  
* Data from https://github.com/JamesSample/icpw2/tree/master/thematic_report_2020/results      
* Sen slope of NO3, TOTN, TOC/TON etc. 1992-2016
* Response variable in all analyses are *whether NO3 decreases or not*     
* Predictors:
    - slope_dep_vs_time: Trend in Tot-N deposition 1992-2016    
    - NO3, TOTN_dep: Medians of NO3, TOTN_dep (Tot-N deposition) 1992-2016   
    - catchment_area (if included in data)      
    - TOC: Medians of TOC 1992-2016 (if included in data)     
    - pre, tmp: mean precipitation + temp   
    - Land cover 
  
Technical details: This html file was created was 

## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of these packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)      # pivot_wider
library(purrr)      # 'map' functions  
library(lubridate)  
library(ggplot2)

# Too many packages, not all are used
library(mapview)
library(visreg)     # visreg
library(rkt)        # Theil -Sen Regression

library(MuMIn)      

# Trees and forests
library(party)                  # ctree
library(evtree)                 # evtree
library(randomForest)
library(randomForestExplainer)  # measure_importance, plot_multi_way_importance
library(pdp)                    # partial, autoplot

library(maps)
my_map <- map_data("world")

library(effects)    # handles lme models  
library(readxl)
library(readr)

knitr::opts_chunk$set(results = 'hold') # collect the results from a chunk  
knitr::opts_chunk$set(warning = FALSE)  

options(width = 95)

```



## 2. Data  
* The data part (part 2) is quite similar in scripts 160, 161 and 162 

### Available files
```{r, echo=FALSE, results='hold'}

if (TRUE){   # in case we don't want this part
  
  # Code for comparing the two median files
  
  # Check medians file #1
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
  file <- "medians_2012-2016_no3.csv"
  fn <- paste0(folder, "/", file)
  df_medians <- read.csv(fn, encoding = "UTF-8")
  cat("Medians results:", sQuote(file), ",n =", nrow(df_medians), "\n\n")  
  cat("Number of values per variable: \n")
  apply(!is.na(df_medians), 2, sum)
  cat("\n")
  
  # Check medians file #2
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
  file <- "medians_2012-2016_toc_totn_no3.csv"
  fn <- paste0(folder, "/", file)
  df_medians <- read.csv(fn, encoding = "UTF-8")
  cat("Medians results:", sQuote(file), ",n =", nrow(df_medians), "\n\n")  
  cat("Number of values per variable: \n")
  apply(!is.na(df_medians), 2, sum)
  cat("\n")
  
}


```

### James' trends and medians     
```{r}

#
# Median results 2012-2016  
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
file <- params$medians_filename
fn <- paste0(folder, "/", file)
df_medians <- read.csv(fn, encoding = "UTF-8")
cat("Medians results:", sQuote(file), ",n =", nrow(df_medians), "\n\n")  
cat("Number of values per variable: \n")
apply(!is.na(df_medians), 2, sum)
cat("\n")

# Station metadata
# WILL BE ADDED TOGETHER WITH LAND COVER, BELOW
# file_meta <- gsub(".csv", "_stations.csv", file)
# fn <- paste0(folder, "/", file_meta)
# df_metadata <- read.csv(fn, encoding = "UTF-8")
# cat("Regression result metadata:", sQuote(file_meta), ",n =",  
#      nrow(df_metadata), "\n\n")


#
# Regression results 1992-2016
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
file <- "trends_1992-2016_no3_results.csv"
fn <- paste0(folder, "/", file)
df_trends <- read.csv(fn, encoding = "UTF-8")
cat("Trends used as predictor variable:", sQuote(file), ",n =", nrow(df_trends), "\n\n")


```
### Start 'dat'  
Using medians    
* Make one line per station   
* Also including some trends for  
```{r}

# table(reg_no3$variable)

# Medians 2012-2016  
df1 <- df_medians %>%
  select(station_id, `NO3.N_µg.l.N`, `TOTN_µg.l.N`, `TON_µg.l.N`, `TOC_mg.C.l`, TOC.TON) %>%
  rename(median_no3 = `NO3.N_µg.l.N`,
         median_totn = `TOTN_µg.l.N`,
         median_ton = `TON_µg.l.N`,
         median_toc = `TOC_mg.C.l`,
         median_tocton = `TOC.TON`) %>%
  mutate(log_median_no3 = log10(median_no3 + 0.1),
         log_median_totn = log10(median_totn),
         log_median_ton = log10(median_ton),
         log_median_toc = log10(median_toc),
         log_median_tocton = log10(median_tocton))

# Some trends
df2 <- df_trends %>% # table()
  filter(variable %in% c("NO3-N_µg/l N", "TOTN_µg/l N", "TOC_mg C/l")) %>%
  select(station_id, variable, median) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "median") %>% # str()
  rename(trend_NO3 = `NO3-N_µg/l N`,
         trend_TOTN = `TOTN_µg/l N`,
         trend_TOC = `TOC_mg C/l`)

cat("\n")
cat("df1, n =", nrow(df1), "\n")
cat("df2, n =", nrow(df2), "\n")

dat <- df1 %>%
  left_join(df2, by = "station_id")

cat("dat, n =", nrow(dat), "\n")

```

### Deposition trends 1992-2006 and median 2012-2016      
```{r, results = 'hold'}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/deposition/totn_dep_trends_icpw_stns.csv"  

# For trends
df_deposition <- read.csv(fn) %>% 
  filter(variable == "totn_mgNpm2" & !is.na(sen_slp)) %>%
  select(station_id, sen_slp, mk_p_val) %>%
              rename(slope_dep_vs_time = sen_slp,
                     p_dep_vs_time = mk_p_val) 

cat("n =", nrow(df_deposition), "\n")

# Deposition medians 2012-2016
fn <- "K:/Prosjekter/langtransporterte forurensninger/O-23300 - ICP-WATERS - HWI/Faglige rapporter/2020 report/Deposition data/Processed/dep_TOTN.csv"
df_deposition_medians <- read_csv(fn) %>%
  filter(year %in% 2012:2016) %>%
  group_by(station_id) %>%
  summarize(TOTN_dep = median(TOTN_dep))

df_deposition <- df_deposition %>%
  left_join(df_deposition_medians)
  
cat("n =", nrow(df_deposition), "\n")

xtabs(~is.na(TOTN_dep), df_deposition)

```

### Add deposition slope and medians to data  
```{r}

dat <- dat %>% 
  left_join(df_deposition,
                 by = "station_id")

cat("dat, n =", nrow(dat), "\n")

# names(dat)

```



### Add climate medians   
```{r}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/climate/cru_climate_trends_icpw_stns.csv"
df_climate_mean <- read_csv(fn) %>%
  select(station_id, variable, median) %>%
  pivot_wider(names_from = "variable", values_from = "median")

# names(df_climate_mean)

# Add
dat <- dat %>%
  left_join(df_climate_mean, by = "station_id")

cat("dat, n =", nrow(dat), "\n")

```




### Combine land cover types   
* Data including UK read using script 159  
* bare_sparse = bare_rock + sparsely_vegetated + glacier   
* Select: coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse   
```{r}

df_landcover3 <- readRDS("Data/159_df_landcover3.rds")

df_landcover3 <- df_landcover3 %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier,
         decid_mixed = deciduous + mixed_forest,
         lake_water = lake + water_ex_lake) %>%
  select(-c(bare_rock, sparsely_vegetated, glacier, 
            deciduous, mixed_forest, 
            lake, water_ex_lake))


```


### Add land cover columns to main data    
```{r}

dat <- left_join(dat, 
                 df_landcover3, 
                 by = "station_id"
)

cat("dat, n =", nrow(dat), "\n")

```



## 3. Plot data      

```{r, fig.width=9, fig.height=7}

gg <- ggplot(dat, aes(TOTN_dep, log_median_no3)) + 
  geom_point(aes(color = country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) 
  
gg

```


## 4. Select data   

### a. Selection of variables    
* Select variables to use, and thereby also cases  
* Also remove PL05, which has dubious values   
```{r}


get_data_for_analysis <- function(data, variable_string){
  variable_string <- gsub(" ", "", variable_string)
  variables <- strsplit(variable_string, split = ",")[[1]]
  # Check if all variables are there
  found <- variables %in% names(data)
  if (sum(!found) > 0)
    stop("Not all variables found in data:", 
      paste(variables[!found], collapse = " ,"), 
      "\n")
  # Data for analyses
  data[variables]
}

cat("-------------------------------------------------------------\n")
cat("Variables: \n")
cat(params$selected_vars)
cat("\n-------------------------------------------------------------\n")

sel <- dat$station_code %in% "PL05"
dat <- dat[!sel,]
message(sum(sel), " station removed - station PL05 (has dubious NO3 data)")  

# debugonce(get_data_for_analysis)
# df_analysis <- get_data_for_analysis(dat, vars)  
df_analysis <- get_data_for_analysis(dat, params$selected_vars)  

# names(dat) %>% paste(collapse = ", ")

cat("Number of missing values per variable: \n")
apply(is.na(df_analysis), 2, sum) 
cat("\n")

# What is missing? (long output)
if (FALSE){
dat %>% 
  split(.$country) %>%
  purrr::map(~apply(is.na(.), 2, mean))
}

cat("Number of complete observations: \n")
complete <- complete.cases(df_analysis)
table(complete)

cat("\n\n")
cat("Number of complete observations by country: \n")
table(dat$country, complete)

# Keep only complete cases
df_analysis <- df_analysis[complete.cases(df_analysis),]

cat("\n\n")
cat("Original data: n =", nrow(dat), "\n")
cat("Analysis: n =", nrow(df_analysis), "\n")


```


### b. Correlations   
```{r, fig.width=10, fig.height=10}

gg <- GGally::ggcorr(df_analysis, method = c("complete.obs", "kendall"), label = TRUE) # +
gg + theme(plot.margin = unit(c(.8, 2, .8, 2.5), "cm"))

# SHOULD also workaccording to ?element_rect (update ggplot2?)
# gg + theme(plot.margin = margin(.6, .5, .6, 1.7, "cm"))


```



## 5. Tree and forest classification


### Split into training and validation data
```{r}

set.seed(123)

x <- runif(nrow(df_analysis))
train <- ifelse(x < 0.9, TRUE, FALSE)

train_set <- df_analysis[train,]  %>% 
  select(-longitude, - latitude) %>%
  as.data.frame()
valid_set <- df_analysis[!train,] %>% 
  select(-longitude, - latitude) %>%
  as.data.frame()

# plot(train_set)

```


### a. Tree classification using 'party'   
```{r, fig.width=10, fig.height=8}

# train_set$X <- 10^train_set$log_median_no3
# (ct = ctree(X ~ ., data = train_set))

(ct = ctree(as.formula(params$tree_formula), 
            data = train_set))

plot(ct, main="Conditional Inference Tree")

```

### b. Evtree (Evolutionary Learning)   
```{r, fig.width=10, fig.height=8}

ev.raw = evtree(as.formula(params$tree_formula), 
                data = train_set)

plot(ev.raw)

```


### c. Random forest  
* Model called 'model1'
```{r}

model1 <- randomForest(as.formula(params$tree_formula), 
                       data = train_set, 
                       mtry = 5,
                       importance = TRUE)

model1

```


#### c1. Predict on training data
Not relevant here  

#### c2. Importance of variables   
* Classification forests result in   
    - mean_min_depth   
    - accuracy_decrease  
    - gini_decrease  
    - no_of_nodes   
    - times_a_root      
* Regression forests result in     
    - mean_min_depth   
    - mse_increase  
    - node_purity_increase  
    - no_of_nodes   
    - times_a_root      
```{r 5c2_calculate_importance}

# Calculation
importance <- measure_importance(model1)

```



```{r}

plot_multi_way_importance(importance, size_measure = "no_of_nodes", no_of_labels = 6)  
plot_multi_way_importance(importance, x_measure = "mse_increase", y_measure = "node_purity_increase",
                          size_measure = "p_value", no_of_labels = 6)

```



#### c3. Random forest, show partial effects  

```{r 5c3_calculate_partial_effects, warning=FALSE, message=FALSE, results='hide', cache = FALSE}

# Which variables to include:
variables_for_plot <- importance %>%
  mutate(variable = levels(variable)[as.numeric(variable)]) %>%
  arrange(desc(mse_increase)) %>%    # gini_decrease for classification trees
  pull(variable) %>%
  head(12)   # pick the first 12 variables (or less)

# Calculation

plotdata <- NULL  # will be list for storing results

max_number_of_plots <- length(variables_for_plot)/2 %>% floor()

for (i in 1:max_number_of_plots){
  varno1 <- c(1,3,5,7,9,11)[i]
  varno2 <- varno1 + 1
  plotdata[[i]] <- model1 %>%
    partial(pred.var = variables_for_plot[c(varno1, varno2)], chull = TRUE, progress = "text",
            which.class = "1", prob = TRUE)
}


```


```{r 5c3_calculate_partial_effects2, warning=FALSE, message=FALSE, results='hide'}

### Extra plots

plotpairs <- params$extra_pairwise_plots %>%
  gsub(" ", "", ., fixed = TRUE) %>%
  strsplit(x, split = ";") %>%
  .[[1]] %>%
  purrr::map(~strsplit(., split = ",")[[1]])

for (plotvar in plotpairs){
  # print(plotvar)
  if (plotvar[1] %in% names(train_set) & plotvar[2] %in% names(train_set)){
    i <- i + 1
    plotdata[[i]] <- model1 %>%
      partial(pred.var = c(plotvar[1], plotvar[2]), chull = TRUE, progress = "text",
             which.class = "1", prob = TRUE)
  }
}


```


```{r, warning=FALSE}

# Plot the plots 
for (i in 1:max_number_of_plots){
  autoplot(plotdata[[i]], contour = TRUE, legend.title = "log(NO3)") %>%
    print()
}

```





## 6. Linear regression      
```{r, results  = "hold"}

fm <- lm(
  as.formula(params$logistic_formula),
  data = df_analysis, 
  na.action = "na.fail")

dredged_models <- dredge(fm)                       # only once
# saveRDS(dredged_models, "Data/162_all_dredged_models.rds")    # save it as it takes a couple of minutes
# dredged_models <- readRDS("Data/162_all_dredged_models.rds")

# cat("\n\nR2: \n")
# mod1 <- get.models(dredged_models, 1)[[1]]  
# summary(mod1)  

```  

### Best models  
```{r}
# subset(dredged_models, delta < 1)

subset(dredged_models, delta < 2)

# Alternative way of showing result (didn't become any better)
# df <- subset(dredged_models, delta < 2)
# select(as.data.frame(df) %>% round(6), -`(Intercept)`, -logLik, -AICc)

```
```{r, echo = FALSE}

#
# Function to get interactions and effects 
#
get_model_variables <- function(model){  
  all_terms <- names(model$coefficients)
  # Remove '(Intercept)'
  all_terms <- all_terms[!all_terms %in% "(Intercept)"]
  sel_interactions <- grepl(":", all_terms, fixed = TRUE)
  # Removing interaction terms leaves us with the variables:
  all_variables <- all_terms[!sel_interactions]
  # Get interactions in a list
  interaction_terms <- all_terms[sel_interactions]
  interaction_list <- interaction_terms %>% strsplit(split = ":")
  # Get additive variables
  interaction_vars <- interaction_list %>% unlist()
  additive_vars <- all_variables[!all_variables %in% interaction_vars]
  list(
    interaction_list = interaction_list,
    additive_vars = additive_vars)
}

# Test
# fit <- lm(sr ~ pop15*dpi + pop75*dpi + ddpi, data = LifeCycleSavings)
# get_model_variables(fit)  

```

### Plots  
```{r, message=FALSE}

# Pick model with lowest AICc
mod1 <- get.models(dredged_models, 1)[[1]]  

modelvars <- get_model_variables(mod1)

# Interactions: 2D plot 
if (length(modelvars$interaction_list) > 0){
  modelvars$interaction_list %>% purrr::walk(
    ~visreg(mod1, .x[1], by = .x[2], scale = "response")
  )
}

# Additive effects: 1D plot
if (length(modelvars$additive_vars) > 0){
  par(mfrow = c(2,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
  for (var in modelvars$additive_vars)
    visreg(mod1, var, scale = "response")  
}

```








