---
title: "`r params$document_title`"
author: "DHJ"
date: "8 5 2020"
output: 
  html_document:
    toc: true    
    toc_float: true
    keep_md: true
params:
  document_title: 
    value: '166x Analyse NO3 decline - test run'
  text_dataset: 
    value: 'Data with slope_dep_vs_time, NO3, and TOTN_dep'
  selected_vars: 
    value: 'slope_no3_vs_time,catchment_area, TOC,slope_dep_vs_time, NO3, TOTN_dep, latitude, longitude, altitude,pre, tmp, slope_pre, slope_tmp, urban, cultivated, coniferous, decid_mixed, total_shrub_herbaceous,wetland, lake_water, bare_sparse'
  response_variable:
    value: 'slope_no3_vs_time'  
  extra_pairwise_plots:
    value: 'TOC,NO3; slope_dep_vs_time,TOTN_dep; altitude,decid_mixed'
  pairwise_plots_same_scale:
    value: 'FALSE'
  logistic_formula: 
    value: 'slope_no3_vs_time ~ TOC*altitude + TOTN_dep*slope_dep_vs_time + NO3 + decid_mixed + coniferous + tmp + lake_water + wetland'

---

```{r, echo = FALSE}  

#
# This file is designed to be run using '166parm_run_markdown.R'  
# The scripts in that file renders 3 result html file: 19\60a, b and c (different variable sets and data)  
#

```


**Analysis of NO3 decrease as continuous variable (not categorical as in 166)**  
  
**`r params$text_dataset`**   

* Response variable: 'Significant /NO3 decline' (locations with signif. *increase* are *not* excluded)  
* Data from https://github.com/JamesSample/icpw2/tree/master/thematic_report_2020/results      
* Sen slope of NO3, TOTN, TOC/TON etc. 1992-2016
* Response variable in all analyses are *whether NO3 decreases or not*     
* Predictors:
    - slope_dep_vs_time: Trend in Tot-N deposition 1992-2016    
    - NO3, TOTN_dep: Medians of NO3, TOTN_dep (Tot-N deposition) 1992-2016   
    - catchment_area (if included in data)      
    - TOC: Medians of TOC       
    - pre, tmp: mean precipitation + temp   
    - Land cover 
  
Technical details: This html file was rendered with `166parm_run_markdown.R` which runs the script `166parm_Time_series_results_James.Rmd` with different inputs, resulting in html files 166a, 166b and 166c.    

## 1. Libraries  
```{r, results='hide', message=FALSE, warning=FALSE}

# All of these packages cn be loaded at once using library(tidyverse). (I just like to be specific.)
library(dplyr)
library(tidyr)      # pivot_wider
library(purrr)      # 'map' functions  
library(lubridate)  
library(ggplot2)

# Too many packages, not all are used
# library(mapview)
library(visreg)     # visreg
library(rkt)        # Theil -Sen Regression

library(MuMIn)      

# Trees and forests
library(party)                  # ctree
library(evtree)                 # evtree
library(randomForest)
library(randomForestExplainer)  # measure_importance, plot_multi_way_importance
library(pdp)                    # partial, autoplot

library(maps)
my_map <- map_data("world")

library(effects)    # handles lme models  
library(readxl)
library(readr)

source("002_Functions.R")
source("160parm_functions.R")

knitr::opts_chunk$set(results = 'hold') # collect the results from a chunk  
knitr::opts_chunk$set(warning = FALSE)  

options(width = 95)

```



## 2. Data

### James' trends and medians     
```{r}

#
# Regression results
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
file <- "trends_1992-2016_no3_results.csv"
fn <- paste0(folder, "/", file)

reg_no3 <- read.csv(fn, encoding = "UTF-8")
cat("Regression results:", sQuote(file), ",n =", nrow(reg_no3), "\n\n")

# Station metadata

# THESE WILL BE ADDED TOGETHER WITH LAND COVER  

# file <- "trends_1992-2016_no3_stations.csv"
# fn <- paste0(folder, "/", file)
# 
# df_metadata <- read.csv(fn, encoding = "UTF-8")
# cat("Regression result metadata:", sQuote(file), ",n =", nrow(df_metadata), "\n\n")
# 
# cat("Countries with trends: \n")
# xtabs(~country, df_metadata)  

#
# Medians NO3
#

if (FALSE){

  # OLD: 2012-2016 medians NO3  

  cat("--------------------------------------------------------------------------\n")
  folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
  file <- "medians_2012-2016_no3.csv"
  fn <- paste0(folder, "/", file)
  # data:
  medians_no3 <- read.csv(fn, encoding = "UTF-8")
  cat("Medians NO3:", sQuote(file), ",n =", nrow(medians_no3), "\n\n")
  file <- "medians_2012-2016_no3_stations.csv"
  fn <- paste0(folder, "/", file)
  # metadata:
  medians_no3_st <- read.csv(fn, encoding = "UTF-8")
  cat("Median metadata (not used):", sQuote(file), "n =", nrow(medians_no3_st), "\n\n")
  # stats:
  cat("Countries with medians: \n")
  xtabs(~country, medians_no3_st)  
  
  #
  #  OLD: 2012-2016 Medians TOC
  #
  cat("--------------------------------------------------------------------------\n")
  
  file <- "medians_2012-2016_toc_totn_no3_nh4.csv"
  fn <- paste0(folder, "/", file)
  
  medians_toc <- read.csv(fn, encoding = "UTF-8")
  cat("Medians TOC:", sQuote(file), ",n =", nrow(medians_toc), "\n\n")
  
  cat("Countries with medians: \n")
  xtabs(~country, medians_no3_st)  
  
}


```

### Start 'dat'  

With slope regression data  
* Make one line per station  
```{r}

# table(reg_no3$variable)

# Slope 
df1 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, sen_slp) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "sen_slp") %>%
  rename(slope_no3_vs_time = `NO3-N_µg/l N`, 
         slope_tocton_vs_time = `TOC/TON`)
  
# Slope p-value
df2 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC/TON")) %>%
  select(station_id, variable, mk_p_val) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "mk_p_val") %>%
  rename(p_no3_vs_time = `NO3-N_µg/l N`, 
         p_tocton_vs_time = `TOC/TON`)

# Medians
df3 <- reg_no3 %>%
  filter(variable %in% c("NO3-N_µg/l N", "TOC_mg C/l")) %>%
  select(station_id, variable, median) %>%
  tidyr::pivot_wider(names_from = "variable", values_from = "median") %>%
  rename(NO3 = `NO3-N_µg/l N`, 
         TOC = `TOC_mg C/l`)

cat("\n")
cat("df1, n =", nrow(df1), "\n")
cat("df2, n =", nrow(df2), "\n")
cat("df3, n =", nrow(df3), "\n")

dat_1_all <- df1 %>%
  full_join(df2, by = "station_id") %>%
  full_join(df3, by = "station_id")

cat("dat_1_all, n =", nrow(dat_1_all), " (may include series without slope_no3_vs_time)\n")

dat_1 <- dat_1_all %>%
  filter(!is.na(slope_no3_vs_time))

cat("dat_1, n =", nrow(dat_1), " (all data with existing values of 'slope_no3_vs_time')\n")

```

```{r}
# dat_1
# 
# str <- "slope_no3_vs_time ~ p_no3_vs_time"
# plot(as.formula(str), data = dat_1)
# lm(as.formula(str), data = dat_1)

```

### Deposition trends and median 1992-2006     
```{r}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/deposition/totn_dep_trends_icpw_stns.csv"  

df_deposition <- read.csv(fn) %>% 
  filter(variable == "totn_mgNpm2")  

cat("n =", nrow(df_deposition), "\n")

```


### Add deposition slope and medians to data  
```{r}

# debugonce(left_join2)
dat_2 <- dat_1 %>% 
  left_join2(df_deposition %>% 
              select(station_id, median, sen_slp, mk_p_val) %>%
              rename(TOTN_dep = median,
                     slope_dep_vs_time = sen_slp,
                     p_dep_vs_time = mk_p_val),
             by = "station_id",
             print_vars = TRUE)

```

### Add medians and station metadata   

```{r}

# dat_2 <- dat_2 %>%
#   left_join(df_metadata, by = "station_id")

# cat("dat_2, n =", nrow(dat_2), "\n")

# Simplify names by removing units
# names(dat_2)
# names(dat_2) <- sub(".N_µg.l.N", "", names(dat_2))
# names(dat_2) <- sub("_mg.C.l", "", names(dat_2))
# names(dat_2) <- sub("_µg.l.P", "", names(dat_2))

# cat("\nVariable names: \n")
# names(dat_2)


```

### Add climate and deposition medians and slopes  
```{r}

fn <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/climate/cru_climate_trends_icpw_stns.csv"

df_climate_mean <- read_csv(fn) %>% 
  select(station_id, variable, median) %>%
  pivot_wider(names_from = "variable", values_from = "median")
cat("\n")
# names(df_climate_mean)

df_climate_slope <- read_csv(fn) %>%
  select(station_id, variable, sen_slp) %>%
  pivot_wider(names_from = "variable", values_from = "sen_slp", names_prefix = "slope_")

# Add
dat_3 <- dat_2 %>%
  left_join2(df_climate_mean, by = "station_id", print_vars = TRUE) %>%
  left_join2(df_climate_slope, by = "station_id", print_vars = TRUE)

```

### Combine land cover types   
* Data including UK read using script 159   
* Note: also includes metadata (country, etc.)
* bare_sparse = bare_rock + sparsely_vegetated + glacier   
* Select: coniferous, deciduous, lake, mixed_forest, wetland, bare_sparse   

```{r}

# df_landcover3_OLD <- readRDS("Data/159_df_landcover3.rds")
df_landcover3 <- readRDS("Data/159_df_meta3.rds")

df_landcover3 <- df_landcover3 %>%
  mutate(bare_sparse = bare_rock + sparsely_vegetated + glacier,
         decid_mixed = deciduous + mixed_forest,
         lake_water = lake + water_ex_lake) %>%
  select(-c(bare_rock, sparsely_vegetated, glacier, deciduous, mixed_forest, lake, water_ex_lake))

```


### Add land cover columns to main data    
```{r}

dat_4 <- left_join2(dat_3, 
                   df_landcover3, 
                   by = "station_id", 
                   print_vars = TRUE
)


```


### Drop locations with >5% cultivated and >5% urban     
- also excluding stations 23517, 38273    
```{r}

cultivated_threshold <- 5
urban_threshold <- 5

dat_5 <- dat_4 %>%
  filter2(!station_id %in% c(23517, 38273), text = "Deleted stations 23517, 38273") %>%
  filter2(cultivated <= cultivated_threshold, 
          text = paste("Deleted stations with >", cultivated_threshold, "% cultivated")) %>%
  filter2(urban <= urban_threshold, 
          text = paste("Deleted stations with >", urban_threshold, "% urban"))

```


### Data set used  
```{r}

dat <- dat_5

```


## 3. Plot slopes    

```{r, fig.width=9, fig.height=7}

ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time)) + 
  geom_point(data = dat %>% filter(p_no3_vs_time < 0.05), size = rel(2)) +
  geom_point(aes(color = country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) 
  
ggplot(dat, aes(slope_dep_vs_time, slope_no3_vs_time,
                color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries")

dat %>%
  filter(!country %in% c("Latvia","Ireland","Italy","Netherlands")) %>%
  ggplot(aes(slope_dep_vs_time, slope_no3_vs_time,
             color = (p_no3_vs_time < 0.05))) + 
  geom_point() +
  facet_wrap(vars(country)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_vline(xintercept = 0, linetype = 2) + 
  labs(title = "A selection of countries") + 
  ylim(-50, 25)


```


## 4. Select data   

### a. Selection of variables  
* Select variables to use, and thereby also cases  
* Saves data both before and after rows with removing missing predictors are removed
```{r}  

add_flag_variable <- function(data, variable_string){
  variable_string <- gsub(" ", "", variable_string)
  variables <- strsplit(variable_string, split = ",")[[1]]
  # Check if all variables are there
  found <- variables %in% names(data)
  if (sum(!found) > 0)
    stop("Not all variables found in data:", 
      paste(variables[!found], collapse = " ,"), 
      "\n")
  # Data for analyses
  complete <- complete.cases(data[variables])
  data$Row_excluded <- !complete
  variables %>% 
    purrr::map_dfr(~data.frame(Var = .x, Missing = sum(is.na(data[[.x]])))) %>%
    print()
  data
}

delete_unused_variables <- function(data, variable_string){
  variable_string <- gsub(" ", "", variable_string)
  variables <- strsplit(variable_string, split = ",")[[1]]
  data[variables]
}

cat("-------------------------------------------------------------\n")
cat("Variables: \n")
cat(params$selected_vars)
cat("\n-------------------------------------------------------------\n")

dat <- dat %>%
  filter2(!station_code %in% "PL05", text = "station PL05 (has dubious NO3 data)")

# debugonce(add_flag_variable)
# df_analysis <- add_flag_variable(dat, vars)  
df_analysis_allrows <- add_flag_variable(dat, params$selected_vars)  

# Save to excel
document_code <- stringr::str_extract(params$document_title, "([^[[:blank:]]]+)")   # extract e.g. "164a1"
fn <- paste0(document_code, "_", params$response_variable, "_data.xlsx")
writexl::write_xlsx(df_analysis_allrows, paste0("Data_analysed/", fn))
cat("\nDataset after removing urban, cultivated, PL05 saved as", sQuote(fn), "\n\n")

cat("Number of rows that will be excluded: \n")
table(df_analysis_allrows$Row_excluded)

cat("\n\n")
cat("Number of complete observations by country: \n")
xtabs(~country + Row_excluded, df_analysis_allrows)

# Keep only complete cases
df_analysis <- df_analysis_allrows %>%
  filter(!Row_excluded)

# Save to excel
fn <- paste0(
  stringr::str_extract(params$document_title, "[^[[:blank:]]]+"),
  "_data.xlsx")
writexl::write_xlsx(df_analysis, paste0("Data_analysed/", fn))

# Remove variables that will note be used
df_analysis <- delete_unused_variables(df_analysis, params$selected_vars)

cat("\n\n")
cat("Data before removing PL05: n =", nrow(dat_5), "\n")
cat("Data after removing PL05: n =", nrow(df_analysis_allrows), "\n")
cat("Data after removing missing predictors: n =", nrow(df_analysis), "\n")


```


### b. Correlations   
```{r, fig.width=10, fig.height=10}

gg <- GGally::ggcorr(
  df_analysis, 
  method = c("complete.obs", "kendall"), 
  label = TRUE,
  hjust = 0.9, angle = -30) # +                    # slanted labels
gg + coord_cartesian(x = c(-2, 20), y = c(-2,22))  # fix margins


```  



## 5. Tree and forest classification


### Split into training and validation data
```{r}

set.seed(123)

x <- runif(nrow(df_analysis))
train <- ifelse(x < 0.9, TRUE, FALSE)

train_set <- df_analysis[train,]  %>% 
  # mutate(no3_decline_f = factor(no3_decline)) %>% 
  select(-longitude, - latitude) %>%
  as.data.frame()
valid_set <- df_analysis[!train,] %>% 
  # mutate(no3_decline_f = factor(no3_decline)) %>% 
  select(-longitude, - latitude) %>%
  as.data.frame()

```



### c. Random forest  
* *For results/interpretation, see separate document '160_randomforest_James_data.html'*  


#### c1a. Predict on training data
```{r}  

model1 <- randomForest(slope_no3_vs_time ~ ., 
                       data = train_set, 
                       mtry = 5,
                       importance = TRUE)

# Predicting in training set
df_plot <- data.frame(
  Predicted = predict(model1, train_set),
  Observed = train_set[[params$response_variable]])
ggplot(df_plot, aes(Predicted, Observed)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  labs(title = "Predictions in training set")

cat("R-square, predicting in training set: \n")
summary(lm(Observed ~ Predicted, data = df_plot))$r.sq

# Predicting in validation set
df_plot <- data.frame(
  Predicted = predict(model1, valid_set),
  Observed = valid_set[[params$response_variable]])
ggplot(df_plot, aes(Predicted, Observed)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  labs(title = "Predictions in validation set")

cat("R-square, predicting in validation set: \n")
summary(lm(Observed ~ Predicted, data = df_plot))$r.sq


```

#### c1b. Model for all data    
```{r}  

full_set <- df_analysis  %>% 
  # mutate(no3_decline = factor(no3_decline)) %>% 
  select(-longitude, - latitude) %>%
  as.data.frame()

model1 <- randomForest(slope_no3_vs_time ~ ., 
                       data = full_set, 
                       mtry = 5,
                       importance = TRUE)


# Prediction
df_plot <- data.frame(
  Predicted = predict(model1, full_set),
  Observed = full_set[[params$response_variable]])
ggplot(df_plot, aes(Predicted, Observed)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  labs(title = "Predictions in entire data set")

cat("R-square, predicting in entire data set: \n")
summary(lm(Observed ~ Predicted, data = df_plot))$r.sq


```

#### c2. Importance of variables
```{r 5c2_calculate_importance}

# Calculation
importance <- measure_importance(model1)

```



```{r}


plot_multi_way_importance(importance, size_measure = "no_of_nodes", no_of_labels = 6)  

if (FALSE){
  # For classification (discrete response variable)
  plot_multi_way_importance(importance, 
                            x_measure = "accuracy_decrease", 
                            y_measure = "gini_decrease", 
                            size_measure = "p_value", no_of_labels = 6)
} else {
  
  # For regression (continuous response variable)
  plot_multi_way_importance(importance, 
                            x_measure = "mse_increase", 
                            y_measure = "node_purity_increase",
                            size_measure = "p_value", no_of_labels = 6)
}

importance %>% 
  arrange(times_a_root)


```



#### c3. Random forest, show partial effects  

```{r 5c3_calculate_partial_effects1, warning=FALSE, message=FALSE, results='hide'}

# Which variables to include:
variables_for_plot <- importance %>%
  mutate(variable = levels(variable)[as.numeric(variable)]) %>%
  arrange(desc(times_a_root)) %>%
  pull(variable) %>%
  head(12)   # pick the first 12 variables (or less)

# Calculation

plotdata <- NULL  # will be list for storing results

max_number_of_plots <- length(variables_for_plot)/2 %>% floor()

for (i in 1:max_number_of_plots){
  varno1 <- c(1,3,5,7,9,11)[i]
  varno2 <- varno1 + 1
  plotdata[[i]] <- model1 %>%
    partial(pred.var = variables_for_plot[c(varno1, varno2)], chull = TRUE, progress = "text",
            which.class = "1", prob = TRUE)
}

```


```{r 5c3_calculate_partial_effects2, warning=FALSE, message=FALSE, results='hide'}

### Extra plots

i <- max_number_of_plots

plotpairs <- params$extra_pairwise_plots %>%
  gsub(" ", "", ., fixed = TRUE) %>%
  strsplit(split = ";") %>%
  .[[1]] %>%
  purrr::map(~strsplit(., split = ",")[[1]])

for (plotvar in plotpairs){
  # print(plotvar)
  if (plotvar[1] %in% names(full_set) & plotvar[2] %in% names(full_set)){
    i <- i + 1
    plotdata[[i]] <- model1 %>%
      partial(pred.var = c(plotvar[1], plotvar[2]), chull = TRUE, progress = "text",
             which.class = "1", prob = TRUE)
  }
}


```

```{r 5c3_plot_partial_effects2}

# Find range of predicted values for each graph
ranges <- plotdata %>% purrr::map_dfc(~range(.$yhat))

# use range of all the ranges
for (i in 1:length(plotdata)){
  
  if (params$pairwise_plots_same_scale == "TRUE"){
  
    gg <- plot_pair_number(i, zrange = range(ranges))
    print(gg)
    
  } else {
    
    gg <- plot_pair_number(i)
    print(gg)

  }
  
  # Save gg object for later plotting / changes
  # Saved in Figures/Partial_plots' with name e.g. "gg_164a1_7.rds" for plot number 7
  fn <- paste0(
    "Figures/Partial_plots/gg_",
    stringr::str_extract(params$document_title, "([^[[:blank:]]]+)"),   # extract e.g. "164a1"
    "_", i, ".rds")
  saveRDS(gg, fn)
  
}
  

```
