---
title: "180_fix_partial_plots"
output: html_document
---

Fixing figures  

## Chapters, figure numbers and model runs used for each    

### 3.1.2	Spatial variability of median NO3 concentrations 2012-2016  
* Figures 164ab1 (median NO3)   
    * Fig. 11: plot no. 7  
    * Fig. 12: plot no. 1  

### 3.1.3	Statistical analysis of NO3 trends 1992-2016  
* Script 166d1x   
    * Fig. 15: plot no. 1  
    * Fig. 16: plot no. 2  
    * Fig. 17: plot no. 7  

### 3.2.2	Spatial variability of TOC/TON median concentration ratios 2012-2016   
* Script 164c1   
    * Fig. 26: plot no. 8  
    * Fig. 27: plot no. 9  
    * Fig. 29: plot no. 7    

### 3.2.3 TOC/TON trends 1992-2016  
* Script 167d1     
    * Fig. 32: plot no. 1  
    * Fig. 33: plot no. 2    
    


## Units  
* N deposition: mgN m(-2)
* N deposition change:  mgN m(-2) yr(-2)
* NO3 concentration median: ug l(-1)
* NO3 concentration change: ug l(-1) yr(-2)  
* TOC concentration median: mg l(-1)  

### Define units/terms including unicode characters  
```{r}

per_sqm <- "m\u207B\u00B2"
per_year <- "yr\u207B\u00B9"
per_yearsquared <- "yr\u207B\u00B2"   # for Sen slope of deposition
per_litre <- "l\u207B\u00B9"
microgram <- "\u03BCg"
no3 <- "NO\u2083"
degrees_celcius <- "C\u00B0"

per_sqm
per_year
per_yearsquared
per_litre
microgram
no3
degrees_celcius

```


## Packages  
```{r}

# library(dplyr)
library(ggplot2)
library(glue)

```

### Defining special characters  
```{r}

name <- "Fred"
age <- 50
anniversary <- as.Date("1991-10-12")

cat("- Test glue() -\n")
glue('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.')


```


## 3.1.2	Spatial variability of median NO3 concentrations 2012-2016  

* *Important: run this chunk before the next chunks*      
* Figures 164ab1 (median NO3)   
    * Fig. 11: plot no. 7  
    * Fig. 12: plot no. 1
    
```{r}

fig_folder <- "Figures/Partial_plots"               # rds files
fig_folder_new <- "Figures/Partial_plots_fixed"     # png files
fn_core <- "gg_164ab1_"
fns <- dir(fig_folder, pattern = fn_core)

# Show all plots
for (i in 1:length(fns)){
  fn <- paste0(fig_folder, "/", fn_core, i,".rds")
  gg <- readRDS(fn)
  # uncomment to show all plots:
  print(gg)
  }

```

### Fig. 11 - plot no. 7   

```{r}

# dir.create("Figures/Partial_plots_fixed")

i <- 7
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)
gg

# legend: Median NO3 concentration (µg N L-1)
# x-akse: mg N m-2 yr-1
# y-akse: mg N m-2 yr-1

gg_new <- gg +
  labs(x = glue("Median N deposition 1992-2016 (mg N {per_sqm} {per_year})"),
       y = glue("Sen slope N deposition (mg N  {per_sqm} {per_year})")) +
  scale_fill_viridis_c(glue("Median {no3} concentration\n({microgram} N {per_litre})"))

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg_new

```


### Fig. 12 - plot no. 1

```{r}

# dir.create("Figures/Partial_plots_fixed")

i <- 1
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)
gg

gg_new <- gg +
  labs(x = glue("Median N deposition 1992-2016 (mg N {per_sqm} {per_year})"),
       y = glue("Air temperature ({degrees_celcius})")) +
  scale_fill_viridis_c(glue("Median {no3} concentration\n({microgram} N {per_litre})"))

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg_new

```


## 3.1.3	Statistical analysis of NO3 trends 1992-2016  

* *Important: run this chunk before the next chunks*      
* Script 166d1x   
    * Fig. 15: plot no. 1  
    * Fig. 16: plot no. 2  
    * Fig. 17: plot no. 7  
    
```{r}

fig_folder <- "Figures/Partial_plots"               # rds files
fig_folder_new <- "Figures/Partial_plots_fixed"     # png files
fn_core <- "gg_166d1x_"
fns <- dir(fig_folder, pattern = fn_core)

# Show all plots
for (i in 1:length(fns)){
  fn <- paste0(fig_folder, "/", fn_core, i,".rds")
  gg <- readRDS(fn)
  # uncomment to show all plots:
  print(gg)
  }

```

### Fig. 15: plot no. 1   
Text in paper says that the x axis is 2012-2016 medians, not 1992-2016. 1992-2016 seems to be correct.

    * 

```{r}

# From the code of script 166:
# Source of 'reg_no3' is (line 111):     
#    https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3    
#    file: "trends_1992-2016_no3_results.csv   
# Variable NO3 is comes from 'reg_no3' (line 201) 
    
# dir.create("Figures/Partial_plots_fixed")

i <- 1
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)
gg

gg_new <- gg +
  labs(x = glue("Median NO\u2083 concentration 1992-2016 (\u03BCg N {per_litre})"),
       y = "Coniferous land cover (%)") +
  scale_fill_viridis_c("NO\u2083 trend\n(Sen slope, \u03BCg N l\u207B\u00B9 yr\u207B\u00B9)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg_new

```


### Fig. 16: plot no. 2
```{r}

# dir.create("Figures/Partial_plots_fixed")

i <- 2
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(
    x = glue("Sen slope N deposition 1992-2016 (mg N {per_sqm} {per_yearsquared})"),
    y = glue("Median N deposition 1992-2016 (mg N {per_sqm} {per_year})")
  ) +
  scale_fill_viridis_c("NO\u2083 trend\n(Sen slope, \u03BCg N l\u207B\u00B9 yr\u207B\u00B9)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new


```


### Fig. 17: plot no. 7

```{r}

# dir.create("Figures/Partial_plots_fixed")

i <- 7
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(
    x = glue("Median TOC concentration 1992-2016 (mg C {per_litre})"),
    y = glue("Median {no3} concentration 1992-2016 ({microgram} N {per_litre})")
    ) +
  scale_fill_viridis_c("NO\u2083 trend\n(Sen slope, \u03BCg N l\u207B\u00B9 yr\u207B\u00B9)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```


## 3.2.2	Spatial variability of TOC/TON median concentration ratios 2012-2016   

* *Important: run this chunk before the next chunks*      
* Script 164c1   

    * Fig. 27: plot no. 8  
    * Fig. 28: plot no. 2  
    * Fig. 30: plot no. 7  
    
```{r}

fig_folder <- "Figures/Partial_plots"               # rds files
fig_folder_new <- "Figures/Partial_plots_fixed"     # png files
fn_core <- "gg_164c1_"
fns <- dir(fig_folder, pattern = fn_core)

# Show all plots
for (i in 1:length(fns)){
  fn <- paste0(fig_folder, "/", fn_core, i,".rds")
  gg <- readRDS(fn)
  # uncomment to show all plots:
  print(gg)
  }

```

### Fig. 27: plot no. 8  

```{r}

# From the code of script 166:
# Source of 'reg_no3' is (line 111):     
#    https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3    
#    file: "trends_1992-2016_no3_results.csv   
# Variable NO3 is comes from 'reg_no3' (line 201) 
    
# dir.create("Figures/Partial_plots_fixed")

i <- 8
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(
    x = glue("Median TOC concentration 2012-2016 (mg C {per_litre})"),
    y = glue("Median TON concentration 2012-2016 ({microgram} N {per_litre})")
    ) +
  scale_fill_viridis_c("TOC/TON ratio\n(unitless)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```

### Fig. 28: plot no. 9  

```{r}

# From the code of script 166:
# Source of 'reg_no3' is (line 111):     
#    https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3    
#    file: "trends_1992-2016_no3_results.csv   
# Variable NO3 is comes from 'reg_no3' (line 201) 
    
# dir.create("Figures/Partial_plots_fixed")

i <- 9
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(x = "Coniferous land cover (%)",
       y = "Lake/water land cover (%)") +
  scale_fill_viridis_c("TOC/TON ratio\n(unitless)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```


### Fig. 30: plot no. 7  

```{r}

# From the code of script 166:
# Source of 'reg_no3' is (line 111):     
#    https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3    
#    file: "trends_1992-2016_no3_results.csv   
# Variable NO3 is comes from 'reg_no3' (line 201) 
    
# dir.create("Figures/Partial_plots_fixed")

i <- 7
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(
    x = glue("Median N deposition 1992-2016 (mg N {per_sqm} {per_year})"),
    y = glue("Sen slope N deposition 1992-2016 (mg N {per_sqm} {per_yearsquared})")
    ) +
  scale_fill_viridis_c("TOC/TON ratio\n(unitless)")

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```


## 3.2.3 TOC/TON trends 1992-2016  

* *Important: run this chunk before the next chunks*      
* Script 167d1     

    * Fig. 33: plot no. 1  
    * Fig. 34: plot no. 2    
    
```{r}

fig_folder <- "Figures/Partial_plots"               # rds files
fig_folder_new <- "Figures/Partial_plots_fixed"     # png files
fn_core <- "gg_167d1_"
fns <- dir(fig_folder, pattern = fn_core)

# Show all plots
for (i in 1:length(fns)){
  fn <- paste0(fig_folder, "/", fn_core, i,".rds")
  gg <- readRDS(fn)
  # uncomment to show all plots:
  print(gg)
  }

```

### Fig. 33: plot no. 1    

```{r}

i <- 1
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

gg_new <- gg +
  labs(
    x = glue("Sen slope TON concentration ({microgram} N {per_litre} {per_year})"),
    y = "Coniferous land cover (%)"
    ) +
  scale_fill_viridis_c(glue("TOC/TON trend\n(Sen slope, ratio {per_year})"))

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```

### Fig. 34: plot no. 2    

```{r}

i <- 2
fn <- paste0(fig_folder, "/", fn_core, i,".rds")
gg <- readRDS(fn)

#     x = glue("Median TOC concentration 1992-2016 (mg {per_litre})"),

gg_new <- gg +
  labs(
    x = glue("Median TOC concentration 1992-2016 (mg C {per_litre})"),
    y = glue("Sen slope TOC concentration (mg C {per_litre} {per_year})")
    ) +
  scale_fill_viridis_c(glue("TOC/TON trend\n(Sen slope, ratio {per_year})"))

fn_new <- paste0(fig_folder_new, "/", fn_core, i,"_fixed.png")
ggsave(fn_new, gg_new, width = 6, height = 4, dpi = 200)

gg
gg_new

```


