---
title: "903_Add_info_to_data_analysed"
output: html_document
---

* The saved excel files in folder 'Data_analysed' lacked some columns (sen_slp, sen_incpt, sen_trend).  
* This scripts adds those data (using left join) and makes new excel files in folder 'Data_analysed2'  

## 1. Packages
```{r}

library(readxl)
library(writexl)
library(dplyr)

# dir.create("Data_analysed2")

```


## 2. Fix files in 'Data_analysed'     

- Read files in 'Data_analysed' (made using scripts 160parm, 161parm, 162parm)  
- Add sen_slp, sen_incpt, sen_trend  
- Write to 'Data_analysed2'  

### a. Output from script 160parm  

#### James' trends and medians     
- Code from original script  
```{r}

#
# Regression results
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_no3"
file <- "trends_1992-2016_no3_results.csv"
fn <- paste0(folder, "/", file)

reg_no3 <- read.csv(fn, encoding = "UTF-8")
cat("Regression results:", sQuote(file), ",n =", nrow(reg_no3), "\n\n")


```
#### Fix files   
- Read files in 'Data_analysed'   
- Add sen_slp, sen_incpt, sen_trend  
- Write to 'Data_analysed2'  
```{r}

# names(reg_no3)

dat_to_add <- reg_no3 %>%
  filter(variable == 'NO3-N_µg/l N') %>%
  select(station_id, sen_slp, sen_incpt, sen_trend, mk_p_val)

# check
# dat_to_add %>% count(station_id) %>% xtabs(~n, .)

filenames <- dir("Data_analysed", pattern = "160.*")

for (filename in filenames){
  # filename <- filenames[1]    # for testing
  dat <- read_excel(paste0("Data_analysed/", filename))
  dat_updated <- dat %>%
    left_join(dat_to_add, by = "station_id") %>%
    select(station_id, sen_slp, sen_incpt, sen_trend, mk_p_val, no3_decline, everything())
  write_xlsx(dat_updated, paste0("Data_analysed2/", filename))
}


```


### b. Output from script 161parm  

#### James' trends and medians     
- Code from original script  
```{r}

#
# Regression results
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/trends_1992-2016_toc_totn_no3_relax_italy"
file <- "trends_1992-2016_toc_totn_no3_relax_italy_results.csv"
fn <- paste0(folder, "/", file)

reg_ratio <- read.csv(fn, encoding = "UTF-8")
cat("Regression slopes and medians from:", sQuote(file), ",n =", nrow(reg_ratio), "\n\n")


```


#### Fix files   
```{r}

# names(reg_no3)

dat_to_add <- reg_ratio %>%                                       # changed compared to a
  filter(variable == 'TOC/TON') %>%                               # changed compared to a
  select(station_id, sen_slp, sen_incpt, sen_trend, mk_p_val)

# check
# dat_to_add %>% count(station_id) %>% xtabs(~n, .)

filenames <- dir("Data_analysed", pattern = "161.*")              # changed compared to a

for (filename in filenames){
  # filename <- filenames[1]    # for testing
  dat <- read_excel(paste0("Data_analysed/", filename))
  dat_updated <- dat %>%
    left_join(dat_to_add, by = "station_id") %>%
     # changed compared to a: 'tocton_decrease':   
    select(station_id, sen_slp, sen_incpt, sen_trend, mk_p_val, tocton_decrease, everything())  
  write_xlsx(dat_updated, paste0("Data_analysed2/", filename))
}


```


### c. Output from script 162parm a + b 

#### James' trends and medians     
- Code from original script    
```{r}

medians_filename <- "medians_2012-2016_no3.csv"   # for a and b files - see 162parm_run_markdown.R

#
# Median results 2012-2016  
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
file <- medians_filename
fn <- paste0(folder, "/", file)
df_medians <- read.csv(fn, encoding = "UTF-8")
cat("Medians results:", sQuote(file), ",n =", nrow(df_medians), "\n\n")  
cat("Number of values per variable: \n")
apply(!is.na(df_medians), 2, sum)
cat("\n")

```
#### Fix files (162a, 162b)    
```{r}

# names(reg_no3)

dat_to_add <- df_medians %>%                                       # changed compared to a
                                                                   # changed compared to a
  select(station_id, `NO3.N_µg.l.N`)                               # changed compared to a

# check
# dat_to_add %>% count(station_id) %>% xtabs(~n, .)

filenames <- dir("Data_analysed", pattern = "162[ab].*")              # changed compared to a

for (filename in filenames){
  # filename <- filenames[1]    # for testing
  dat <- read_excel(paste0("Data_analysed/", filename))
  dat_updated <- dat %>%
    left_join(dat_to_add, by = "station_id") %>%
     # changed compared to a: 'NO3.N_µg.l.N':   
    select(station_id, `NO3.N_µg.l.N`, everything())  
  write_xlsx(dat_updated, paste0("Data_analysed2/", filename))
}


```



### d. Output from script 162parm c + d 

#### James' trends and medians     
- Code from original script    
```{r}

medians_filename <- "medians_2012-2016_no3.csv"            # for a and b files - see 162parm_run_markdown.R
medians_filename <- "medians_2012-2016_toc_totn_no3.csv"   # for c and d files - see 162parm_run_markdown.R

#
# Median results 2012-2016  
#
folder <- "https://github.com/JamesSample/icpw2/raw/master/thematic_report_2020/results/medians_2012-2016"
file <- medians_filename
fn <- paste0(folder, "/", file)
df_medians <- read.csv(fn, encoding = "UTF-8")
cat("Medians results:", sQuote(file), ",n =", nrow(df_medians), "\n\n")  
cat("Number of values per variable: \n")
apply(!is.na(df_medians), 2, sum)
cat("\n")

```
#### Fix files (162c, 162d)    
```{r}

# names(reg_no3)

dat_to_add <- df_medians %>%                                       # changed compared to a
                                                                   # changed compared to a
  select(station_id, `TOC_mg.C.l`, `TOTN_µg.l.N`, `TOC.TON`)                               # changed compared to a

# check
# dat_to_add %>% count(station_id) %>% xtabs(~n, .)

filenames <- dir("Data_analysed", pattern = "162[cd].*")              # changed compared to a

for (filename in filenames){
  # filename <- filenames[1]    # for testing
  dat <- read_excel(paste0("Data_analysed/", filename))
  dat_updated <- dat %>%
    left_join(dat_to_add, by = "station_id") %>%
     # changed compared to a: 'NO3.N_µg.l.N':   
    select(station_id, `TOC_mg.C.l`, `TOTN_µg.l.N`, `TOC.TON`, everything())  
  write_xlsx(dat_updated, paste0("Data_analysed2/", filename))
}


```
